---
title: "Authoritarianism index"
author: "Craig Jolley"
date: "March 16, 2016"
output: html_document
---

This index will measure the a respondant's inclination toward authoritarian political attitudes. The following variables were measured in all three countries and correlate strongly with each other:

- `dem2`: Support for democracy or dictatorship (1=either is OK, 2=democracy, 3=dictatorship)
- `dem11`: Support for mano dura policies (1=iron fist, 2=broad participation)
- `jc10`: Coup is justified when crime is high (1=yes, 2=no)
- `jc13`: Coup is justified when corruption is high (1=yes, 2=no)
- `jc15a`: President justified in governing without legislature during crisis (1=yes, 2=no)

To start off, we'll load the necessary packages:

```{r, message=FALSE}
library(ggplot2)
library(mice)
library(plyr) 
library(GGally) 
```

We'll load all three 2014 datasets and construct a data frame containing the columns we want.

```{r}
setwd("C:/Users/Craig/Desktop/SMA/VSFS/LAPOP-SMA/indices")
lapop.2014.GTM <- read.csv('../../2014-GTM.csv')
lapop.2014.SLV <- read.csv('../../2014-SLV.csv')
lapop.2014.HND <- read.csv('../../2014-HND.csv')
v <- c('dem2','dem11','jc13','jc10','jc15a')
my_data <- rbind(lapop.2014.GTM[,v],lapop.2014.SLV[,v],lapop.2014.HND[,v])
#dem2 has sort of an odd scale -- switch 1 and 2
my_data$dem2[my_data$dem2==1] <- 100
my_data$dem2[my_data$dem2==2] <- 1
my_data$dem2[my_data$dem2==100] <- 2
is.na(my_data[my_data>16]) <- TRUE
```

We'll use multiple imputation to fill in missing data, but let's omit missing values first to be sure that everything makes sense.

```{r}
my_complete <- na.omit(my_data)
pr_complete <- prcomp(my_complete,center=TRUE,scale=FALSE)
plot(pr_complete)
summary(pr_complete)
```

41% of variance is concentrated in the first component.

```{r}
set.seed(12345)
my_imp <- mice(my_data,printFlag = F) 
my_imp
```

Most variables had fairly similar numbers of missing values; people were least willing to talk about expanded presidential powers (`jc15a`) and most willing to talk about mano dura policies (`dem11`).

What does PCA look like for one of the five imputed datasets?

```{r}
pr <- lapply(1:5,function(x) prcomp(complete(my_imp,x),scale=FALSE,center=TRUE))
plot(pr[[1]])
summary(pr[[1]])
```

PCA results look very similar to what we saw without imputation, which is good.

```{r, fig.width=10, fig.height=10}
pc1 <- pr[[1]]$x[,1]
pc2 <- pr[[1]]$x[,2]
qplot(pc1,pc2) + theme_classic()
```

The distribution of PC1 seems to fall into three bands, suggesting that the index is dominated by one variable that takes on three values.

```{r, fig.width=10, fig.height=10}
all_pc1 <- data.frame(llply(1:5, function(i) pr[[i]]$x[,1]))
names(all_pc1) <- c('imp1','imp2','imp3','imp4','imp5')
ggpairs(all_pc1) + theme_classic()
```

All of our imputations are really well-correlated, and the lowest-valued band is the most heavily-populated one.

```{r, fig.width=10, fig.height=5}
all_pc1$avg <- rowMeans(all_pc1)
q25 <- quantile(all_pc1$avg)[2]
q75 <- quantile(all_pc1$avg)[4]
qchange <- data.frame(low=colMeans(my_data[all_pc1$avg <= q25,],na.rm=TRUE),
                      hi=colMeans(my_data[all_pc1$avg >= q75,],na.rm=TRUE))
qchange$slope <- qchange$hi - qchange$low
qchange$label <- rownames(qchange)
qchange$even <- rank(qchange$hi)/nrow(qchange)*max(qchange$hi)
ggplot(data=qchange) +
  geom_segment(aes(x=0,y=qchange$low,xend=1,yend=qchange$hi,
                   color=qchange$slope),size=1) +
  scale_color_gradient(low='#FF420E',high='#89DA59') +
  geom_text(aes(x=1,y=(qchange$even+qchange$hi)/2,label=qchange$label,
                hjust=0,color=qchange$slope),size=5) +
  annotate("text",x=0,y=0,label='Average in lowest quartile',hjust=0) +
  annotate("text",x=1,y=0,label='Average in highest quartile',hjust=1) +
  scale_x_continuous(limits=c(0,1.5)) +
  theme_classic() +
  theme(legend.position='none',
        axis.ticks=element_blank(),
        axis.line=element_blank(),
        axis.text.x=element_blank(),
        axis.title=element_blank(),
        text=element_text(size=20))
```

What are we seeing here? Compared to the lowest quantile, the highest quantile of our proposed index has much higher values for `dem2` and decreasing values for everything else. Recall that lower values for `dem2` indicate a preference for democracy, while higher values of the binary variables indicate less-authoritarian attitudes. Our scale in pointing in the right direction.

Also note that `dem2` is the only variable that takes on three values; its strong slope in this plot suggests that it is the dominant factor in our index.

Let's normalize and average PC1:

```{r}
all_pc1$norm <- scale(all_pc1$avg)
```

To confirm our suspicions about `dem2`, let's see how all variables contribute to the composite score.

```{r,warning=FALSE}
predict_data <- data.frame(dem2=c(1,3),dem11=c(2,1),jc13=c(2,1),
                          jc10=c(2,1),jc15a=c(2,1))
# The first row of predict_data should be the least-authoritarian response
# possible, while the second is the most-authoritarian.
minmax <- rowMeans(sapply(1:5, function(i) predict(pr[[i]],predict_data)[,1]))
minmax <- (minmax-mean(all_pc1$avg)) / sd(all_pc1$avg)
# should match quantile(all_pc1$norm)
predict_data2 <- data.frame(2-diag(5))
names(predict_data2) <- names(my_data)
predict_data2[,'dem2'] <- 3 - predict_data2[,'dem2']
scores <- rowMeans(sapply(1:5, function(i) 
  predict(pr[[i]],predict_data2)[,1]))
scores <- (scores-mean(all_pc1$avg)) / sd(all_pc1$avg)
diff <- data.frame(d=(scores - minmax[1]),n=names(predict_data))
ggplot(diff,aes(x=n,y=d)) +
  geom_bar(stat='identity',fill='skyblue') +
  coord_flip() +
  theme_classic() +
  theme(text=element_text(size=20),
        axis.title.y=element_blank(),
        axis.title.x=element_blank())
```

While all five variables have an impact on the composite index, `dem2` has by far the strongest influence.

We can be more rigorous about this by calculating p-values. We'll have to do this differently for `dem2` (which takes on three values) than for the ones that only take on two. 

```{r}
bin_cor <- function(var2) {
  # Determine the correlation between all_pc1$norm and a binary variable 
  # var2. This uses a Fisher test conditioned on the values of var2.
  tmp <- data.frame(v1=all_pc1$norm,v2=my_data[,var2])
  tmp$v2 <- tmp$v2 - min(tmp$v2,na.rm=TRUE) # convert to 0-1 scale
  if (sum(tmp$v2==0,na.rm=TRUE) > 1 & sum(tmp$v2==1,na.rm=TRUE) > 1) {
    tt <- t.test(tmp$v1[tmp$v2==0],tmp$v1[tmp$v2==1])
    res=data.frame(var=var2,est=tt$estimate[2]-tt$estimate[1],pval=tt$p.value,
                 stringsAsFactors=FALSE)
  }
  res
}

my_data$dem2_lo <- as.numeric(my_data$dem2 >= 2)
my_data$dem2_hi <- as.numeric(my_data$dem2 > 2)
bin_vars <- c('dem2_lo','dem2_hi','dem11','jc13','jc10','jc15a')
ldply(bin_vars,bin_cor)
```

All are significantly-correlated (especially `dem2`).