---
title: "Community Activity index"
author: "Jorge Perez"
output: html_document
---

The following is a proposed index, measuring the level of civic and political participation, and civil socity involvement of the individuals interrogated. Variables such as 'Attendance to religious meetings' and 'Sports Activities' were excluded, due to them not directly falling into the categories previously listed. In all, do individuals participate in making their community's social and political environment better? If so, how often?

* `cp5`: Tried to solve a community problem (1=Once a week, 2=Once or twice a month, 3=Once or twice a year, 4=Never)
* `cp7`: Attended meetings of Parent Association 
* `cp8`: Attended meetings of Community Improvement Association 
* `cp13`: Attended meetings of a Political Party 
* `cp20`: Attended meetings of Women's Group 
* `honcp22`: Help for security groups 
* `honcp21a`: Help for neighborhood associations 

The two latter elements are particular to Honduras, and are not present for the data on El Salvador and Guatemala. 

```{r, message=FALSE}
library(ggplot2)
library(mice)
library(plyr) # nice for re-formatting data
library(GGally) # for plot matrices

setwd("C:/Users/Craig/Desktop/SMA/VSFS/LAPOP-SMA/indices")
lapop.2014.HND <- read.csv("../../HND-2014.csv")
my_data <- lapop.2014.HND[,c('cp5','cp7','cp8','cp13','cp20','honcp22','honcp21a')]
is.na(my_data[my_data>16]) <- TRUE
nrow(na.omit(my_data))
my_data <- my_data[,order(names(my_data))] # alphabetize columns
```
Ignoring missing values, we are looking at 762 rows left

```{r}
sum(is.na(my_data)) / (nrow(my_data)*ncol(my_data))
```
7.5% of values missing!

```{r}
my_complete <- na.omit(my_data)
pr_complete <- prcomp(my_complete,center=TRUE,scale=FALSE)
plot(pr_complete)
summary(pr_complete)
```
Without any imputation, 32.7% of variance is concentrated on the first principal component.

```{r}
my_imp <- mice(my_data,printFlag = F,seed=12345) 
nmis <- my_imp$nmis
nm <- data.frame(nmis=nmis/nrow(my_data),var=names(nmis))
ggplot(nm,aes(x=var,y=nmis)) +
  geom_bar(stat = 'identity', fill = '#90AFC5') +
  ylab('Fraction of imputed values') +
  coord_flip() +
  theme_classic() +
  theme(legend.position='none',
        axis.ticks=element_blank(),
        axis.line=element_blank(),
        axis.title.y=element_blank(),
        text=element_text(size=20))
```

None of the variables stand out as needing too much imputation, with the CLEAR exception of `cp20`, which was question conditioned on the subject being female. Presumably, what the imputation is telling us is whether the individual in question is the kind of guy who would go to women's groups if men were invited.

```{r}
pr <- lapply(1:5,function(x) prcomp(complete(my_imp,x),scale=FALSE,center=TRUE))
plot(pr[[1]])
summary(pr[[1]])
```
With the imputation, we don't have much change. 32.5% of variance is concentrated on the first principal component.

```{r}
pc1 <- pr[[1]]$x[,1]
pc2 <- pr[[1]]$x[,2]
qplot(pc1,pc2) + theme_classic()
```

```{r}
all_pc1 <- data.frame(llply(1:5, function(i) pr[[i]]$x[,1]))
names(all_pc1) <- c('imp1','imp2','imp3','imp4','imp5')
ggpairs(all_pc1) + theme_classic()
```

We have a problem here; some of our imputed datasets differ from each other by a sign change. We'll have to flip these so that all have a high correlation with `imp1`, rather than an anticorrelation.

```{r}
for (i in 2:ncol(all_pc1)) {
  if (cor(all_pc1[,1],all_pc1[,2]) < 0) {
    all_pc1[,2] <- -1 * all_pc1[,2]
  }
}
ggpairs(all_pc1) + theme_classic()
```

Much better.

```{r,fig.height=5,fig.width=10}
all_pc1$avg <- rowMeans(all_pc1)
qchange <- data.frame(low=colMeans(my_data[all_pc1$avg < quantile(all_pc1$avg)[2],],na.rm=TRUE),
                      hi=colMeans(my_data[all_pc1$avg > quantile(all_pc1$avg)[4],],na.rm=TRUE))
qchange$slope <- qchange$hi - qchange$low
qchange$label <- rownames(qchange)
qchange$even <- rank(qchange$hi)/nrow(qchange)*max(qchange$hi)
ggplot(data=qchange) +
  geom_segment(aes(x=0,y=qchange$low,xend=1,yend=qchange$hi,
                   color=qchange$slope),size=1) +
  scale_color_gradient(low='#FF420E',high='#89DA59') +
  geom_text(aes(x=1,y=(qchange$even+qchange$hi)/2,label=qchange$label,
                hjust=0,color=qchange$slope),size=5) +
  annotate("text",x=0,y=0,label='Average in lowest quartile',hjust=0) +
  annotate("text",x=1,y=0,label='Average in highest quartile',hjust=1) +
  scale_x_continuous(limits=c(0,1.5)) +
  theme_classic() +
  theme(legend.position='none',
        axis.ticks=element_blank(),
        axis.line=element_blank(),
        axis.text.x=element_blank(),
        axis.title=element_blank(),
        text=element_text(size=20))
```

In the lowest quartile, nearly everyone answered '4', indicating no participation at all. Among the most engaged quartile, parents' groups (`cp7`) appear to be the most popular community activity. Our index appears to be pointing in the right direction.

```{r}
all_pc1$norm <- scale(all_pc1$avg) 
predict_data <- data.frame(cp5=c(4,1),cp7=c(4,1),cp8=c(4,1),
                           cp13=c(4,1),cp20=c(4,1),honcp22=c(4,1),
                           honcp21a=c(4,1))
# The first row is the responses that would come from the least-engaged person
# possible; the second row is the responses from a maximally-engaged person.
minmax <- rowMeans(sapply(1:5, function(i) predict(pr[[i]],predict_data)[,1]))
minmax <- (minmax - mean(all_pc1$avg)) / sd(all_pc1$avg) 
# Comparing minmax to quantile(all_pc1$norm), we can see that some people are
# pretty close to the minimum level of engagement, and no one is near the 
# maximum.

predict_data2 <- data.frame(4-diag(7))
names(predict_data2) <- names(my_data)
scores <- rowMeans(sapply(1:5, function(i) predict(pr[[i]],predict_data2)[,1]))
scores <- (scores-mean(all_pc1$avg)) / sd(all_pc1$avg) 
diff <- data.frame(d=(scores - minmax[1]),n=names(predict_data))
ggplot(diff,aes(x=n,y=d)) +
  geom_bar(stat='identity',fill='skyblue') +
  coord_flip() +
  theme_classic() +
  theme(text=element_text(size=20),
        axis.title.y=element_blank(),
        axis.title.x=element_blank())
```

All variables have a strong influence on the index.

```{r}
lmp <- function (modelobject) {
    if (class(modelobject) != "lm") stop("Not an object of class 'lm' ")
    f <- summary(modelobject)$fstatistic
    p <- pf(f[1],f[2],f[3],lower.tail=F)
    attributes(p) <- NULL
    return(p)
}
sapply(c('cp5','cp7','cp8','cp13','cp20','honcp22','honcp21a'),
       function(x) lmp(lm(all_pc1$norm ~ my_data[,x])))
```

All p-values are highly-significant.

```{r}
cp7 <- data.frame(q=my_data$cp7,w=all_pc1$norm)
cp7 <- na.omit(cp7)
my_lm <- lm(data=cp7, q ~ w)
summary(my_lm)
ggplot(cp7,aes(x=w,y=q)) +
  geom_jitter(color='tomato3',size=5,alpha=0.2) +
  theme_classic() +
  geom_smooth(method='lm',size=2,color='royalblue') +
  annotate('text',label=paste("R-sq:",round(summary(my_lm)$r.squared,2)),  
           size=10,x=-0.5,y=2.5,hjust=0,vjust=0,color='royalblue') +
  ylab('cp7') +
  xlab('Composite index') +
  theme(text=element_text(size=20)) 
```

```{r}
cp13 <- data.frame(q=my_data$cp13,w=all_pc1$norm)
cp13 <- na.omit(cp13)
my_lm <- lm(data=cp13, q ~ w)
summary(my_lm)
ggplot(cp13,aes(x=w,y=q)) +
  geom_jitter(color='tomato3',size=5,alpha=0.2) +
  theme_classic() +
  geom_smooth(method='lm',size=2,color='royalblue') +
  annotate('text',label=paste("R-sq:",round(summary(my_lm)$r.squared,2)),  
           size=10,x=-0.5,y=2.5,hjust=0,vjust=0,color='royalblue') +
  ylab('cp13') +
  xlab('Composite index') +
  theme(text=element_text(size=20)) 
```

```{r}
cp20 <- data.frame(q=my_data$cp20,w=all_pc1$norm)
cp20 <- na.omit(cp20)
my_lm <- lm(data=cp20, q ~ w)
summary(my_lm)
ggplot(cp20,aes(x=w,y=q)) +
  geom_jitter(color='tomato3',size=5,alpha=0.2) +
  theme_classic() +
  geom_smooth(method='lm',size=2,color='royalblue') +
  annotate('text',label=paste("R-sq:",round(summary(my_lm)$r.squared,2)),  
           size=10,x=-0.5,y=2.5,hjust=0,vjust=0,color='royalblue') +
  ylab('cp20') +
  xlab('Composite index') +
  theme(text=element_text(size=20)) 
```

Finally, check these values against the comments in `make_indices.R` to be sure we're calculating the same index:

```{r}
quantile(all_pc1$norm)
```