---
title: "Community Activity index"
author: "Jorge Perez"
output: html_document
---

The following is a proposed index, measuring the level of civic and political participation, and civil socity involvement of the individuals interrogated. Variables such as 'Attendance to religious meetings' and 'Sports Activities' were excluded, due to them not directly falling into the categories previously listed. In all, do individuals participate in making their community's social and political environment better? If so, how often?

##HONDURAS##

* `cp5`: Tried to solve a community problem (1=Once a week, 2=Once or twice a month, 3=Once or twice a year, 4=Never)
* `cp7`: Attended meetings of Parent Association (1=Once a week, 2=Once or twice a month, 3=Once or twice a year, 4=Never)
* `cp8`: Attended meetings of Community Improvement Association (1=Once a week, 2=Once or twice a month, 3=Once or twice a year, 4=Never)
* `cp13`: Attended meetings of a Political Party (1=Once a week, 2=Once or twice a month, 3=Once or twice a year, 4=Never)
* `cp20`: Attended meetings of Women's Group (1=Once a week, 2=Once or twice a month, 3=Once or twice a year, 4=Never)
* `honcp22`: Help for security groups (1=Once a week, 2=Once or twice a month, 3=Once or twice a year, 4=Never)
* `honcp21a`: Help for neighborhood associations (1=Once a week, 2=Once or twice a month, 3=Once or twice a year, 4=Never)

The two latter elements are particular to Honduras, and are not present for the data on El Salvador and Guatemala. They may eventually be removed for a composite index.

```{r, message=FALSE}
library(ggplot2)
library(mice)
library(plyr) # nice for re-formatting data
library(GGally) # for plot matrices
set.seed(12345) # makes imputation reproducible
```

```{r}
lapop.2014.HND <- read.csv("../HND-2014.csv")
my_data <- lapop.2014.HND[,c('cp5','cp7','cp8','cp13','cp20','honcp22','honcp21a')]
```

```{r}
is.na(my_data[my_data>16]) <- TRUE
nrow(na.omit(my_data))
```
Ignoring missing values, we are looking at 762 rows left

```{r}
sum(is.na(my_data)) / (nrow(my_data)*ncol(my_data))
```
7.5% of values missing!

```{r}
my_complete <- na.omit(my_data)
pr_complete <- prcomp(my_complete,center=TRUE,scale=FALSE)
plot(pr_complete)
summary(pr_complete)
```
Without any imputation, 32.7% of variance is concentrated on the first principal component.

```{r}
my_imp <- mice(my_data,printFlag = F) 
my_imp
```
None of the variables stand out as needing too much imputation, with the CLEAR exception of `cp20`, which was question conditioned on the subject being female. Presumably, what the imputation is telling us is that the individual in question is the kind of guy who would go to women's groups if men were invited.

```{r}
pr <- lapply(1:5,function(x) prcomp(complete(my_imp,x),scale=FALSE,center=TRUE))
plot(pr[[1]])
summary(pr[[1]])
```
With the imputation, we don't have much change. 32.5% of variance is concentrated on the first principal component.

```{r}
pc1 <- pr[[1]]$x[,1]
pc2 <- pr[[1]]$x[,2]
qplot(pc1,pc2)
```

```{r}
all_pc1 <- data.frame(llply(1:5, function(i) pr[[i]]$x[,1]))
names(all_pc1) <- c('imp1','imp2','imp3','imp4','imp5')
ggpairs(all_pc1) + theme_classic()
```

And yes! We have some great correlations. This may get imputed with rather high confidence, given that most people don't attend these meetings much or at all - this also is common among the most engaged, whose presence is not as frequent in these kind of meetings!

```{r}
all_pc1$avg <- rowMeans(all_pc1)
quantile(all_pc1$avg)
colMeans(my_data[all_pc1$avg <= quantile(all_pc1$avg)[2],],na.rm=TRUE)
colMeans(my_data[all_pc1$avg >= quantile(all_pc1$avg)[4],],na.rm=TRUE)
```
When comparing the lower 25% quartile to the higher 75% quartile, all of the variables show that the subjects are not often engaged in these kinds of activities; in some cases, they are never present.

* Some initiative in trying to solve a community problem `cp5`; however, judging by the number of people that are never present, I am assuming that these meetings have some exclusivity. Not surprising!
* The case above is also applicable to:
* `cp7`: Meetings to Parent Association
* `cp8`: Meetings to Community Improvement Association
* `honcp21a`: Help for neighborhood associations.

In the instance that there is a:

* Meetings of a Political Party (`b13`)
* Meeting of a Women's Group (`b20`)
* Help for Security Groups (`honcp22`)

The participation is slim to none.

Note that high values of the index correspond to larger numbers, which means *less* frequent engagement. Let's reverse the index so that everything is pointing in the right direction. 

It would be interesting to look more into the profile of the people that do engage in these activities - particularly, the men who imputed values of 1 or 2 in women's groups meetings.

```{r}
all_pc1$norm <- -scale(all_pc1$avg) 
predict_data <- data.frame(cp5=c(4,1),cp7=c(4,1),cp8=c(4,1),
                           cp13=c(4,1),cp20=c(4,1),honcp22=c(4,1),
                           honcp21a=c(4,1))
# The first row is the responses that would come from the least-engaged person
# possible; the second row is the responses from a maximally-engaged person.
minmax <- rowMeans(sapply(1:5, function(i) predict(pr[[i]],predict_data)[,1]))
minmax <- -(minmax - mean(all_pc1$avg)) / sd(all_pc1$avg) 
# Comparing minmax to quantile(all_pc1$norm), we can see that some people are
# pretty close to the minimum level of engagement, and no one is near the 
# maximum.

predict_data2 <- data.frame(4-diag(7))
names(predict_data2) <- names(my_data)
scores <- rowMeans(sapply(1:5, function(i) predict(pr[[i]],predict_data2)[,1]))
scores <- -(scores-mean(all_pc1$avg)) / sd(all_pc1$avg) 
diff <- data.frame(d=(scores - minmax[1]),n=names(predict_data))
ggplot(diff,aes(x=n,y=d)) +
  geom_bar(stat='identity',fill='skyblue') +
  coord_flip() +
  theme_classic() +
  theme(text=element_text(size=20),
        axis.title.y=element_blank(),
        axis.title.x=element_blank())
```

All variables have a strong influence on the index.

```{r}
lmp <- function (modelobject) {
    if (class(modelobject) != "lm") stop("Not an object of class 'lm' ")
    f <- summary(modelobject)$fstatistic
    p <- pf(f[1],f[2],f[3],lower.tail=F)
    attributes(p) <- NULL
    return(p)
}
sapply(c('cp5','cp7','cp8','cp13','cp20','honcp22','honcp21a'),
       function(x) lmp(lm(all_pc1$norm ~ my_data[,x])))
```

All p-values are highly-significant.

```{r}
cp7 <- data.frame(q=my_data$cp7,w=all_pc1$norm)
cp7 <- na.omit(cp7)
my_lm <- lm(data=cp7, q ~ w)
summary(my_lm)
ggplot(cp7,aes(x=w,y=q)) +
  geom_jitter(color='tomato3',size=5,alpha=0.2) +
  theme_classic() +
  geom_smooth(method='lm',size=2,color='royalblue') +
  annotate('text',label=paste("R-sq:",round(summary(my_lm)$r.squared,2)),  
           size=10,x=-0.5,y=2.5,hjust=0,vjust=0,color='royalblue') +
  ylab('cp7') +
  xlab('Composite index') +
  theme(text=element_text(size=20)) 
```

```{r}
cp13 <- data.frame(q=my_data$cp13,w=all_pc1$norm)
cp13 <- na.omit(cp13)
my_lm <- lm(data=cp13, q ~ w)
summary(my_lm)
ggplot(cp13,aes(x=w,y=q)) +
  geom_jitter(color='tomato3',size=5,alpha=0.2) +
  theme_classic() +
  geom_smooth(method='lm',size=2,color='royalblue') +
  annotate('text',label=paste("R-sq:",round(summary(my_lm)$r.squared,2)),  
           size=10,x=-0.5,y=2.5,hjust=0,vjust=0,color='royalblue') +
  ylab('cp13') +
  xlab('Composite index') +
  theme(text=element_text(size=20)) 
```

```{r}
cp20 <- data.frame(q=my_data$cp20,w=all_pc1$norm)
cp20 <- na.omit(cp20)
my_lm <- lm(data=cp20, q ~ w)
summary(my_lm)
ggplot(cp20,aes(x=w,y=q)) +
  geom_jitter(color='tomato3',size=5,alpha=0.2) +
  theme_classic() +
  geom_smooth(method='lm',size=2,color='royalblue') +
  annotate('text',label=paste("R-sq:",round(summary(my_lm)$r.squared,2)),  
           size=10,x=-0.5,y=2.5,hjust=0,vjust=0,color='royalblue') +
  ylab('cp20') +
  xlab('Composite index') +
  theme(text=element_text(size=20)) 
```

Before, moving with the other two countries, let us drop `honcp22`and `honcp21a`from the index, since they are exclusive to the studies done in Honduras. Without them, we still have an index for all three countries in the Northern Triangle.

##Guatemala##

* `cp5`: Tried to solve a community problem (1=Once a week, 2=Once or twice a month, 3=Once or twice a year, 4=Never)
* `cp7`: Attended meetings of Parent Association (1=Once a week, 2=Once or twice a month, 3=Once or twice a year, 4=Never)
* `cp8`: Attended meetings of Community Improvement Association (1=Once a week, 2=Once or twice a month, 3=Once or twice a year, 4=Never)
* `cp13`: Attended meetings of a Political Party (1=Once a week, 2=Once or twice a month, 3=Once or twice a year, 4=Never)
* `cp20`: Attended meetings of Women's Group (1=Once a week, 2=Once or twice a month, 3=Once or twice a year, 4=Never)

```{r}
library(ggplot2)
library(mice)
library(plyr) # nice for re-formatting data
library(GGally) # for plot matrices
set.seed(12345) # makes imputation reproducible
```

```{r}
lapop.2014.GTM <- read.csv("../GTM-2014.csv")
my_data <- lapop.2014.GTM[,c('cp5','cp7','cp8','cp13','cp20')]
```

```{r}
is.na(my_data[my_data>16]) <- TRUE
nrow(na.omit(my_data))
```
Ignoring missing data, we are left with 748 rows.

```{r}
sum(is.na(my_data)) / (nrow(my_data)*ncol(my_data))
```
About 10.1% of data missing!

```{r}
my_complete <- na.omit(my_data)
pr_complete <- prcomp(my_complete,center=TRUE,scale=FALSE)
plot(pr_complete)
summary(pr_complete)
```
Without any imputation, 43.75% of variance is concentrated on the first principal component.

```{r}
my_imp <- mice(my_data,printFlag = F) 
my_imp
```
Running the imputations, `cp20`is the column with the LARGEST amount of missing cells. Again, our comments made for this variable, on the Honduras index, are applicable here. Furthermore, it seems to be drawing up a pattern for the region.

```{r}
pr <- lapply(1:5,function(x) prcomp(complete(my_imp,x),scale=FALSE,center=TRUE))
plot(pr[[1]])
summary(pr[[1]])
```
With imputations, we have quite a decrease: 41% of variance is concentrated on the first principal component.

```{r}
pc1 <- pr[[1]]$x[,1]
pc2 <- pr[[1]]$x[,2]
qplot(pc1,pc2)
```

For my untrained eyes, I am seeing four separate clusters here, but unlike all the other ones that I have seen, this one seems to be a bit more "cohesive".

```{r}
all_pc1 <- data.frame(llply(1:5, function(i) pr[[i]]$x[,1]))
names(all_pc1) <- c('imp1','imp2','imp3','imp4','imp5')
ggpairs(all_pc1) + theme_classic()
```

Although not perfect, the correlations seem quite strong.

```{r}
all_pc1$avg <- rowMeans(all_pc1)
quantile(all_pc1$avg)
colMeans(my_data[all_pc1$avg < quantile(all_pc1$avg)[2],],na.rm=TRUE)
colMeans(my_data[all_pc1$avg > quantile(all_pc1$avg)[4],],na.rm=TRUE)
```
OK! 
A few things I am going to highlight:

* Where Guatemala is concerned, our subjects for the most part attend political party meetings (`cp13`) on a VERY RARE to non-existant basis.
* The male subjects rarely attend or are not at all present in Women's Groups meetings, though again, it would be interesting to look into the kind of guy that does go to these meetings.
* Ditto for trying to solve community problems (`cp5`)
* Finally, attendance seems to be more regular to Parents' Associations (`cp7`)and Community Improvement Meetings (`cp8`). I am personally not aware of what these meetings comprise or of the profile of the people that do go to them, though.

```{r}
all_pc1$norm <- -scale(all_pc1$avg) 
predict_data <- data.frame(cp5=c(4,1),cp7=c(4,1),cp8=c(4,1),
                           cp13=c(4,1),cp20=c(4,1))
# The first row is the responses that would come from the least-engaged person
# possible; the second row is the responses from a maximally-engaged person.
minmax <- rowMeans(sapply(1:5, function(i) predict(pr[[i]],predict_data)[,1]))
minmax <- -(minmax - mean(all_pc1$avg)) / sd(all_pc1$avg) 
# Comparing minmax to quantile(all_pc1$norm), we can see that some people are
# pretty close to the minimum level of engagement, and no one is near the 
# maximum.

predict_data2 <- data.frame(4-diag(5))
names(predict_data2) <- names(my_data)
scores <- rowMeans(sapply(1:5, function(i) predict(pr[[i]],predict_data2)[,1]))
scores <- -(scores-mean(all_pc1$avg)) / sd(all_pc1$avg) 
diff <- data.frame(d=(scores - minmax[1]),n=names(predict_data))
ggplot(diff,aes(x=n,y=d)) +
  geom_bar(stat='identity',fill='skyblue') +
  coord_flip() +
  theme_classic() +
  theme(text=element_text(size=20),
        axis.title.y=element_blank(),
        axis.title.x=element_blank())
```

Strong influence on score from all variables, but the index is largely dominated by `cp7`and `cp8`.

```{r}
lmp <- function (modelobject) {
    if (class(modelobject) != "lm") stop("Not an object of class 'lm' ")
    f <- summary(modelobject)$fstatistic
    p <- pf(f[1],f[2],f[3],lower.tail=F)
    attributes(p) <- NULL
    return(p)
}
sapply(c('cp5','cp7','cp8','cp13','cp20'),
       function(x) lmp(lm(all_pc1$norm ~ my_data[,x])))
```
All of the p-values are small and deserve to be considered, but there's something new here: `cp8`shows a value of 0. Not close to 0, but literally the number zero!

```{r}
cp13 <- data.frame(q=my_data$cp13,w=all_pc1$norm)
cp13 <- na.omit(cp13)
my_lm <- lm(data=cp13, q ~ w)
summary(my_lm)
ggplot(cp13,aes(x=w,y=q)) +
  geom_jitter(color='tomato3',size=5,alpha=0.2) +
  theme_classic() +
  geom_smooth(method='lm',size=2,color='royalblue') +
  annotate('text',label=paste("R-sq:",round(summary(my_lm)$r.squared,2)),  
           size=10,x=-0.5,y=2.5,hjust=0,vjust=0,color='royalblue') +
  ylab('cp13') +
  xlab('Composite index') +
  theme(text=element_text(size=20)) 
```

##El Salvador##

* `cp5`: Tried to solve a community problem (1=Once a week, 2=Once or twice a month, 3=Once or twice a year, 4=Never)
* `cp7`: Attended meetings of Parent Association (1=Once a week, 2=Once or twice a month, 3=Once or twice a year, 4=Never)
* `cp8`: Attended meetings of Community Improvement Association (1=Once a week, 2=Once or twice a month, 3=Once or twice a year, 4=Never)
* `cp13`: Attended meetings of a Political Party (1=Once a week, 2=Once or twice a month, 3=Once or twice a year, 4=Never)
* `cp20`: Attended meetings of Women's Group (1=Once a week, 2=Once or twice a month, 3=Once or twice a year, 4=Never)

```{r}
library(ggplot2)
library(mice)
library(plyr) # nice for re-formatting data
library(GGally) # for plot matrices
set.seed(12345) # makes imputation reproducible
```

```{r}
lapop.2014.SLV <- read.csv("../SLV-2014.csv")
my_data <- lapop.2014.SLV[,c('cp5','cp7','cp8','cp13','cp20')]
```

```{r}
is.na(my_data[my_data>16]) <- TRUE
nrow(na.omit(my_data))
```
With missing values, we are left with 835 rows.

```{r}
sum(is.na(my_data)) / (nrow(my_data)*ncol(my_data))
```
Approximately 9% of data missing.

```{r}
my_complete <- na.omit(my_data)
pr_complete <- prcomp(my_complete,center=TRUE,scale=FALSE)
plot(pr_complete)
summary(pr_complete)
```
Without any imputation, 35.5% of variance concentrated on first principal component.

```{r}
my_imp <- mice(my_data,printFlag = F) 
my_imp
```
After running the imputations, we clearly see that much like Guatemala and El Salvador, column `cp20`is largely missing data. However, if the responses to this question are like all the other indeces, then the imputation should be done with high confidence.

```{r}
pr <- lapply(1:5,function(x) prcomp(complete(my_imp,x),scale=FALSE,center=TRUE))
plot(pr[[1]])
summary(pr[[1]])
```
With imputation, there is a slight change, with 34.2% of variance concentrated on the first principal component.

```{r}
pc1 <- pr[[1]]$x[,1]
pc2 <- pr[[1]]$x[,2]
qplot(pc1,pc2)
```

What I'm seeing here is 4 separate clusters forming, with a much more solid "foundation" on the higher values.
```{r}
all_pc1 <- data.frame(llply(1:5, function(i) pr[[i]]$x[,1]))
names(all_pc1) <- c('imp1','imp2','imp3','imp4','imp5')
ggpairs(all_pc1) + theme_classic()
```
And despite the large amount of missing cells in column `cp20`, the correlations are still pretty strong.

```{r}
all_pc1$avg <- rowMeans(all_pc1)
quantile(all_pc1$avg)
colMeans(my_data[all_pc1$avg <= quantile(all_pc1$avg)[2],],na.rm=TRUE)
colMeans(my_data[all_pc1$avg >= quantile(all_pc1$avg)[4],],na.rm=TRUE)
```
So, the NaN's that we got to see in the "Extorsion Index" are popping up here again, which makes me uncomfortable to continue with this index. Judging from the values that I have gotten on the lower 25% quartile, then much like Honduras and Guatemala, the participation of individuals in these kinds of activities, committees or gatherings goes from rarely to not at all. Interesting to see that `cp5`(Tried to solve a community problem) has a lower value (~2.47) relative to all the other variables, which would allow me to say that there is more participation in this area. Then again, with the "NaN"s, I do not feel confident enough to assert thiswith authority.

```{r}
all_pc1$norm <- -scale(all_pc1$avg) 
predict_data <- data.frame(cp5=c(4,1),cp7=c(4,1),cp8=c(4,1),
                           cp13=c(4,1),cp20=c(4,1))
# The first row is the responses that would come from the least-engaged person
# possible; the second row is the responses from a maximally-engaged person.
minmax <- rowMeans(sapply(1:5, function(i) predict(pr[[i]],predict_data)[,1]))
minmax <- -(minmax - mean(all_pc1$avg)) / sd(all_pc1$avg) 
# Comparing minmax to quantile(all_pc1$norm), we can see that some people are
# pretty close to the minimum level of engagement, and no one is near the 
# maximum.

predict_data2 <- data.frame(4-diag(5))
names(predict_data2) <- names(my_data)
scores <- rowMeans(sapply(1:5, function(i) predict(pr[[i]],predict_data2)[,1]))
scores <- -(scores-mean(all_pc1$avg)) / sd(all_pc1$avg) 
diff <- data.frame(d=(scores - minmax[1]),n=names(predict_data))
ggplot(diff,aes(x=n,y=d)) +
  geom_bar(stat='identity',fill='skyblue') +
  coord_flip() +
  theme_classic() +
  theme(text=element_text(size=20),
        axis.title.y=element_blank(),
        axis.title.x=element_blank())
```

All variables seem to have a strong influence on the index score, but `cp5` seems to be on the lead here, followed by `cp8`and `cp7`, in that order.
```{r}
lmp <- function (modelobject) {
    if (class(modelobject) != "lm") stop("Not an object of class 'lm' ")
    f <- summary(modelobject)$fstatistic
    p <- pf(f[1],f[2],f[3],lower.tail=F)
    attributes(p) <- NULL
    return(p)
}
sapply(c('cp5','cp7','cp8','cp13','cp20'),
       function(x) lmp(lm(all_pc1$norm ~ my_data[,x])))
```
Yes! The values here are very small, but notice how EPICALLY tiny the p-value for `cp5`is!
