---
title: "Community Activity index"
author: "Craig Jolley"
output: html_document
---

Extending the community participation index over the entire region. 

* `cp5`: Tried to solve a community problem (1=Once a week, 2=Once or twice a month, 3=Once or twice a year, 4=Never)
* `cp7`: Attended meetings of Parent Association (1=Once a week, 2=Once or twice a month, 3=Once or twice a year, 4=Never)
* `cp8`: Attended meetings of Community Improvement Association (1=Once a week, 2=Once or twice a month, 3=Once or twice a year, 4=Never)
* `cp13`: Attended meetings of a Political Party (1=Once a week, 2=Once or twice a month, 3=Once or twice a year, 4=Never)
* `cp20`: Attended meetings of Women's Group (1=Once a week, 2=Once or twice a month, 3=Once or twice a year, 4=Never)

```{r, message=FALSE}
library(ggplot2)
library(mice)
library(plyr) 
library(GGally) 
```

```{r}
setwd("C:/Users/Craig/Desktop/SMA/VSFS/LAPOP-SMA/indices")
lapop.2014.GTM <- read.csv('../../2014-GTM.csv')
lapop.2014.SLV <- read.csv('../../2014-SLV.csv')
lapop.2014.HND <- read.csv('../../2014-HND.csv')
v <- c('cp5','cp7','cp8','cp13','cp20')
my_data <- rbind(lapop.2014.GTM[,v],lapop.2014.SLV[,v],lapop.2014.HND[,v])
is.na(my_data[my_data>16]) <- TRUE
my_data <- my_data[,order(names(my_data))] # alphabetize columns
```

We'll use multiple imputation to fill in missing data, but let's omit missing values first to be sure that everything makes sense.

```{r}
my_complete <- na.omit(my_data)
pr_complete <- prcomp(my_complete,center=TRUE,scale=FALSE)
plot(pr_complete)
summary(pr_complete)
```
Without any imputation, 38.4% of variance is concentrated on the first principal component.

```{r}
my_imp <- mice(my_data,printFlag = F,seed=12345) 
nmis <- my_imp$nmis
nm <- data.frame(nmis=nmis/nrow(my_data),var=names(nmis))
ggplot(nm,aes(x=var,y=nmis)) +
  geom_bar(stat = 'identity', fill = '#90AFC5') +
  ylab('Fraction of imputed values') +
  coord_flip() +
  theme_classic() +
  theme(legend.position='none',
        axis.ticks=element_blank(),
        axis.line=element_blank(),
        axis.title.y=element_blank(),
        text=element_text(size=20))
```

None of the variables stand out as needing too much imputation, with the CLEAR exception of `cp20`, which was question conditioned on the subject being female. Presumably, what the imputation is telling us is that the individual in question is the kind of guy who would go to women's groups if men were invited.

```{r}
pr <- lapply(1:5,function(x) prcomp(complete(my_imp,x),scale=FALSE,center=TRUE))
plot(pr[[1]])
summary(pr[[1]])
```

With the imputation, we don't have much change; 37% of variance is concentrated on the first principal component.

```{r}
pc1 <- pr[[1]]$x[,1]
pc2 <- pr[[1]]$x[,2]
qplot(pc1,pc2) + theme_classic()
```

Variation along PC1 seems to be fairly continuous, indicating that this index isn't dominated by a single variable.

```{r}
all_pc1 <- data.frame(llply(1:5, function(i) pr[[i]]$x[,1]))
names(all_pc1) <- c('imp1','imp2','imp3','imp4','imp5')
ggpairs(all_pc1) + theme_classic()
```

Our different imputed datasets correlate strongly, and the distribution skews toward low values of PC1.

```{r, fig.height=5, fig.width=10}
all_pc1$avg <- rowMeans(all_pc1)
q25 <- quantile(all_pc1$avg)[2]
q75 <- quantile(all_pc1$avg)[4]
qchange <- data.frame(low=colMeans(my_data[all_pc1$avg <= q25,],na.rm=TRUE),
                      hi=colMeans(my_data[all_pc1$avg >= q75,],na.rm=TRUE))
qchange$slope <- qchange$hi - qchange$low
qchange$label <- rownames(qchange)
qchange$even <- rank(qchange$hi)/nrow(qchange)*max(qchange$hi)
ggplot(data=qchange) +
  geom_segment(aes(x=0,y=qchange$low,xend=1,yend=qchange$hi,
                   color=qchange$slope),size=1) +
  scale_color_gradient(low='#FF420E',high='#89DA59') +
  geom_text(aes(x=1,y=qchange$hi,label=qchange$label,
                hjust=0,color=qchange$slope),size=5) +
  annotate("text",x=0,y=0,label='Average in lowest quartile',hjust=0) +
  annotate("text",x=1,y=0,label='Average in highest quartile',hjust=1) +
  scale_x_continuous(limits=c(0,1.5)) +
  theme_classic() +
  theme(legend.position='none',
        axis.ticks=element_blank(),
        axis.line=element_blank(),
        axis.text.x=element_blank(),
        axis.title=element_blank(),
        text=element_text(size=20))
```

All five variables decrease as the index increases; recall that larger values correspond to less-frequent participation, so our index is pointing in the right direction. Because `cp7` (parent's association attendance) has the steepest slope, we suspect that it will have the strongest influence on the composite index.

```{r}
all_pc1$norm <- scale(all_pc1$avg) 
predict_data <- data.frame(cp5=c(4,1),cp7=c(4,1),cp8=c(4,1),
                           cp13=c(4,1),cp20=c(4,1))
# The first row is the responses that would come from the least-engaged person
# possible; the second row is the responses from a maximally-engaged person.
minmax <- rowMeans(sapply(1:5, function(i) predict(pr[[i]],predict_data)[,1]))
minmax <- (minmax - mean(all_pc1$avg)) / sd(all_pc1$avg) 
# Comparing minmax to quantile(all_pc1$norm), we can see that some people are
# pretty close to the minimum level of engagement, and no one is near the 
# maximum.

predict_data2 <- data.frame(4-diag(5))
names(predict_data2) <- names(my_data)
scores <- rowMeans(sapply(1:5, function(i) predict(pr[[i]],predict_data2)[,1]))
scores <- (scores-mean(all_pc1$avg)) / sd(all_pc1$avg) 
diff <- data.frame(d=(scores - minmax[1]),n=names(predict_data))
ggplot(diff,aes(x=n,y=d)) +
  geom_bar(stat='identity',fill='skyblue') +
  coord_flip() +
  theme_classic() +
  theme(text=element_text(size=20),
        axis.title.y=element_blank(),
        axis.title.x=element_blank())
```

All variables influence the index; the weakest influence comes from `cp13` (Attended meetings of a political party), apparently because even the most-engaged quartile hardly did this at all.

Since all variables take on multiple values, we will check p-values using linear regression.

```{r}
lmp <- function (modelobject) {
    if (class(modelobject) != "lm") stop("Not an object of class 'lm' ")
    f <- summary(modelobject)$fstatistic
    p <- pf(f[1],f[2],f[3],lower.tail=F)
    attributes(p) <- NULL
    return(p)
}
sapply(c('cp5','cp7','cp8','cp13','cp20'),
       function(x) lmp(lm(all_pc1$norm ~ my_data[,x])))
```

All p-values are highly-significant.

```{r}
cp7 <- data.frame(q=my_data$cp7,w=all_pc1$norm)
cp7 <- na.omit(cp7)
my_lm <- lm(data=cp7, q ~ w)
summary(my_lm)
ggplot(cp7,aes(x=w,y=q)) +
  geom_jitter(color='tomato3',size=5,alpha=0.2) +
  theme_classic() +
  geom_smooth(method='lm',size=2,color='royalblue') +
  annotate('text',label=paste("R-sq:",round(summary(my_lm)$r.squared,2)),  
           size=10,x=-0.5,y=2.5,hjust=0,vjust=0,color='royalblue') +
  ylab('cp7') +
  xlab('Composite index') +
  theme(text=element_text(size=20)) 
```

```{r}
cp13 <- data.frame(q=my_data$cp13,w=all_pc1$norm)
cp13 <- na.omit(cp13)
my_lm <- lm(data=cp13, q ~ w)
summary(my_lm)
ggplot(cp13,aes(x=w,y=q)) +
  geom_jitter(color='tomato3',size=5,alpha=0.2) +
  theme_classic() +
  geom_smooth(method='lm',size=2,color='royalblue') +
  annotate('text',label=paste("R-sq:",round(summary(my_lm)$r.squared,2)),  
           size=10,x=-0.5,y=2.5,hjust=0,vjust=0,color='royalblue') +
  ylab('cp13') +
  xlab('Composite index') +
  theme(text=element_text(size=20)) 
```

```{r}
cp20 <- data.frame(q=my_data$cp20,w=all_pc1$norm)
cp20 <- na.omit(cp20)
my_lm <- lm(data=cp20, q ~ w)
summary(my_lm)
ggplot(cp20,aes(x=w,y=q)) +
  geom_jitter(color='tomato3',size=5,alpha=0.2) +
  theme_classic() +
  geom_smooth(method='lm',size=2,color='royalblue') +
  annotate('text',label=paste("R-sq:",round(summary(my_lm)$r.squared,2)),  
           size=10,x=-0.5,y=2.5,hjust=0,vjust=0,color='royalblue') +
  ylab('cp20') +
  xlab('Composite index') +
  theme(text=element_text(size=20)) 
```

Finally, check these values against the comments in `make_indices.R` to be sure we're calculating the same index:

```{r}
quantile(all_pc1$norm)
```

