---
title: "Fear index"
author: "Craig Jolley"
date: "March 16, 2016"
output: html_document
---


```{r}
setwd("C:/Users/Craig/Desktop/SMA/VSFS/LAPOP-SMA/indices")
lapop.2014.GTM <- read.csv("../../GTM-2014.csv")
```

The following indicators related to fear and violence were measured in Guatemala:

- vic40: Out of fear of crime, has limited the places to go shopping (0=no, 1=yes) 
- vic41: Out of fear of crime, has limited the places to go for recreation (0=no, 1=yes) 
- vic43: Out of fear of crime, has felt the need to change neighborhoods (0=no, 1=yes) 
- vic45: Out of fear of crime, has changed jobs (0=no, 1=yes) 
- fear10: Avoided Walking Through Dangerous Areas  (0=no, 1=yes) 
- fear6e: Insecurity on public transportation (1 = a lot, 4 = not at all) 
- fear6f: Insecurity at Schools (1 = a lot, 4 = not at all) 
- vic1ext: Victim of Crime (1=yes, 2=no) 
- vic1exta: Victim of Crime (Frequency) (999999 if vic1ext=2, otherwise max=15) 
- vic1hogar: Other Victim of Crime in Household (1=yes, 2=no)
- aoj11: Perception of Neighborhood Security (1 = safe, 4 = unsafe) 
- pese1: Perception of Neighborhood Violence (1=high, 3=low) 
- pese2: Trend in Neighborhood Violence (1=high, 3=low) 
- aoj17: Gang Presence in Neighborhood (1 = a lot, 4 = none) 
- diso7: Youth Loitering a Problem in the Neighborhood (1 = very serious, 5 = not a problem) 
- diso8: Youth in Gangs a Problem in the Neighborhood (1 = very serious, 5 = not a problem) 
- diso10: Illegal Drug Trafficking a Problem in the Neighborhood (1 = very serious, 5 = not a problem) 
- diso18: Gang Fights a Problem in the Neighborhood (1 = very serious, 5 = not a problem) 
- diso14: Drug Addicts a Problem in the Neighborhood (1 = very serious, 5 = not a problem) 
- diso16: Assaults a Problem in the Neighborhood (1 = very serious, 5 = not a problem) 
- diso17: Shootings a Problem in the Neighborhood (1 = very serious, 5 = not a problem) 
- vicbar1: Burglaries in the Neighborhood (1=yes, 2=no) 
- vicbar1f: Number of Occurrences of Burglaries in Neighborhood (1 = weekly, 3=yearly) 
- vicbar3: Sales of Illegal Drugs in Neighborhood (1=yes, 2=no) 
- vicbar4: Extortion or Blackmail in the Neighborhood (1=yes, 2=no)
- vicbar7: Murders in the Neighborhood (1=yes, 2=no) 

Note that `vic44` appears to be on topic, but turns out not to correlate strongly with this index; we won't include it.

```{r,message=FALSE}
library(ggplot2)
library(mice)
library(plyr)
library(GGally)
set.seed(12345)
my_data <- lapop.2014.GTM[,c('vic40', 'vic41','vic43',
                             'fear10','fear6e','fear6f','vic1ext','vic1exta',
                             'vic1hogar','aoj11','pese1','pese2','aoj17',
                             'diso7','diso8','diso10','diso18','diso14',
                             'diso16','diso17','vicbar1','vicbar1f',
                             'vicbar3','vicbar4','vicbar7')]
my_data[my_data$vic1exta > 800000,'vic1exta'] <- 0
my_data$a4 <- as.numeric(lapop.2014.GTM$a4 %in% c(5,14,27,30,57))
is.na(my_data[my_data>800000]) <- TRUE

my_complete <- na.omit(my_data)
pr_complete <- prcomp(my_complete, center = TRUE, scale = FALSE)
plot(pr_complete)

summary(pr_complete)
```

40% of total variance winds up in the first component.

```{r}
my_imp <- mice(my_data, printFlag = F)
my_imp
```

The only variable that got a lot of imputation was `vicbar1f` -- number of burglaries in the neighborhood. I can imagine a lot of people don't think they know how many burglaries happen in their neighborhood.

Moving onto the PCA:
```{r}
pr <- lapply(1:5,function(x) prcomp(complete(my_imp,x),scale=FALSE,center=TRUE))
plot(pr[[1]])
summary(pr[[1]])
```

Imputation actually increased the fraction of variance in the first component to 44%.

```{r}
pc1 <- pr[[1]]$x[,1]
pc2 <- pr[[1]]$x[,2]
qplot(pc1,pc2) + theme_classic()
```

Fairly continuous distribution along PC1 means that no single variable dominates.

```{r}
all_pc1 <- data.frame(llply(1:5, function(i) pr[[i]]$x[,1]))
names(all_pc1) <- c('imp1', 'imp2', 'imp3', 'imp4', 'imp5')
ggpairs(all_pc1) + theme_classic()
```

Correlations for different imputed datasets are about 0.997 – imputation is consistent and distribution of PC1 looks well-behaved.

```{r,fig.height=8,fig.width=10}
all_pc1$avg <- rowMeans(all_pc1)
qchange <- data.frame(low=colMeans(my_data[all_pc1$avg < quantile(all_pc1$avg)[2],],na.rm=TRUE),
                      hi=colMeans(my_data[all_pc1$avg > quantile(all_pc1$avg)[4],],na.rm=TRUE))
qchange$slope <- qchange$hi - qchange$low
qchange$label <- rownames(qchange)
qchange$even <- rank(qchange$hi)/nrow(qchange)*max(qchange$hi)
ggplot(data=qchange) +
  geom_segment(aes(x=0,y=qchange$low,xend=1,yend=qchange$hi,
                   color=qchange$slope),size=1) +
  scale_color_gradient(low='#FF420E',high='#89DA59') +
  geom_text(aes(x=1,y=(qchange$even+qchange$hi)/2,label=qchange$label,
                hjust=0,color=qchange$slope),size=5) +
  annotate("text",x=0,y=0,label='Average in lowest quartile',hjust=0) +
  annotate("text",x=1,y=0,label='Average in highest quartile',hjust=1) +
  scale_x_continuous(limits=c(0,1.5)) +
  theme_classic() +
  theme(legend.position='none',
        axis.ticks=element_blank(),
        axis.line=element_blank(),
        axis.text.x=element_blank(),
        axis.title=element_blank(),
        text=element_text(size=20))
```

The `diso` variables all increase dramatically, while `aoj11` shows a dramatic decrease. This suggests that high values of the index indicate not being particularly concerned about crime and violence; we'll have to reverse the scale.

```{r}
all_pc1$norm <- scale(-all_pc1$avg) # sign flip -- now high values mean more fear
predict_data <- data.frame(vic40 = c(0,1), vic41 = c(0,1), vic43 = c(0,1), 
                           fear10 = c(0,1), fear6e=c(4,1),
                           fear6f = c(4,1), vic1ext = c(2,1),
                           vic1exta=c(0,max(my_data$vic1exta)),
                           vic1hogar=c(2,1),aoj11=c(1,4),pese1=c(3,1),
                           pese2=c(3,1),aoj17=c(4,1),diso7=c(5,1),diso8=c(5,1),
                           diso10=c(5,1),diso18=c(5,1),diso14=c(5,1),
                           diso16=c(5,1),diso17=c(5,1),vicbar1=c(2,1),
                           vicbar1f=c(3,1),vicbar3=c(2,1),vicbar4=c(2,1),
                           vicbar7=c(2,1),a4=c(0,1))

minmax <- rowMeans(sapply(1:5, function(i) predict(pr[[i]],predict_data)[,1]))
minmax <- -(minmax - mean(all_pc1$avg)) / sd(all_pc1$avg)

fear <- data.frame(w=all_pc1$norm) 

predict_data2 <- data.frame(diag(26))
names(predict_data2) <- names(predict_data)
# right now, you have something that would feed 0's and 1's into your score
# prediction -- this doesn't really make sense because many of your questions
# are on a scale from 1-2 or 1-4. Here's one (slightly clunky) way to fix it:
predict_data2$fear6f <- 4 - 3*predict_data2$fear6f 
predict_data2$fear6e <- 4 - 3*predict_data2$fear6e 
predict_data2$vic1ext <- 2 - predict_data2$vic1ext
predict_data2$vic1hogar <- 2 - predict_data2$vic1hogar
predict_data2$aoj11 <- 1 + 3*predict_data2$aoj11
predict_data2$aoj17 <- 4 - 3*predict_data2$aoj17
predict_data2$pese1 <- 3 - 2*predict_data2$pese1
predict_data2$pese2 <- 3 - 2*predict_data2$pese2
predict_data2$diso7 <- 5 - 4*predict_data2$diso7
predict_data2$diso8 <- 5 - 4*predict_data2$diso8
predict_data2$diso10 <- 5 - 4*predict_data2$diso10
predict_data2$diso18 <- 5 - 4*predict_data2$diso18
predict_data2$diso14 <- 5 - 4*predict_data2$diso14
predict_data2$diso16 <- 5 - 4*predict_data2$diso16
predict_data2$diso17 <- 5 - 4*predict_data2$diso17
predict_data2$vicbar1 <- 2 - predict_data2$vicbar1
predict_data2$vicbar1f <- 3 - 2*predict_data2$vicbar1f
predict_data2$vicbar3 <- 2 - predict_data2$vicbar3
predict_data2$vicbar4 <- 2 - predict_data2$vicbar4
predict_data2$vicbar7 <- 2 - predict_data2$vicbar7
```

Now, in each row of predict_data2, all values are those that will give a lower score, except for one.
```{r}
scores <- rowMeans(sapply(1:5, function(i) predict(pr[[i]],predict_data2)[,1]))
scores <- (mean(all_pc1$avg)-scores)/sd(all_pc1$avg)
```

All of the scores are reasonably close to the minimum value stored in minmax.

```{r}
diff <- data.frame(d=(scores-minmax[1]), n = names(predict_data))
ggplot(diff, aes(x=n,y=d)) + 
  geom_bar(stat = 'identity', fill = 'skyblue') + 
  coord_flip() + 
  theme_classic() + 
  theme(text = element_text(size=20), 
        axis.title.y = element_blank(),
        axis.title.x = element_blank())
```

The `diso` variables show the largest effect.

```{r}
fear10 <- data.frame(q=my_data$fear10,w=fear$w)
my_lm <- lm(data=fear10, q ~ w)
summary(my_lm)
ggplot(fear10,aes(x=w,y=q)) +
  geom_jitter(color='tomato3',size=5,alpha=0.2) +
  theme_classic() +
  geom_smooth(method='lm',size=2,color='royalblue') +
  annotate('text',label=paste("R-sq:",round(summary(my_lm)$r.squared,2)),  
  size=10,x=-1.5,y=1,hjust=0,vjust=0,color='royalblue') +
  ylab('fear10') +
  xlab('Fear index') +
  theme(text=element_text(size=20))
```

```{r}
fear6f <- data.frame(q=my_data$fear6f,w=fear$w)
my_lm <- lm(data=fear6f, q ~ w)
summary(my_lm)

ggplot(fear6f,aes(x=w,y=q)) +
  geom_jitter(color='tomato3',size=5,alpha=0.2) +
  theme_classic() +
  geom_smooth(method='lm',size=2,color='royalblue') +
  annotate('text',label=paste("R-sq:",round(summary(my_lm)$r.squared,2)),  
  size=10,x=0.5,y=4,hjust=0,vjust=0,color='royalblue') +
  ylab('fear10') +
  xlab('Fear index') +
  theme(text=element_text(size=20))
```

```{r}
diso8 <- data.frame(q=my_data$diso8,w=fear$w)
my_lm <- lm(data=diso8, q ~ w)
summary(my_lm)

ggplot(diso8,aes(x=w,y=q)) +
  geom_jitter(color='tomato3',size=5,alpha=0.2) +
  theme_classic() +
  geom_smooth(method='lm',size=2,color='royalblue') +
  annotate('text',label=paste("R-sq:",round(summary(my_lm)$r.squared,2)),  
  size=10,x=0.5,y=4,hjust=0,vjust=0,color='royalblue') +
  ylab('diso8') +
  xlab('Fear index') +
  theme(text=element_text(size=20))
```

We'll use linear regression to get p-values for variables that can have more than 2 values, and a 2-sammple t-test for the binary variables.

```{r}
lmp <- function (modelobject) {
  if (class(modelobject) != "lm") stop("Not an object of class 'lm' ")
    f <- summary(modelobject)$fstatistic
    p <- pf(f[1],f[2],f[3],lower.tail=F)
    attributes(p) <- NULL
    return(p)
}
lm_data <- my_data[,c('fear6f','fear6e','vic1exta','aoj11','aoj17','pese1','pese2',
                      'diso7','diso8','diso10','diso14','diso16','diso17',
                      'diso18','vicbar3')]
sapply(lm_data, function(x) lmp(lm(x ~ all_pc1$norm)))

bin_cor <- function(var) {
  tt <- t.test(all_pc1$norm[var==min(var,na.rm=TRUE)],
               all_pc1$norm[var==max(var,na.rm=TRUE)])
  tt$p.value
}

bin_vars <- c('vic40','vic41','vic43','vic1ext','vic1hogar',
                       'vicbar1','vicbar3','vicbar4','vicbar7','a4')
sapply(bin_vars,function(x) bin_cor(my_data[,x]))
```

All are highly significant.

