---
title: "Fear index"
author: "Craig Jolley"
date: "March 16, 2016"
output: html_document
---

```{r}
setwd("C:/Users/Craig/Desktop/SMA/VSFS/LAPOP-SMA/indices")
lapop.2014.SLV <- read.csv("../../SLV-2014.csv")
```

- fear10: Has avoided walking through dangerous areas
- vic44: Organized in Neighborhood for Security (0 = no, 1=yes) 
- vic1exta: Victim of Crime (Frequency) (999999 if vic1ext=2, otherwise max=15) 
- vic1hogar: Other Victim of Crime in Household (1=yes, 2=no)
- aoj11: Perception of Neighborhood Security (1 = safe, 4 = unsafe) 
- pese1: Perception of Neighborhood Violence (1=high, 3=low) 
- pese2: Trend in Neighborhood Violence (1=high, 3=low) 
- aoj17: Gang Presence in Neighborhood (1 = a lot, 4 = none) 
- diso7: Youth Loitering a Problem in the Neighborhood (1 = very serious, 5 = not a problem) 
- diso8: Youth in Gangs a Problem in the Neighborhood (1 = very serious, 5 = not a problem) 
- diso10: Illegal Drug Trafficking a Problem in the Neighborhood (1 = very serious, 5 = not a problem) 
- diso18: Gang Fights a Problem in the Neighborhood (1 = very serious, 5 = not a problem) 
- diso14: Drug Addicts a Problem in the Neighborhood (1 = very serious, 5 = not a problem) 
- diso16: Assaults a Problem in the Neighborhood (1 = very serious, 5 = not a problem) 
- diso17: Shootings a Problem in the Neighborhood (1 = very serious, 5 = not a problem) 
- elsdiso18: Gang recruitment at school exits a problem in the neighborhood (1 = very serious, 5 = not a problem)
- elsdiso19: School dropouts because of gangs a serious problem (1 = very serious, 5 = not a problem)
- vicbar1: Burglaries in the Neighborhood (1=yes, 2=no) 
- vicbar1f: Number of Occurrences of Burglaries in Neighborhood (1 = weekly, 3=yearly) - vicbar3: Sales of Illegal Drugs in Neighborhood (1=yes, 2=no) 
- vicbar4: Extortion or Blackmail in the Neighborhood (1=yes, 2=no)
- vicbar7: Murders in the Neighborhood (1=yes, 2=no) 
- vicbar7f: Number of occurrences of murders
- a4 == 5,14,27,30,57: Biggest problem related to crime/violence/security

`vic44` appears to be on-topic, but actually doesn't correlate as strongly as the other variables do, so we're leaving it out.

```{r,message=FALSE}
library(ggplot2)
library(mice)
library(plyr)
library(GGally)
set.seed(12345)
my_data <- lapop.2014.SLV[,c('fear10','vic1ext','vic1exta',
                             'vic1hogar','aoj11','pese1','pese2','aoj17',
                             'diso7','diso8','diso10','diso18','diso14',
                             'diso16','diso17','vicbar1','vicbar1f',
                             'vicbar3','vicbar4','vicbar7','vicbar7f',
                             'elsdiso18','elsdiso19')]
my_data$vic1exta[my_data$vic1exta > 800000] <- 0
my_data$a4 <- as.numeric(lapop.2014.SLV$a4 %in% c(5,14,27,30,57))
is.na(my_data[my_data>800000]) <- TRUE

my_complete <- na.omit(my_data)
pr_complete <- prcomp(my_complete, center = TRUE, scale = FALSE)
plot(pr_complete)
summary(pr_complete)
```

38% of total variance winds up in the first component.

```{r}
my_imp <- mice(my_data, printFlag = F)
my_imp
```

The only variable that needed a lot of imputation was `vicbar7f` -- the number of murders in the neighborhood. Let's look at the PCA for one of our imputed datasets.

```{r}
pr <- lapply(1:5,function(x) prcomp(complete(my_imp,x),scale=FALSE,center=TRUE))
plot(pr[[1]])
summary(pr[[1]])
```

Imputation decreased the fraction of variance in the first component to 48%.

```{r}
pc1 <- pr[[1]]$x[,1]
pc2 <- pr[[1]]$x[,2]
qplot(pc1,pc2) + theme_classic()
```

The distribution along PC1 looks pretty continuous, while the PC2 distribution looks very skewed. Fortunately, we're only using PC1.

```{r}
all_pc1 <- data.frame(llply(1:5, function(i) pr[[i]]$x[,1]))
names(all_pc1) <- c('imp1', 'imp2', 'imp3', 'imp4', 'imp5')
ggpairs(all_pc1) + theme_classic()
all_pc1$avg <- rowMeans(all_pc1)
```

Correlations are high, and the distribution is skewed toward high values.

```{r,fig.height=8,fig.width=10}
qchange <- data.frame(low=colMeans(my_data[all_pc1$avg < quantile(all_pc1$avg)[2],],
                                   na.rm=TRUE),
                      hi=colMeans(my_data[all_pc1$avg > quantile(all_pc1$avg)[4],],
                                  na.rm=TRUE))
qchange$slope <- qchange$hi - qchange$low
qchange$label <- rownames(qchange)
qchange$even <- rank(qchange$hi)/nrow(qchange)*max(qchange$hi)
ggplot(data=qchange) +
  geom_segment(aes(x=0,y=qchange$low,xend=1,yend=qchange$hi,
                   color=qchange$slope),size=1) +
  scale_color_gradient(low='#FF420E',high='#89DA59') +
  geom_text(aes(x=1,y=(qchange$even+qchange$hi)/2,label=qchange$label,
                hjust=0,color=qchange$slope),size=5) +
  annotate("text",x=0,y=0,label='Average in lowest quartile',hjust=0) +
  annotate("text",x=1,y=0,label='Average in highest quartile',hjust=1) +
  scale_x_continuous(limits=c(0,1.5)) +
  theme_classic() +
  theme(legend.position='none',
        axis.ticks=element_blank(),
        axis.line=element_blank(),
        axis.text.x=element_blank(),
        axis.title=element_blank(),
        text=element_text(size=20))
```

The `diso` variables all increase dramatically, while `aoj11` shows a dramatic decrease. This suggests that high values of the index indicate not being particularly concerned about crime and violence; we'll have to reverse the scale.

```{r}
all_pc1$norm <- scale(-all_pc1$avg) # sign flip -- now high values mean more fear
predict_data <- data.frame(fear10 = c(0,1), vic1ext = c(2,1),
                           vic1exta=c(0,max(my_data$vic1exta)),
                           vic1hogar=c(2,1),aoj11=c(1,4),pese1=c(3,1),
                           pese2=c(3,1),aoj17=c(4,1),diso7=c(5,1),diso8=c(5,1),
                           diso10=c(5,1),diso18=c(5,1),diso14=c(5,1),
                           diso16=c(5,1),diso17=c(5,1),vicbar1=c(2,1),
                           vicbar1f=c(3,1),vicbar3=c(2,1),vicbar4=c(2,1),
                           vicbar7=c(2,1),vicbar7f=c(0,3),elsdiso18=c(5,1),
                           elsdiso19=c(5,1),a4=c(0,1))

minmax <- rowMeans(sapply(1:5, function(i) predict(pr[[i]],predict_data)[,1]))
minmax <- (mean(all_pc1$avg) - minmax) / sd(all_pc1$avg)
# quantile(all_pc1$norm)
fear <- data.frame(w=all_pc1$norm) 
```

```{r}
predict_data2 <- data.frame(diag(24))
names(predict_data2) <- names(predict_data)
# right now, you have something that would feed 0's and 1's into your score
# prediction -- this doesn't really make sense because many of your questions
# are on a scale from 1-2 or 1-4. Here's one (slightly clunky) way to fix it:
predict_data2$fear10 <- 4 - 3*predict_data2$fear10 
predict_data2$vic1ext <- 2 - predict_data2$vic1ext
predict_data2$vic1exta <- 20*predict_data2$vic1exta
predict_data2$vic1hogar <- 2 - predict_data2$vic1hogar
predict_data2$aoj11 <- 1 + 3*predict_data2$aoj11
predict_data2$aoj17 <- 4 - 3*predict_data2$aoj17
predict_data2$pese1 <- 3 - 2*predict_data2$pese1
predict_data2$pese2 <- 3 - 2*predict_data2$pese2
predict_data2$diso7 <- 5 - 4*predict_data2$diso7
predict_data2$diso8 <- 5 - 4*predict_data2$diso8
predict_data2$diso10 <- 5 - 4*predict_data2$diso10
predict_data2$diso18 <- 5 - 4*predict_data2$diso18
predict_data2$diso14 <- 5 - 4*predict_data2$diso14
predict_data2$diso16 <- 5 - 4*predict_data2$diso16
predict_data2$diso17 <- 5 - 4*predict_data2$diso17
predict_data2$elsdiso18 <- 5 - 4*predict_data2$elsdiso18
predict_data2$elsdiso19 <- 5 - 4*predict_data2$elsdiso19
predict_data2$vicbar1 <- 2 - predict_data2$vicbar1
predict_data2$vicbar1f <- 3 - 2*predict_data2$vicbar1f
predict_data2$vicbar3 <- 2 - predict_data2$vicbar3
predict_data2$vicbar4 <- 2 - predict_data2$vicbar4
predict_data2$vicbar7 <- 2 - predict_data2$vicbar7
predict_data2$vicbar7f <- 3*predict_data2$vicbar7f
```

Now, in each row of predict_data2, all values are those that will give a lower score, except for one.

```{r}
scores <- rowMeans(sapply(1:5, function(i) predict(pr[[i]],predict_data2)[,1]))
scores <- (mean(all_pc1$avg)-scores)/sd(all_pc1$avg)
```

Now you can see that all of the scores are reasonably close to the minimum value stored in minmax.
```{r}
diff <- data.frame(d=(scores-minmax[1]), n = names(predict_data))
ggplot(diff, aes(x=n,y=d)) + 
  geom_bar(stat = 'identity', fill = 'skyblue') + 
  coord_flip() + 
  theme_classic() + 
  theme(text = element_text(size=20), 
        axis.title.y = element_blank(),
        axis.title.x = element_blank())
```

All variables seem to have an effect, with a somewhat smaller one for `vicbar7f` (number of occurrences of murders) and `fear10` (avoided walking through dangerous areas). While murders in the neighborhood seem like something that would contribute to fear, most people answered that they aren't very common and this variable was frequently imputed because people didn't answer it at all.

```{r}
fear10 <- data.frame(q=my_data$fear10,w=fear$w)
my_lm <- lm(data=fear10, q ~ w)
ggplot(fear10,aes(x=w,y=q)) +
  geom_jitter(color='tomato3',size=5,alpha=0.2) +
  theme_classic() +
  geom_smooth(method='lm',size=2,color='royalblue') +
  annotate('text',label=paste("R-sq:",round(summary(my_lm)$r.squared,2)),  
  size=10,x=-1.5,y=1,hjust=0,vjust=0,color='royalblue') +
  ylab('fear10') +
  xlab('Fear index') +
  theme(text=element_text(size=20))
```

```{r}
elsdiso19 <- data.frame(q=my_data$elsdiso19,w=fear$w)
my_lm <- lm(data=elsdiso19, q ~ w)
ggplot(elsdiso19,aes(x=w,y=q)) +
  geom_jitter(color='tomato3',size=5,alpha=0.2) +
  theme_classic() +
  geom_smooth(method='lm',size=2,color='royalblue') +
  annotate('text',label=paste("R-sq:",round(summary(my_lm)$r.squared,2)),  
  size=10,x=0.5,y=4,hjust=0,vjust=0,color='royalblue') +
  ylab('elsdiso19') +
  xlab('Fear index') +
  theme(text=element_text(size=20))
```

We can make this all more rigorous by examining p-values:

```{r}
lmp <- function (modelobject) {
  if (class(modelobject) != "lm") stop("Not an object of class 'lm' ")
    f <- summary(modelobject)$fstatistic
    p <- pf(f[1],f[2],f[3],lower.tail=F)
    attributes(p) <- NULL
    return(p)
}
lm_data <- my_data[,c('vic1exta','aoj11','aoj17','pese1','pese2',
                      'diso7','diso8','diso10','diso14','diso16','diso17',
                      'diso18','elsdiso18','elsdiso19','vicbar7f','vicbar1f',
                      'vicbar3')]
sapply(lm_data, function(x) lmp(lm(x ~ all_pc1$norm)))

bin_cor <- function(var) {
  tt <- t.test(all_pc1$norm[var==min(var,na.rm=TRUE)],
               all_pc1$norm[var==max(var,na.rm=TRUE)])
  tt$p.value
}

bin_vars <- c('fear10','vic1ext','vic1hogar','vicbar1','vicbar3',
              'vicbar4','vicbar7')
sapply(bin_vars,function(x) bin_cor(my_data[,x]))
```

Finally, check these values against the comments in `make_indices.R` to be sure we're calculating the same index:

```{r}
quantile(all_pc1$norm)
```
