---
title: "Trust in Government Index"
author: "Jorge L. Perez"
date: "31 October 2015"
output: html_document
---

The following is a proposal for a Trust in Government Index, for which the following variables were considered to be the most relevant in determining the citizens' trust in all civic, public, political and governmental structures:

* `b1`: Courts Guarantee a Fair Trial (1=Not at all, 7=A lot)
* `b2`: Respect for Political Institutions (1=Not at all, 7=A lot)
* `b3`: Respect for Basic Rights (1=Not at all, 7=A lot)
* `b4`: Pride in Political System (1=Not at all, 7=A lot)
* `b6`: People Should Support the Political System (1=Not at all, 7=A lot)
* `b10a`: Trust in Judicial System (1=Not at all, 7=A lot)
* `b11`: Trust in the National Electoral Institution (1=Not at all, 7=A lot)
* `b12`: Trust in Armed Forces (1=Not at all, 7=A lot)
* `b13`: Trust in National Legislature (1=Not at all, 7=A lot)
* `b18`: Trust in National Police (1=Not at all, 7=A lot)
* `b21`: Trust in Political Parties (1=Not at all, 7=A lot)
* `b21a`: Trust in Executive (1=Not at all, 7=A lot)
* `b32`: Trust in Local Government (1=Not at all, 7=A lot) 
* `b37`: Trust in the Media (1=Not at all, 7=A lot)
* `b47a`: Trust in Elections (1=Not at all, 7=A lot)
* `b14`: Trust in the Central Government (1=Not at all, 7=A lot)
* `b15`: Trust in the Public Ministry (1=Not at all, 7=A lot)
* `b19`: Trust in Superior Court of Accounts (1=Not at all, 7=A lot)
* `b46`: Trust in National Anticorruption Council (1=Not at all, 7=A lot) 
* `honb51`: Trust in Commission for Human Rights (1=Not at all, 7=A lot)
* `venb11`: Trust in the secret vote in Honduras (1=Not at all, 7=A lot)
* `venhonb51`: Trust that election results reflect the votes issued (1=Not at all, 7=A lot)
* `venhonvb10`: Level of agreement for having international observers at elections (1=Not at all, 7=A lot)
* `epp1`: Political parties represent voters well (1=Not at all, 7=A lot)
* `epp3`: Political parties listen to voters (1=Not at all, 7=A lot)
* `n9`: Evaluation of administration's handling of corruption (1=Not at all, 7=A lot)
* `n11`: Evaluation of administration's handling of citizen security (1=Not at all, 7=A lot) 
* `n15`: Evaluation of administration's handling of economy (1=Not at all, 7=A lot)
* `b3milx`: Armed Forces respect human rights (1=Not at all, 7=A lot) 

Variables measuring the levels of trust in religious institutions, such as the Catholic and the Protestant Churches, were left out of this proposed index, since they do not partake in the structures of government and political and civil affairs - regardless, of their influence in the citizens' daily lives. 

```{r}
library(ggplot2)
library(mice)
library(plyr) # nice for re-formatting data
library(GGally) # for plot matrices
set.seed(12345) # makes imputation reproducible

setwd("C:/Users/Craig/Desktop/SMA/VSFS/LAPOP-SMA/indices")
lapop.2014.HND <- read.csv("../../HND-2014.csv")
my_data <- lapop.2014.HND[,c('b1','b2','b3','b4','b6','b10a','b11','b12','b13',
                             'b18','b21','b21a','b32', 'b37','b47a','b14',
                             'b15','b19','b46','honb51','venb11', 'venhonb51', 
                             'venhonvb10','epp1','epp3','n9','n11','n15',
                             'b3milx')]
```

Firstly, how many missing values do we have here?

```{r}
is.na(my_data[my_data>16]) <- TRUE
nrow(na.omit(my_data))
```
So, ignoring missing values, we have 1,120 rows left

```{r}
sum(is.na(my_data)) / (nrow(my_data)*ncol(my_data))
```

Around 2.3% of missing variables.

```{r}
my_complete <- na.omit(my_data)
pr_complete <- prcomp(my_complete,center=TRUE,scale=FALSE)
plot(pr_complete)
summary(pr_complete)
```

If we don't do any imputation, 39.9% of variance is contained in the first principal component; the others all appear to be much lower.

```{r}
my_imp <- mice(my_data,printFlag = F) 
my_imp
```

Some variables stand out as needing more imputation that others:

* `b1`: Some people chose not to talk about their opinion on whether or not Courts deliver a fair trial.
* `b19` (Trust in the Superior Court of Accounts) is a little higher than others, possibly because some people don't know enough about it to know whether to trust it or not.
* `b46` (Trust in National Anticorruption Council) either reflects the lack of awareness of this institution or the perception of corruption, in general. 
* `honb51`(Trust in Commission on Human Rights) again, this may also be a reflection of the lack of awareness of a special instution such as this one, or the overall perception of how human rights are being treated in Honduras.
* I am asserting this guess for the latter two, since there have been calls to implement a body like the CICIG in Honduras (i.e. CICIH).

```{r}
pr <- lapply(1:5,function(x) prcomp(complete(my_imp,x),scale=FALSE,center=TRUE))
plot(pr[[1]])
summary(pr[[1]])
```

After imputation, things don't change much -- 38.6% of variance is in the first PC.

```{r}
pc1 <- pr[[1]]$x[,1]
pc2 <- pr[[1]]$x[,2]
qplot(pc1,pc2) + theme_classic()
```

The distribution along PC1 seems smooth, indicating that this isn't dominated by a single variable.

```{r}
all_pc1 <- data.frame(llply(1:5, function(i) pr[[i]]$x[,1]))
names(all_pc1) <- c('imp1','imp2','imp3','imp4','imp5')
ggpairs(all_pc1) + theme_classic()
```

Correlations between imputations are high; distributions are well-behaved.

```{r,fig.height=5,fig.width=10}
all_pc1$avg <- rowMeans(all_pc1)
q25 <- quantile(all_pc1$avg)[2]
q75 <- quantile(all_pc1$avg)[4]
qchange <- data.frame(low=colMeans(my_data[all_pc1$avg <= q25,],na.rm=TRUE),
                      hi=colMeans(my_data[all_pc1$avg >= q75,],na.rm=TRUE))
qchange$slope <- qchange$hi - qchange$low
qchange$label <- rownames(qchange)
qchange$even <- rank(qchange$hi)/nrow(qchange)*max(c(qchange$low,qchange$hi))
ggplot(data=qchange) +
  geom_segment(aes(x=0,y=qchange$low,xend=1,yend=qchange$hi,
                   color=qchange$slope),size=1) +
  scale_color_gradient(low='#FF420E',high='#89DA59') +
  geom_text(aes(x=1,y=qchange$even,label=qchange$label,
                hjust=0,color=qchange$slope),size=5) +
  annotate("text",x=0,y=0,label='Average in lowest quartile',hjust=0) +
  annotate("text",x=1,y=0,label='Average in highest quartile',hjust=1) +
  scale_x_continuous(limits=c(0,1.5)) +
  theme_classic() +
  theme(legend.position='none',
        axis.ticks=element_blank(),
        axis.line=element_blank(),
        axis.text.x=element_blank(),
        axis.title=element_blank(),
        text=element_text(size=20))
```

All variables tend to decrease with PC1; recall that higher values indicate more trust in government institutions. We'll need to reverse the direction of the index.

```{r}
all_pc1$norm <- scale(-all_pc1$avg)
predict_data <- data.frame(b1=c(1,7),b2=c(1,7),b3=c(1,7),b4=c(1,7),
                           b6=c(1,7),b10a=c(1,7),b11=c(1,7),b12=c(1,7),
                           b13=c(1,7),b18=c(1,7),b21=c(1,7),b21a=c(1,7),
                           b32=c(1,7),b37=c(1,7),b47a=c(1,7),b14=c(1,7),
                           b15=c(1,7), b19=c(1,7), b46=c(1,7), honb51=c(1,7), 
                           venb11=c(1,7), venhonb51=c(1,7), venhonvb10=c(1,7), 
                           epp1=c(1,7), epp3=c(1,7), n9=c(1,7), n11=c(1,7), 
                           n15=c(1,7), b3milx=c(1,7))
minmax <- rowMeans(sapply(1:5, function(i) predict(pr[[i]],predict_data)[,1]))
minmax <- -(minmax-mean(all_pc1$avg)) / sd(all_pc1$avg)
# at this point, make sure to compare values of minmax to 
# quantile(all_pc1$norm)
# to make sure that you got all of your minus signs right! 

predict_data2 <- data.frame(diag(29)+1)
names(predict_data2) <- names(my_data)
scores <- rowMeans(sapply(1:5, function(i) predict(pr[[i]],predict_data2)[,1]))
scores <- -(scores-mean(all_pc1$avg)) / sd(all_pc1$avg)
diff <- data.frame(d=(scores - minmax[1]),n=names(predict_data))
ggplot(diff,aes(x=n,y=d)) +
  geom_bar(stat='identity',fill='skyblue') +
  coord_flip() +
  theme_classic() +
  theme(text=element_text(size=20),
        axis.title.y=element_blank(),
        axis.title.x=element_blank())
```

All variables seem to have a roughly equal on the Trust Index. Note that the average influence on the index for a 1-point change in any individual variable is only about 0.03 -- considering that the minimum and maximum values of the index differ by about 5, this is a sign that large changes in several variables make the difference between a high and a low score.

We can be more rigorous here by calculating the p-values. Because all variables take on multiple values, we can use linear regression.

```{r}
lmp <- function (modelobject) {
    if (class(modelobject) != "lm") stop("Not an object of class 'lm' ")
    f <- summary(modelobject)$fstatistic
    p <- pf(f[1],f[2],f[3],lower.tail=F)
    attributes(p) <- NULL
    return(p)
}
sapply(c('b1','b2','b3', 'b4','b6','b10a','b11','b12','b13',
                             'b18','b21','b21a','b32', 'b37','b47a','b14','b15','b19','b46', 'honb51', 'venb11', 'venhonvb10', 'epp1', 'epp3', 'n9', 'n11', 'n15', 'b3milx'),
       function(x) lmp(lm(all_pc1$norm ~ my_data[,x])))

```

All variables correlate strongly with the composite index.

```{r}
b37 <- data.frame(q=my_data$b37,w=all_pc1$norm)
b37 <- na.omit(b37)
my_lm <- lm(data=b37, q ~ w)
summary(my_lm)
ggplot(b37,aes(x=w,y=q)) +
  geom_jitter(color='tomato3',size=5,alpha=0.2) +
  theme_classic() +
  geom_smooth(method='lm',size=2,color='royalblue') +
  annotate('text',label=paste("R-sq:",round(summary(my_lm)$r.squared,2)),  
           size=10,x=-0.5,y=2.5,hjust=0,vjust=0,color='royalblue') +
  ylab('b37') +
  xlab('Composite index') +
  theme(text=element_text(size=20))
```

