---
title: "Trust in Government Index"
author: "Jorge L. Perez"
date: "31 October 2015"
output: html_document
---

The following is a proposal for a Trust in Government Index, for which the following variables measured in El Salvador were considered to be the most relevant in determining the citizens' trust in all civic, public, political and governmental structures:

* `b1`: Courts Guarantee Fair Trial (1=Not at all, 7=A lot)
* `b2`: Respect for Political Institutions (1=Not at all, 7=A lot)
* `b3`: Respect for Basic Rights (1=Not at all, 7=A lot)
* `b4`: Pride in Political System (1=Not at all, 7=A lot)
* `b6`: People Should Support the Political System (1=Not at all, 7=A lot)
* `b10a`: Trust in Judicial System (1=Not at all, 7=A lot)
* `b11`: Trust in the National Electoral Institution (1=Not at all, 7=A lot)
* `b12`: Trust in Armed Forces (1=Not at all, 7=A lot)
* `b13`: Trust in the National Legislature (1=Not at all, 7=A lot)
* `b18`: Trust in National Police (1=Not at all, 7=A lot)
* `b21`: Trust in Political Parties (1=Not at all, 7=A lot)
* `b21a`: Trust in Executive (1=Not at all, 7=A lot)
* `b32`: Trust in Local Government (1=Not at all, 7=A lot)
* `b47a`: Trust in Elections (1=Not at all, 7=A lot)
* `esb48`: Trust in the Institute of Access to Public Information (1=Not at all, 7=A lot)
* `n9`: Evaluation of Administration's Handling of Corruption (1=Not at all, 7=A lot)
* `n11`: Evaluation of Administration's Handling of Citizen Security (1=Not at all, 7=A lot)
* `n15`: Evaluation of Administration's Handling of Economy (1=Not at all, 7=A lot)
* `epp1`: Political Parties Represent Voters Well (1=Not at all, 7=A lot)
* `epp3`: Political Parties Listen to Voters (1=Not at all, 7=A lot)
* `b3milx`: Armed Forces Respect Human Rights (1=Not at all, 7=A lot)
* `pr4`: Government Respects Private Property (1=Not at all, 7=A lot)
* `epn3a`: Central Governement Represents Your Interests (1=A lot, 4=Not at all)
* `epn3b`: Legislative Representative Represents Your Interests (1=A lot, 4=Not at all)
* `epn3c`: Mayor and City Council Represent Your Interests (1=A lot, 4=Not at all)

As is to be expected with a LAPOP study, there are questions that were adapted to reflect the national reality of a country. Here we have examples of a few variables that we did not see for Honduras, but that were selected for this proposed Trust Index for El Salvador:

* `esb48` - which presents an institution particular to El Salvador's civil and political structure
* `epn3a`, `epn3b`, `epn3c` - which reflect confidence on these institutions, and their democratic nature
* `pr4`- reflecting confidence in the government's role in society


```{r, message=FALSE}
library(ggplot2)
library(mice)
library(plyr) 
library(GGally) 

setwd("C:/Users/Craig/Desktop/SMA/VSFS/LAPOP-SMA/indices")
lapop.2014.SLV <- read.csv('../../2014-SLV.csv')

my_data <- lapop.2014.SLV[,c('b1','b2','b3','b4','b6','b10a','b11','b12','b13',
                             'b18','b21','b21a','b32', 'b47a','esb48','n9',
                             'n11','n15','epp1','epp3','b3milx','pr4','epn3a',
                             'epn3b','epn3c','aoj18')]

# aoj18 has a funny scale that we'll have to fix; 
# 1 = police protect against crime
# 2 = police participate in crime
# 3 = neither or both
my_data[lapop.2014.SLV$aoj18==2,'aoj18'] <- 3
my_data[lapop.2014.SLV$aoj18==3,'aoj18'] <- 2

is.na(my_data[my_data>16]) <- TRUE
my_data <- my_data[,order(names(my_data))] # alphabetize columns
nrow(na.omit(my_data))
```

With missing values, we are left with 1,198 rows.

```{r}
sum(is.na(my_data)) / (nrow(my_data)*ncol(my_data))
```

Around 1.2% of missing values.

```{r}
my_complete <- na.omit(my_data)
pr_complete <- prcomp(my_complete,center=TRUE,scale=TRUE)
plot(pr_complete)
summary(pr_complete)
```

Without any imputation, about 36% of my variance is concentrated in my first principal component.

```{r}
my_imp <- mice(my_data,printFlag = F,seed=12345) 
nmis <- my_imp$nmis
nm <- data.frame(nmis=nmis/nrow(my_data),var=names(nmis))
ggplot(nm,aes(x=var,y=nmis)) +
  geom_bar(stat = 'identity', fill = '#90AFC5') +
  ylab('Fraction of imputed values') +
  coord_flip() +
  theme_classic() +
  theme(legend.position='none',
        axis.ticks=element_blank(),
        axis.line=element_blank(),
        axis.title.y=element_blank(),
        text=element_text(size=20))
```

None of the variables seems to stand out as needing much imputation, with the exception of `aoj18`. The question of whether the police protect from crime or cause it was a hard one for some to answer.

```{r}
pr <- lapply(1:5,function(x) prcomp(complete(my_imp,x),scale=TRUE,center=TRUE))
plot(pr[[1]])
summary(pr[[1]])
```

Following the imputation, not much has changed. 36.75% of variance is concentrated on the first principal component.

```{r}
pc1 <- pr[[1]]$x[,1]
pc2 <- pr[[1]]$x[,2]
qplot(pc1,pc2) + theme_classic()
```

The distribution along PC1 seems smooth, indicating that this isn't dominated by a single variable.

```{r}
all_pc1 <- data.frame(llply(1:5, function(i) pr[[i]]$x[,1]))
names(all_pc1) <- c('imp1','imp2','imp3','imp4','imp5')
ggpairs(all_pc1) + theme_classic()
```

Correlations between imputations are high; distributions are well-behaved.

```{r,fig.height=5,fig.width=10}
all_pc1$avg <- rowMeans(all_pc1)
q25 <- quantile(all_pc1$avg)[2]
q75 <- quantile(all_pc1$avg)[4]
qchange <- data.frame(low=colMeans(my_data[all_pc1$avg <= q25,],na.rm=TRUE),
                      hi=colMeans(my_data[all_pc1$avg >= q75,],na.rm=TRUE))
qchange$slope <- qchange$hi - qchange$low
qchange$label <- rownames(qchange)
qchange$even <- rank(qchange$hi)/nrow(qchange)*max(c(qchange$low,qchange$hi))
ggplot(data=qchange) +
  geom_segment(aes(x=0,y=qchange$low,xend=1,yend=qchange$hi,
                   color=qchange$slope),size=1) +
  scale_color_gradient(low='#FF420E',high='#89DA59') +
  geom_text(aes(x=1,y=qchange$even,label=qchange$label,
                hjust=0,color=qchange$slope),size=5) +
  annotate("text",x=0,y=0,label='Average in lowest quartile',hjust=0) +
  annotate("text",x=1,y=0,label='Average in highest quartile',hjust=1) +
  scale_x_continuous(limits=c(0,1.5)) +
  theme_classic() +
  theme(legend.position='none',
        axis.ticks=element_blank(),
        axis.line=element_blank(),
        axis.text.x=element_blank(),
        axis.title=element_blank(),
        text=element_text(size=20))
```

All variables increase with PC1, except for `epn3a` through `epn3c`. For most variables, higher values indicate a higher level of trust in government, while the opposite is true for the `epn` variables and `aoj18`. Our index is pointing in the right direction.

```{r}
all_pc1$norm <- scale(all_pc1$avg) 
predict_data <- data.frame(b1=c(1,7),b2=c(1,7),b3=c(1,7),b4=c(1,7),
                           b6=c(1,7),b10a=c(1,7),b11=c(1,7),b12=c(1,7),
                           b13=c(1,7),b18=c(1,7),b21=c(1,7),b21a=c(1,7),
                           b32=c(1,7),b47a=c(1,7),esb48=c(1,7),
                           n9=c(1,7), n11=c(1,7), n15=c(1,7),epp1=c(1,7), 
                           epp3=c(1,7),b3milx=c(1,7), pr4=c(1,7), epn3a=c(4,1), 
                           epn3b=c(4,1),epn3c=c(4,1),aoj18=c(3,1))
minmax <- rowMeans(sapply(1:5, function(i) predict(pr[[i]],predict_data)[,1]))
minmax <- (minmax-mean(all_pc1$avg)) / sd(all_pc1$avg) 
# compares well with quantile(all_pc1$norm)

predict_data2 <- data.frame(diag(26)+1)
names(predict_data2) <- names(my_data)
predict_data2$epn3a <- 5 - predict_data2$epn3a
predict_data2$epn3b <- 5 - predict_data2$epn3b
predict_data2$epn3c <- 5 - predict_data2$epn3c
scores <- rowMeans(sapply(1:5, function(i) predict(pr[[i]],predict_data2)[,1]))
scores <- (scores-mean(all_pc1$avg)) / sd(all_pc1$avg) 
# sanity check: are all scores just a little above minmax[1]?
diff <- data.frame(d=(scores - minmax[1]),n=names(predict_data))
ggplot(diff,aes(x=n,y=d)) +
  geom_bar(stat='identity',fill='skyblue') +
  coord_flip() +
  theme_classic() +
  theme(text=element_text(size=20),
        axis.title.y=element_blank(),
        axis.title.x=element_blank())
```

Most variables influence the score to a similar degree.

Now onto a calculation of p-values, we see that-

```{r}
lmp <- function (modelobject) {
    if (class(modelobject) != "lm") stop("Not an object of class 'lm' ")
    f <- summary(modelobject)$fstatistic
    p <- pf(f[1],f[2],f[3],lower.tail=F)
    attributes(p) <- NULL
    return(p)
}
sapply(c('aoj18','b1','b2','b3', 'b4','b6','b10a','b11','b12','b13',
                             'b18','b21','b21a','b32', 'b47a','esb48','n9','n11','n15', 'epp1', 'epp3', 'b3milx', 'pr4', 'epn3a', 'epn3b','epn3c'),
       function(x) lmp(lm(all_pc1$norm ~ my_data[,x])))
```

all of the variables correlate strongly with the proposed index.

```{r}
epn3c <- data.frame(q=my_data$epn3c,w=all_pc1$norm)
epn3c <- na.omit(epn3c)
my_lm <- lm(data=epn3c, q ~ w)
summary(my_lm)
ggplot(epn3c,aes(x=w,y=q)) +
  geom_jitter(color='tomato3',size=5,alpha=0.2) +
  theme_classic() +
  geom_smooth(method='lm',size=2,color='royalblue') +
  annotate('text',label=paste("R-sq:",round(summary(my_lm)$r.squared,2)),  
           size=10,x=-0.5,y=2.5,hjust=0,vjust=0,color='royalblue') +
  ylab('epn3c') +
  xlab('Composite index') +
  theme(text=element_text(size=20))
```


```{r}
epn3a <- data.frame(q=my_data$epn3a,w=all_pc1$norm)
epn3a <- na.omit(epn3a)
my_lm <- lm(data=epn3a, q ~ w)
summary(my_lm)
ggplot(epn3a,aes(x=w,y=q)) +
  geom_jitter(color='tomato3',size=5,alpha=0.2) +
  theme_classic() +
  geom_smooth(method='lm',size=2,color='royalblue') +
  annotate('text',label=paste("R-sq:",round(summary(my_lm)$r.squared,2)),  
           size=10,x=-0.5,y=2.5,hjust=0,vjust=0,color='royalblue') +
  ylab('epn3a') +
  xlab('Composite index') +
  theme(text=element_text(size=20))
```

Finally, check these values against the comments in `make_indices.R` to be sure we're calculating the same index:

```{r}
quantile(all_pc1$norm)
```