---
title: "Wealth time series"
author: "Craig Jolley"
date: "January 27, 2016"
output: html_document
---

I'm curious which of the variables in our trust in government index have historical information available, and how these vary over time.

```{r}
setwd("C:/Users/Craig/Dropbox/SMA/VSFS/LAPOP-SMA")
source('make_trends.R')
source('make_indices.R')
crit_common_ts <- intersect(names(lapop.trends),crit_common) # all of them!
crit_ts_idx <- make_idx(lapop.trends,crit_common_ts,sgn=1)
```

The new index uses ```length(crit_common_ts)``` variables, rather than the previous ```length(crit_common)```, and is constructed using ```nrow(lapop.trends)``` observations rather than the previous ```nrow(lapop.2014.all)```. For the 2014 data (where both indices were calculated), how well do they agree?

```{r}
df1 <- data.frame(idnum=lapop.trends$idnum,crit_ts=crit_ts_idx)
df2 <- data.frame(idnum=lapop.2014.all$idnum,crit=crit_all)
j <- join(df2,df1,by='idnum')
my_lm <- lm(data=j,crit_ts~crit)
ggplot(data=j,aes(x=crit,y=crit_ts)) +
  geom_point(color='tomato',alpha=0.3) +
  theme_classic() +
  geom_smooth(method='lm',size=2,color='royalblue') +
  annotate('text',label=paste("R-sq:",round(summary(my_lm)$r.squared,3)),
             size=10,x=-1.5,y=1,hjust=0,vjust=0,color='royalblue') +
  xlab('2014 sympathy for critics') +
  ylab('2004-2014 sympathy for critics')
```

Almost perfectly.

```{r}
df3 <- data.frame(year=lapop.trends$year,pais=lapop.trends$pais,crit_idx=crit_ts_idx)
plotme <- ddply(df3,.(year,pais),summarize,mean=mean(crit_idx),
      lo=t.test(crit_idx)$conf.int[1],hi=t.test(crit_idx)$conf.int[2])
pais_text <- c('','Guatemala','El Salvador','Honduras')
plotme$pais <- pais_text[plotme$pais]
ggplot(plotme,aes(x=year,y=mean,group=pais,color=pais)) +
  geom_point(size=5) +
  geom_line(size=3) +
  geom_errorbar(aes(ymin=lo,ymax=hi,width=0.5)) +
  ylab('Average critic-sympathy index') +
  scale_x_continuous(limits=c(2004,2014)) +
  theme_classic()
```

The overall trend is downward, particularly in Guatemala.