Adding to the list of indices, the following is a proposed Extorsion index. The variables chosen represent questions that relate to the request for a bribe from any civil or governmental institution. (Beginning on page 51 of codebook PDF.)

##Guatemala##

* `exc2`: Police Officer asked for a bribe (0=No, 1=Yes)
* `exc6`: Government Employee asked for a bribe (0=No, 1=Yes)
* `exc20`: Soldier requested a bribe (0=No, 1=Yes)
* `exc11`: Bribe requested to process document in Municipal Government (0=No, 1=Yes)
* `exc13`: Asked to pay a bribe at work (0=No, 1=Yes)
* `exc14`: Asked to pay a bribe at the courts (0=No, 1=Yes)
* `exc15`: Asked to pay a bribe to use health services (0=No, 1=Yes)
* `exc16`: Asked to pay a bribe at school (0=No, 1=Yes)
* `exc7`: Perception of public corruption (1=common, 4=uncommon)

```{r, message=FALSE}
library(ggplot2)
library(mice)
library(plyr) # nice for re-formatting data
library(GGally) # for plot matrices
set.seed(12345) # makes imputation reproducible
```

```{r}
lapop.2014.GTM <- read.csv("../GTM-2014.csv")
my_data <- lapop.2014.GTM[,c('exc2','exc6','exc20','exc11','exc13','exc14',
                             'exc15','exc16','exc7')]
```

Since many of these questions are asking people to report on rare events, we should start by looking to see whether there are any that co-occur.

```{r}
test_data <- lapop.2014.GTM[,c('exc2','exc6','exc20','exc11','exc13','exc14',
                               'exc15','exc16')]
test_data[test_data>16] <- 0 # assume that non-responses mean no bribe paid
colMeans(test_data)
jitter_data <- data.frame(police=jitter(test_data$police,factor=2),
                          gov=jitter(test_data$gov,factor=2),
                          soldier=jitter(test_data$soldier,factor=2),
                          muni=jitter(test_data$muni,factor=2),
                          work=jitter(test_data$work,factor=2),
                          courts=jitter(test_data$courts,factor=2),
                          health=jitter(test_data$health,factor=2),
                          school=jitter(test_data$school,factor=2))
ggpairs(jitter_data) + theme_classic()
```

It's really disturbing that 14% of people report having paid bribes to the police. The plot is a little confusing, but the main thing to take away from it is that, for each possible pair of bribes, there are at least a few people who paid both. This gives me some hope that this could be the basis for a composite index. 

Because there are so many NA responses, let's try assuming that NA means 'no' for these 8 indices, and only impute exc7.

```{r}
is.na(my_data$exc7[my_data$exc7>16]) <- TRUE
my_data[my_data>16] <- 0
nrow(na.omit(my_data))
```
Since we've made this assumption, we have 1408 rows with complete data. 

```{r}
sum(is.na(my_data)) / (nrow(my_data)*ncol(my_data))
```
0.7% of values missing

```{r}
my_complete <- na.omit(my_data)
pr_complete <- prcomp(my_complete,center=TRUE,scale=FALSE)
plot(pr_complete)
summary(pr_complete)
```
Without any imputation, 72% of variance is concentrated on the first principal component. This seems suspiciously good, like only one variable is doing all the work.

```{r}
my_imp <- mice(my_data,printFlag = F) 
my_imp
```
As expected, only exc7 requires any imputation.

```{r}
pr <- lapply(1:5,function(x) prcomp(complete(my_imp,x),scale=FALSE,center=TRUE))
plot(pr[[1]])
summary(pr[[1]])
```
With the imputation, we start seeing a slight change, with 61.3% of variance concentrated on the first principal component.

```{r}
pc1 <- pr[[1]]$x[,1]
pc2 <- pr[[1]]$x[,2]
qplot(pc1,pc2)
```

This isn't cool. Our PC1 is pretty much completely determined by a single variable with four levels -- based on the last iteration we can be pretty sure that it's `exc7`. What happens if we just take it out, so that we're looking at people's actual experiences with corruption rather than their perception of it?

```{r}
pr_noexc7 <- prcomp(test_data,scale=FALSE,center=TRUE)
plot(pr_noexc7)
summary(pr_noexc7) # 48% of variance in the first component -- not bad, considering
qplot(pr_noexc7$x[,1],pr_noexc7$x[,2])
```

Now we're seeing one binary variable that really dominates. If you look up at the previous PC1/PC2 plot, this was probably the one that dominated PC2.

If our real question is "how much extorsion are people experiencing?", then it might be easier to just look at a sum of the types of bribes any one person was asked to pay.

```{r}
ex_sum <- rowSums(test_data)
plotme <- data.frame(x=ex_sum,y=my_data$exc7)
ggplot(plotme,aes(x=x,y=y)) +
  geom_jitter(color='tomato3',size=5,alpha=0.2) +
  theme_classic() +
  ylab('exc7') +
  xlab('Sum of other variables') +
  theme(text=element_text(size=20)) 
my_lm <- lm(data=plotme, y ~ x)
summary(my_lm)
```

While the plot seems to show a trend that people who think extortion is a serious problem (exc7 = 1 or 2) paid more bribes (high sum), the linear regression doesn't necessarily support this (low R-squared, high p-value). Most people seem to think corruption is a serious problem, and most of the people who think this aren't admitting to paying bribes.

Is there a relationship between `exc7` and any of the individual extortion variables?

```{r}
t.test(my_data[my_data$exc2==1,'exc7'],my_data[my_data$exc2==0,'exc7'],alternative="less")$p.value
t.test(my_data[my_data$exc6==1,'exc7'],my_data[my_data$exc6==0,'exc7'],alternative="less")$p.value
t.test(my_data[my_data$exc11==1,'exc7'],my_data[my_data$exc11==0,'exc7'],alternative="less")$p.value
t.test(my_data[my_data$exc13==1,'exc7'],my_data[my_data$exc13==0,'exc7'],alternative="less")$p.value
t.test(my_data[my_data$exc14==1,'exc7'],my_data[my_data$exc14==0,'exc7'],alternative="less")$p.value
t.test(my_data[my_data$exc15==1,'exc7'],my_data[my_data$exc15==0,'exc7'],alternative="less")$p.value
t.test(my_data[my_data$exc16==1,'exc7'],my_data[my_data$exc16==0,'exc7'],alternative="less")$p.value
t.test(my_data[my_data$exc20==1,'exc7'],my_data[my_data$exc20==0,'exc7'],alternative="less")$p.value
```

So only `exc2` has a correlation that is even borderline-significant. I think it might be time to kick `exc7` out of this index -- it doesn't have much to do with the others.

```{r}
pc1_noexc7 <- pr_noexc7$x[,1]
quantile(pc1_noexc7)
mean(pc1_noexc7 == min(pc1_noexc7))
colMeans(my_data[pc1_noexc7 == min(pc1_noexc7),],na.rm=TRUE)
colMeans(my_data[pc1_noexc7 > quantile(pc1_noexc7)[4],],na.rm=TRUE)
```

The people with the minimum value of the composite index (79% of the population) didn't pay any bribes, but are only slightly less likely to consider corruption a problem than the people in the upper quartile, the majority of whom have paid at least one.

```{r}
pc1_norm <- scale(pc1_noexc7) 
predict_data <- data.frame(exc2=c(0,1),exc6=c(0,1),exc20=c(0,1),exc11=c(0,1),
                           exc13=c(0,1),exc14=c(0,1),exc15=c(0,1),exc16=c(0,1))
minmax <- predict(pr_noexc7,predict_data)[,1]
minmax <- (minmax - mean(pc1_noexc7)) / sd(pc1_noexc7)  

predict_data2 <- data.frame(diag(8))
names(predict_data2) <- names(predict_data)
scores <- predict(pr_noexc7,predict_data2)[,1]
scores <- (scores-mean(pc1_noexc7)) / sd(pc1_noexc7) 
diff <- data.frame(d=(scores - minmax[1]),n=names(predict_data))
ggplot(diff,aes(x=n,y=d)) +
  geom_bar(stat='identity',fill='skyblue') +
  coord_flip() +
  theme_classic() +
  theme(text=element_text(size=20),
        axis.title.y=element_blank(),
        axis.title.x=element_blank())
```

`exc2` and `exc6` have a strong influence, the others not as much.

I can't get the logistic regression calculations to work, so we'll do this using a t-test instead.

```{r}
t.test(pc1_norm[my_data$exc2==1],pc1_norm[my_data$exc2==0])$p.value
t.test(pc1_norm[my_data$exc6==1],pc1_norm[my_data$exc6==0])$p.value
t.test(pc1_norm[my_data$exc11==1],pc1_norm[my_data$exc11==0])$p.value
t.test(pc1_norm[my_data$exc13==1],pc1_norm[my_data$exc13==0])$p.value
t.test(pc1_norm[my_data$exc14==1],pc1_norm[my_data$exc14==0])$p.value
t.test(pc1_norm[my_data$exc15==1],pc1_norm[my_data$exc15==0])$p.value
t.test(pc1_norm[my_data$exc16==1],pc1_norm[my_data$exc16==0])$p.value
t.test(pc1_norm[my_data$exc20==1],pc1_norm[my_data$exc20==0])$p.value
```

`exc2`, `exc6`, `exc11`, and `exc13` are unquestionably significant; the others are borderline (or worse). Does our ex_sum variable do any better?

```{r}
t.test(ex_sum[my_data$exc2==1],ex_sum[my_data$exc2==0])$p.value
t.test(ex_sum[my_data$exc6==1],ex_sum[my_data$exc6==0])$p.value
t.test(ex_sum[my_data$exc11==1],ex_sum[my_data$exc11==0])$p.value
t.test(ex_sum[my_data$exc13==1],ex_sum[my_data$exc13==0])$p.value
t.test(ex_sum[my_data$exc14==1],ex_sum[my_data$exc14==0])$p.value
t.test(ex_sum[my_data$exc15==1],ex_sum[my_data$exc15==0])$p.value
t.test(ex_sum[my_data$exc16==1],ex_sum[my_data$exc16==0])$p.value
t.test(ex_sum[my_data$exc20==1],ex_sum[my_data$exc20==0])$p.value
```

This time they're all significant. So if we want one number that captures people's experience of extortion, this might be our best choice. 

```{r}
exc2 <- data.frame(q=my_data$exc2,w=ex_sum)
exc2 <- na.omit(exc2)
my_lm <- lm(data=exc2, q ~ w)
summary(my_lm)
ggplot(exc2,aes(x=w,y=q)) +
  geom_jitter(color='tomato3',size=5,alpha=0.2) +
  theme_classic() +
  geom_smooth(method='lm',size=2,color='royalblue') +
  annotate('text',label=paste("R-sq:",round(summary(my_lm)$r.squared,2)),  
           size=10,x=-0.5,y=1,hjust=0,vjust=0,color='royalblue') +
  ylab('exc2') +
  xlab('Simple sum') +
  theme(text=element_text(size=20)) 
```

```{r}
exc14 <- data.frame(q=my_data$exc14,w=ex_sum)
exc14 <- na.omit(exc14)
my_lm <- lm(data=exc14, q ~ w)
summary(my_lm)
ggplot(exc14,aes(x=w,y=q)) +
  geom_jitter(color='tomato3',size=5,alpha=0.2) +
  theme_classic() +
  geom_smooth(method='lm',size=2,color='royalblue') +
  annotate('text',label=paste("R-sq:",round(summary(my_lm)$r.squared,2)),  
           size=10,x=-0.5,y=1,hjust=0,vjust=0,color='royalblue') +
  ylab('exc14') +
  xlab('Simple sum') +
  theme(text=element_text(size=20)) 
```

##El Salvador##

* `exc2`: Police Officer asked for a bribe (0=No, 1=Yes)
* `exc6`: Government Employee asked for a bribe (0=No, 1=Yes)
* `exc20`: Soldier requested a bribe (0=No, 1=Yes)
* `exc11`: Bribe requested to process document in Municipal Government (0=No, 1=Yes)
* `exc13`: Asked to pay a bribe at work (0=No, 1=Yes)
* `exc14`: Asked to pay a bribe at the courts (0=No, 1=Yes)
* `exc15`: Asked to pay a bribe to use health services (0=No, 1=Yes)
* `exc16`: Asked to pay a bribe at school (0=No, 1=Yes)
* `exc7`: Perception of public corruption (1=common, 4=uncommon)
(Starting on page 57 of El Salvador codebook)

```{r}
library(ggplot2)
library(mice)
library(plyr) # nice for re-formatting data
library(GGally) # for plot matrices
set.seed(12345) # makes imputation reproducible
```
```{r}
my_data <- lapop.2014.SLV[,c('exc2','exc6','exc20','exc11','exc13','exc14',
                             'exc15','exc16','exc7')]
```

```{r}
test_data <- lapop.2014.SLV[,c('exc2','exc6','exc20','exc11','exc13','exc14',
                               'exc15','exc16')]
test_data[test_data>16] <- 0 # assume that non-responses mean no bribe paid
colMeans(test_data)
jitter_data <- data.frame(police=jitter(test_data$police,factor=2),
                          gov=jitter(test_data$gov,factor=2),
                          soldier=jitter(test_data$soldier,factor=2),
                          muni=jitter(test_data$muni,factor=2),
                          work=jitter(test_data$work,factor=2),
                          courts=jitter(test_data$courts,factor=2),
                          health=jitter(test_data$health,factor=2),
                          school=jitter(test_data$school,factor=2))
ggpairs(jitter_data) + theme_classic()
```

Unlike, the previous work on this index for Guatemala, I do not seem to be getting a plot for this command; which will also make it difficult for me to see the possibles pairs of bribes. However, the amount of reported extorsions from the police amounts to approximately 5.4% of people. This is still the highest reported occasion of extorsion for El Salvador, based on the variables that are provided by the LAPOP study. 

```{r}
is.na(my_data$exc7[my_data$exc7>16]) <- TRUE
my_data[my_data>16] <- 0
nrow(na.omit(my_data))
```

Following the same logic as above on line 49, we have 1,482 rows with complete data.

```{r}
sum(is.na(my_data)) / (nrow(my_data)*ncol(my_data))
```

O.2% of data missing.

```{r}
my_complete <- na.omit(my_data)
pr_complete <- prcomp(my_complete,center=TRUE,scale=FALSE)
plot(pr_complete)
summary(pr_complete)
```

Without any imputation, 87.5% of variance is concentrated on the first principal component, which, much like the work above, also gives the impression that one variable is doing all of the work.

```{r}
my_imp <- mice(my_data,printFlag = F) 
my_imp
```

And not surprising, only `exc7`needs imputation out of all the variables listed.

```{r}
pr <- lapply(1:5,function(x) prcomp(complete(my_imp,x),scale=FALSE,center=TRUE))
plot(pr[[1]])
summary(pr[[1]])
```

With imputation, 86.7% of variance is concentrated on the first principal component. By this time, I may need to check with you, if you are getting the same result, since the calculations for this command did not yield the same result for me on the index for Guatemala.

```{r}
pc1 <- pr[[1]]$x[,1]
pc2 <- pr[[1]]$x[,2]
qplot(pc1,pc2)
```
I am getting a similar reading for this plot graph, as with Guatemala, and I can safely assume that `exc7`is dominating this index.

```{r}
pr_noexc7 <- prcomp(test_data,scale=FALSE,center=TRUE)
plot(pr_noexc7)
summary(pr_noexc7) # 41.5% of variance in the first component 
qplot(pr_noexc7$x[,1],pr_noexc7$x[,2])
```
And hmmm, this is different. 

So, what is the sum that we get for the types of bribes that any one person was asked to fork over?

```{r}
ex_sum <- rowSums(test_data)
plotme <- data.frame(x=ex_sum,y=my_data$exc7)
ggplot(plotme,aes(x=x,y=y)) +
  geom_jitter(color='tomato3',size=5,alpha=0.2) +
  theme_classic() +
  ylab('exc7') +
  xlab('Sum of other variables') +
  theme(text=element_text(size=20)) 
my_lm <- lm(data=plotme, y ~ x)
summary(my_lm)
```
My reading may be wrong, and I will welcome all corrections, but as far as I can see, the majority of people do believe that corruption is a problem (`exc7`). FYI, the results show a low r-squared value and a high p-value.

What about relationships between `exc7`and other variables?

```{r}
t.test(my_data[my_data$exc2==1,'exc7'],my_data[my_data$exc2==0,'exc7'],alternative="less")$p.value
t.test(my_data[my_data$exc6==1,'exc7'],my_data[my_data$exc6==0,'exc7'],alternative="less")$p.value
t.test(my_data[my_data$exc11==1,'exc7'],my_data[my_data$exc11==0,'exc7'],alternative="less")$p.value
t.test(my_data[my_data$exc13==1,'exc7'],my_data[my_data$exc13==0,'exc7'],alternative="less")$p.value
t.test(my_data[my_data$exc14==1,'exc7'],my_data[my_data$exc14==0,'exc7'],alternative="less")$p.value
t.test(my_data[my_data$exc15==1,'exc7'],my_data[my_data$exc15==0,'exc7'],alternative="less")$p.value
t.test(my_data[my_data$exc16==1,'exc7'],my_data[my_data$exc16==0,'exc7'],alternative="less")$p.value
t.test(my_data[my_data$exc20==1,'exc7'],my_data[my_data$exc20==0,'exc7'],alternative="less")$p.value
```
`exc2`has a borderline-significant correlation, but `exc14` on the other hand! The correlation looks to be epically strong!

```{r}
pc1_noexc7 <- pr_noexc7$x[,1]
quantile(pc1_noexc7)
mean(pc1_noexc7 == min(pc1_noexc7))
colMeans(my_data[pc1_noexc7 == min(pc1_noexc7),],na.rm=TRUE)
colMeans(my_data[pc1_noexc7 > quantile(pc1_noexc7)[4],],na.rm=TRUE)
```

```{r}
pc1_norm <- scale(pc1_noexc7) 
predict_data <- data.frame(exc2=c(0,1),exc6=c(0,1),exc20=c(0,1),exc11=c(0,1),
                           exc13=c(0,1),exc14=c(0,1),exc15=c(0,1),exc16=c(0,1))
minmax <- predict(pr_noexc7,predict_data)[,1]
minmax <- (minmax - mean(pc1_noexc7)) / sd(pc1_noexc7)  

predict_data2 <- data.frame(diag(8))
names(predict_data2) <- names(predict_data)
scores <- predict(pr_noexc7,predict_data2)[,1]
scores <- (scores-mean(pc1_noexc7)) / sd(pc1_noexc7) 
diff <- data.frame(d=(scores - minmax[1]),n=names(predict_data))
ggplot(diff,aes(x=n,y=d)) +
  geom_bar(stat='identity',fill='skyblue') +
  coord_flip() +
  theme_classic() +
  theme(text=element_text(size=20),
        axis.title.y=element_blank(),
        axis.title.x=element_blank())
```
`exc2`, followed by `exc6`largely influence the score here. Time for the t-test!

```{r}
t.test(pc1_norm[my_data$exc2==1],pc1_norm[my_data$exc2==0])$p.value
t.test(pc1_norm[my_data$exc6==1],pc1_norm[my_data$exc6==0])$p.value
t.test(pc1_norm[my_data$exc11==1],pc1_norm[my_data$exc11==0])$p.value
t.test(pc1_norm[my_data$exc13==1],pc1_norm[my_data$exc13==0])$p.value
t.test(pc1_norm[my_data$exc14==1],pc1_norm[my_data$exc14==0])$p.value
t.test(pc1_norm[my_data$exc15==1],pc1_norm[my_data$exc15==0])$p.value
t.test(pc1_norm[my_data$exc16==1],pc1_norm[my_data$exc16==0])$p.value
t.test(pc1_norm[my_data$exc20==1],pc1_norm[my_data$exc20==0])$p.value
```
`exc2`, `exc6`and `exc20`are largely significant, while `exc11`and `exc15`are borderline significant. How about an ex_sum variable?

```{r}
t.test(ex_sum[my_data$exc2==1],ex_sum[my_data$exc2==0])$p.value
t.test(ex_sum[my_data$exc6==1],ex_sum[my_data$exc6==0])$p.value
t.test(ex_sum[my_data$exc11==1],ex_sum[my_data$exc11==0])$p.value
t.test(ex_sum[my_data$exc13==1],ex_sum[my_data$exc13==0])$p.value
t.test(ex_sum[my_data$exc14==1],ex_sum[my_data$exc14==0])$p.value
t.test(ex_sum[my_data$exc15==1],ex_sum[my_data$exc15==0])$p.value
t.test(ex_sum[my_data$exc16==1],ex_sum[my_data$exc16==0])$p.value
t.test(ex_sum[my_data$exc20==1],ex_sum[my_data$exc20==0])$p.value
```
This time, all but `exc14`are largely significant.

```{r}
exc2 <- data.frame(q=my_data$exc2,w=ex_sum)
exc2 <- na.omit(exc2)
my_lm <- lm(data=exc2, q ~ w)
summary(my_lm)
ggplot(exc2,aes(x=w,y=q)) +
  geom_jitter(color='tomato3',size=5,alpha=0.2) +
  theme_classic() +
  geom_smooth(method='lm',size=2,color='royalblue') +
  annotate('text',label=paste("R-sq:",round(summary(my_lm)$r.squared,2)),  
           size=10,x=-0.5,y=1,hjust=0,vjust=0,color='royalblue') +
  ylab('exc2') +
  xlab('Simple sum') +
  theme(text=element_text(size=20)) 
```

```{r}
exc20 <- data.frame(q=my_data$exc20,w=ex_sum)
exc20 <- na.omit(exc20)
my_lm <- lm(data=exc20, q ~ w)
summary(my_lm)
ggplot(exc20,aes(x=w,y=q)) +
  geom_jitter(color='tomato3',size=5,alpha=0.2) +
  theme_classic() +
  geom_smooth(method='lm',size=2,color='royalblue') +
  annotate('text',label=paste("R-sq:",round(summary(my_lm)$r.squared,2)),  
           size=10,x=-0.5,y=1,hjust=0,vjust=0,color='royalblue') +
  ylab('exc20') +
  xlab('Simple sum') +
  theme(text=element_text(size=20)) 
```

##Honduras##

* `exc2`: Police Officer asked for a bribe (0=No, 1=Yes)
* `exc6`: Government Employee asked for a bribe (0=No, 1=Yes)
* `exc20`: Soldier requested a bribe (0=No, 1=Yes)
* `exc11`: Bribe requested to process document in Municipal Government (0=No, 1=Yes)
* `exc13`: Asked to pay a bribe at work (0=No, 1=Yes)
* `exc14`: Asked to pay a bribe at the courts (0=No, 1=Yes)
* `exc15`: Asked to pay a bribe to use health services (0=No, 1=Yes)
* `exc16`: Asked to pay a bribe at school (0=No, 1=Yes)
* `exc7`: Perception of public corruption (1=common, 4=uncommon)
(Starting on page 79 of the Honduras codebook)

```{r}
library(ggplot2)
library(mice)
library(plyr) # nice for re-formatting data
library(GGally) # for plot matrices
set.seed(12345) # makes imputation reproducible
```

```{r}
my_data <- lapop.2014.HND[,c('exc2','exc6','exc20','exc11','exc13','exc14',
                             'exc15','exc16','exc7')]
```

```{r}
test_data <- lapop.2014.HND[,c('exc2','exc6','exc20','exc11','exc13','exc14',
                               'exc15','exc16')]
test_data[test_data>16] <- 0 # assume that non-responses mean no bribe paid
colMeans(test_data)
jitter_data <- data.frame(police=jitter(test_data$police,factor=2),
                          gov=jitter(test_data$gov,factor=2),
                          soldier=jitter(test_data$soldier,factor=2),
                          muni=jitter(test_data$muni,factor=2),
                          work=jitter(test_data$work,factor=2),
                          courts=jitter(test_data$courts,factor=2),
                          health=jitter(test_data$health,factor=2),
                          school=jitter(test_data$school,factor=2))
ggpairs(jitter_data) + theme_classic()
```
Same issue with the plot graph here as with El Salvador. The one change that I can report is that of the reported extorsion cases, 15.4% were police-related.

```{r}
is.na(my_data$exc7[my_data$exc7>16]) <- TRUE
my_data[my_data>16] <- 0
nrow(na.omit(my_data))
```
Assuming that the NAs mean "no" for all variables with the exception of `exc7`, we are left with 1,511 rows with complete data.

```{r}
sum(is.na(my_data)) / (nrow(my_data)*ncol(my_data))
```
0.36% of data missing.

```{r}
my_complete <- na.omit(my_data)
pr_complete <- prcomp(my_complete,center=TRUE,scale=FALSE)
plot(pr_complete)
summary(pr_complete)
```
Without any imputation, 68% of variance is concentrated in the first principal component, and while things may possibily point to one variable doing all the work, my intuition also tells me that I am to expect some changes compared to the last two indeces.

```{r}
my_imp <- mice(my_data,printFlag = F) 
my_imp
```
Of all the variables, only `exc7`needs imputation, and not as much as it did in the last two indeces.

```{r}
pr <- lapply(1:5,function(x) prcomp(complete(my_imp,x),scale=FALSE,center=TRUE))
plot(pr[[1]])
summary(pr[[1]])
```
A slight increase, with 68.4% of variance concentrated on the first principal component, following imputations.

```{r}
pc1 <- pr[[1]]$x[,1]
pc2 <- pr[[1]]$x[,2]
qplot(pc1,pc2)
```

```{r}
pr_noexc7 <- prcomp(test_data,scale=FALSE,center=TRUE)
plot(pr_noexc7)
summary(pr_noexc7) # 44.1% of variance in the first component 
qplot(pr_noexc7$x[,1],pr_noexc7$x[,2])
```
Safe to assume that one variable is dominating?

```{r}
ex_sum <- rowSums(test_data)
plotme <- data.frame(x=ex_sum,y=my_data$exc7)
ggplot(plotme,aes(x=x,y=y)) +
  geom_jitter(color='tomato3',size=5,alpha=0.2) +
  theme_classic() +
  ylab('exc7') +
  xlab('Sum of other variables') +
  theme(text=element_text(size=20)) 
my_lm <- lm(data=plotme, y ~ x)
summary(my_lm)
```
It is overwhelming to see just how many people regard corruption to be a problem (`exc7= 1 or 2), regardless of whether they were forced to pay a bribe or not. 

So, how correlated are all of the variables with `exc7`?

```{r}
t.test(my_data[my_data$exc2==1,'exc7'],my_data[my_data$exc2==0,'exc7'],alternative="less")$p.value
t.test(my_data[my_data$exc6==1,'exc7'],my_data[my_data$exc6==0,'exc7'],alternative="less")$p.value
t.test(my_data[my_data$exc11==1,'exc7'],my_data[my_data$exc11==0,'exc7'],alternative="less")$p.value
t.test(my_data[my_data$exc13==1,'exc7'],my_data[my_data$exc13==0,'exc7'],alternative="less")$p.value
t.test(my_data[my_data$exc14==1,'exc7'],my_data[my_data$exc14==0,'exc7'],alternative="less")$p.value
t.test(my_data[my_data$exc15==1,'exc7'],my_data[my_data$exc15==0,'exc7'],alternative="less")$p.value
t.test(my_data[my_data$exc16==1,'exc7'],my_data[my_data$exc16==0,'exc7'],alternative="less")$p.value
t.test(my_data[my_data$exc20==1,'exc7'],my_data[my_data$exc20==0,'exc7'],alternative="less")$p.value
```
And there would seem to be no correlation at all, this time.

```{r}
pc1_noexc7 <- pr_noexc7$x[,1]
quantile(pc1_noexc7)
mean(pc1_noexc7 == min(pc1_noexc7))
colMeans(my_data[pc1_noexc7 == min(pc1_noexc7),],na.rm=TRUE)
colMeans(my_data[pc1_noexc7 > quantile(pc1_noexc7)[4],],na.rm=TRUE)
```

```{r}
pc1_norm <- scale(pc1_noexc7) 
predict_data <- data.frame(exc2=c(0,1),exc6=c(0,1),exc20=c(0,1),exc11=c(0,1),
                           exc13=c(0,1),exc14=c(0,1),exc15=c(0,1),exc16=c(0,1))
minmax <- predict(pr_noexc7,predict_data)[,1]
minmax <- (minmax - mean(pc1_noexc7)) / sd(pc1_noexc7)  

predict_data2 <- data.frame(diag(8))
names(predict_data2) <- names(predict_data)
scores <- predict(pr_noexc7,predict_data2)[,1]
scores <- (scores-mean(pc1_noexc7)) / sd(pc1_noexc7) 
diff <- data.frame(d=(scores - minmax[1]),n=names(predict_data))
ggplot(diff,aes(x=n,y=d)) +
  geom_bar(stat='identity',fill='skyblue') +
  coord_flip() +
  theme_classic() +
  theme(text=element_text(size=20),
        axis.title.y=element_blank(),
        axis.title.x=element_blank())
```
`exc2`, `exc6`and `exc20`largely influence the score. The case is not so much for `exc14`.
What does the t.test tell us about the variables?

```{r}
t.test(pc1_norm[my_data$exc2==1],pc1_norm[my_data$exc2==0])$p.value
t.test(pc1_norm[my_data$exc6==1],pc1_norm[my_data$exc6==0])$p.value
t.test(pc1_norm[my_data$exc11==1],pc1_norm[my_data$exc11==0])$p.value
t.test(pc1_norm[my_data$exc13==1],pc1_norm[my_data$exc13==0])$p.value
t.test(pc1_norm[my_data$exc14==1],pc1_norm[my_data$exc14==0])$p.value
t.test(pc1_norm[my_data$exc15==1],pc1_norm[my_data$exc15==0])$p.value
t.test(pc1_norm[my_data$exc16==1],pc1_norm[my_data$exc16==0])$p.value
t.test(pc1_norm[my_data$exc20==1],pc1_norm[my_data$exc20==0])$p.value
```
All of the variables are either very significant or borderline-significant.
How will this change with an ex_sum variable?

```{r}
t.test(ex_sum[my_data$exc2==1],ex_sum[my_data$exc2==0])$p.value
t.test(ex_sum[my_data$exc6==1],ex_sum[my_data$exc6==0])$p.value
t.test(ex_sum[my_data$exc11==1],ex_sum[my_data$exc11==0])$p.value
t.test(ex_sum[my_data$exc13==1],ex_sum[my_data$exc13==0])$p.value
t.test(ex_sum[my_data$exc14==1],ex_sum[my_data$exc14==0])$p.value
t.test(ex_sum[my_data$exc15==1],ex_sum[my_data$exc15==0])$p.value
t.test(ex_sum[my_data$exc16==1],ex_sum[my_data$exc16==0])$p.value
t.test(ex_sum[my_data$exc20==1],ex_sum[my_data$exc20==0])$p.value
```
This time, they are all largely significant!

```{r}
exc2 <- data.frame(q=my_data$exc2,w=ex_sum)
exc2 <- na.omit(exc2)
my_lm <- lm(data=exc2, q ~ w)
summary(my_lm)
ggplot(exc2,aes(x=w,y=q)) +
  geom_jitter(color='tomato3',size=5,alpha=0.2) +
  theme_classic() +
  geom_smooth(method='lm',size=2,color='royalblue') +
  annotate('text',label=paste("R-sq:",round(summary(my_lm)$r.squared,2)),  
           size=10,x=-0.5,y=1,hjust=0,vjust=0,color='royalblue') +
  ylab('exc2') +
  xlab('Simple sum') +
  theme(text=element_text(size=20)) 
```

```{r}
exc16 <- data.frame(q=my_data$exc16,w=ex_sum)
exc16 <- na.omit(exc16)
my_lm <- lm(data=exc16, q ~ w)
summary(my_lm)
ggplot(exc16,aes(x=w,y=q)) +
  geom_jitter(color='tomato3',size=5,alpha=0.2) +
  theme_classic() +
  geom_smooth(method='lm',size=2,color='royalblue') +
  annotate('text',label=paste("R-sq:",round(summary(my_lm)$r.squared,2)),  
           size=10,x=-0.5,y=1,hjust=0,vjust=0,color='royalblue') +
  ylab('exc16') +
  xlab('Simple sum') +
  theme(text=element_text(size=20)) 
```
