Adding to the list of indices, the following is a proposed Extorsion index. The variables chosen represent questions that relate to the request for a bribe from any civil or governmental institution. (Beginning on page 51 of codebook PDF.)

* `exc2`: Police Officer asked for a bribe (0=No, 1=Yes)
* `exc6`: Government Employee asked for a bribe (0=No, 1=Yes)
* `exc20`: Soldier requested a bribe (0=No, 1=Yes)
* `exc11`: Bribe requested to process document in Municipal Government (0=No, 1=Yes)
* `exc13`: Asked to pay a bribe at work (0=No, 1=Yes)
* `exc14`: Asked to pay a bribe at the courts (0=No, 1=Yes)
* `exc15`: Asked to pay a bribe to use health services (0=No, 1=Yes)
* `exc16`: Asked to pay a bribe at school (0=No, 1=Yes)
* `exc7`: Perception of public corruption (1=common, 4=uncommon)

```{r, message=FALSE}
library(ggplot2)
library(mice)
library(plyr) # nice for re-formatting data
library(GGally) # for plot matrices
set.seed(12345) # makes imputation reproducible
```

```{r}
lapop.2014.GTM <- read.csv("../GTM-2014.csv")
my_data <- lapop.2014.GTM[,c('exc2','exc6','exc20','exc11','exc13','exc14',
                             'exc15','exc16','exc7')]
```

Since many of these questions are asking people to report on rare events, we should start by looking to see whether there are any that co-occur.

```{r}
#test_data <- lapop.2014.GTM[,c('exc2','exc6','exc20','exc11','exc13','exc14',
#                               'exc15','exc16')]
#test_data[test_data>16] <- 0 # assume that non-responses mean no bribe paid
#colMeans(test_data)
#jitter_data <- data.frame(police=jitter(test_data$police,factor=2),
##                          gov=jitter(test_data$gov,factor=2),
#                          soldier=jitter(test_data$soldier,factor=2),
#                          muni=jitter(test_data$muni,factor=2),
#                          work=jitter(test_data$work,factor=2),
#                          courts=jitter(test_data$courts,factor=2),
#                          health=jitter(test_data$health,factor=2),
#                          school=jitter(test_data$school,factor=2))
#ggpairs(jitter_data) + theme_classic()
```

It's really disturbing that 14% of people report having paid bribes to the police. The plot is a little confusing, but the main thing to take away from it is that, for each possible pair of bribes, there are at least a few people who paid both. This gives me some hope that this could be the basis for a composite index. 

Because there are so many NA responses, let's try assuming that NA means 'no' for these 8 indices, and only impute exc7.

```{r}
is.na(my_data$exc7[my_data$exc7>16]) <- TRUE
my_data[my_data>16] <- 0
nrow(na.omit(my_data))
```
Since we've made this assumption, we have 1408 rows with complete data. 

```{r}
sum(is.na(my_data)) / (nrow(my_data)*ncol(my_data))
```
0.7% of values missing

```{r}
my_complete <- na.omit(my_data)
pr_complete <- prcomp(my_complete,center=TRUE,scale=FALSE)
plot(pr_complete)
summary(pr_complete)
```
Without any imputation, 72% of variance is concentrated on the first principal component. This seems suspiciously good, like only one variable is doing all the work.

```{r}
my_imp <- mice(my_data,printFlag = F) 
my_imp
```
As expected, only exc7 requires any imputation.

```{r}
pr <- lapply(1:5,function(x) prcomp(complete(my_imp,x),scale=FALSE,center=TRUE))
plot(pr[[1]])
summary(pr[[1]])
```
With the imputation, we start seeing a slight change, with 61.3% of variance concentrated on the first principal component.

```{r}
pc1 <- pr[[1]]$x[,1]
pc2 <- pr[[1]]$x[,2]
qplot(pc1,pc2)
```

This isn't cool. Our PC1 is pretty much completely determined by a single variable with four levels -- based on the last iteration we can be pretty sure that it's `exc7`. What happens if we just take it out, so that we're looking at people's actual experiences with corruption rather than their perception of it?

```{r}
#pr_noexc7 <- prcomp(test_data,scale=FALSE,center=TRUE)
#plot(pr_noexc7)
#summary(pr_noexc7) # 48% of variance in the first component -- not bad, considering
#qplot(pr_noexc7$x[,1],pr_noexc7$x[,2])
```

Now we're seeing one binary variable that really dominates. If you look up at the previous PC1/PC2 plot, this was probably the one that dominated PC2.

If our real question is "how much extorsion are people experiencing?", then it might be easier to just look at a sum of the types of bribes any one person was asked to pay.

```{r}
ex_sum <- rowSums(test_data)
plotme <- data.frame(x=ex_sum,y=my_data$exc7)
ggplot(plotme,aes(x=x,y=y)) +
  geom_jitter(color='tomato3',size=5,alpha=0.2) +
  theme_classic() +
  ylab('exc7') +
  xlab('Sum of other variables') +
  theme(text=element_text(size=20)) 
my_lm <- lm(data=plotme, y ~ x)
summary(my_lm)
```

While the plot seems to show a trend that people who think extortion is a serious problem (exc7 = 1 or 2) paid more bribes (high sum), the linear regression doesn't necessarily support this (low R-squared, high p-value). Most people seem to think corruption is a serious problem, and most of the people who think this aren't admitting to paying bribes.

Is there a relationship between `exc7` and any of the individual extortion variables?

```{r}
t.test(my_data[my_data$exc2==1,'exc7'],my_data[my_data$exc2==0,'exc7'],alternative="less")$p.value
t.test(my_data[my_data$exc6==1,'exc7'],my_data[my_data$exc6==0,'exc7'],alternative="less")$p.value
t.test(my_data[my_data$exc11==1,'exc7'],my_data[my_data$exc11==0,'exc7'],alternative="less")$p.value
t.test(my_data[my_data$exc13==1,'exc7'],my_data[my_data$exc13==0,'exc7'],alternative="less")$p.value
t.test(my_data[my_data$exc14==1,'exc7'],my_data[my_data$exc14==0,'exc7'],alternative="less")$p.value
t.test(my_data[my_data$exc15==1,'exc7'],my_data[my_data$exc15==0,'exc7'],alternative="less")$p.value
t.test(my_data[my_data$exc16==1,'exc7'],my_data[my_data$exc16==0,'exc7'],alternative="less")$p.value
t.test(my_data[my_data$exc20==1,'exc7'],my_data[my_data$exc20==0,'exc7'],alternative="less")$p.value
```

So only `exc2` has a correlation that is even borderline-significant. I think it might be time to kick `exc7` out of this index -- it doesn't have much to do with the others.

```{r}
pc1_noexc7 <- pr_noexc7$x[,1]
quantile(pc1_noexc7)
mean(pc1_noexc7 == min(pc1_noexc7))
colMeans(my_data[pc1_noexc7 == min(pc1_noexc7),],na.rm=TRUE)
colMeans(my_data[pc1_noexc7 > quantile(pc1_noexc7)[4],],na.rm=TRUE)
```

The people with the minimum value of the composite index (79% of the population) didn't pay any bribes, but are only slightly less likely to consider corruption a problem than the people in the upper quartile, the majority of whom have paid at least one.

```{r}
pc1_norm <- scale(pc1_noexc7) 
predict_data <- data.frame(exc2=c(0,1),exc6=c(0,1),exc20=c(0,1),exc11=c(0,1),
                           exc13=c(0,1),exc14=c(0,1),exc15=c(0,1),exc16=c(0,1))
minmax <- predict(pr_noexc7,predict_data)[,1]
minmax <- (minmax - mean(pc1_noexc7)) / sd(pc1_noexc7)  

predict_data2 <- data.frame(diag(8))
names(predict_data2) <- names(predict_data)
scores <- predict(pr_noexc7,predict_data2)[,1]
scores <- (scores-mean(pc1_noexc7)) / sd(pc1_noexc7) 
diff <- data.frame(d=(scores - minmax[1]),n=names(predict_data))
ggplot(diff,aes(x=n,y=d)) +
  geom_bar(stat='identity',fill='skyblue') +
  coord_flip() +
  theme_classic() +
  theme(text=element_text(size=20),
        axis.title.y=element_blank(),
        axis.title.x=element_blank())
```

`exc2` and `exc6` have a strong influence, the others not as much.

I can't get the logistic regression calculations to work, so we'll do this using a t-test instead.

```{r}
t.test(pc1_norm[my_data$exc2==1],pc1_norm[my_data$exc2==0])$p.value
t.test(pc1_norm[my_data$exc6==1],pc1_norm[my_data$exc6==0])$p.value
t.test(pc1_norm[my_data$exc11==1],pc1_norm[my_data$exc11==0])$p.value
t.test(pc1_norm[my_data$exc13==1],pc1_norm[my_data$exc13==0])$p.value
t.test(pc1_norm[my_data$exc14==1],pc1_norm[my_data$exc14==0])$p.value
t.test(pc1_norm[my_data$exc15==1],pc1_norm[my_data$exc15==0])$p.value
t.test(pc1_norm[my_data$exc16==1],pc1_norm[my_data$exc16==0])$p.value
t.test(pc1_norm[my_data$exc20==1],pc1_norm[my_data$exc20==0])$p.value
```

`exc2`, `exc6`, `exc11`, and `exc13` are unquestionably significant; the others are borderline (or worse). Does our ex_sum variable do any better?

```{r}
t.test(ex_sum[my_data$exc2==1],ex_sum[my_data$exc2==0])$p.value
t.test(ex_sum[my_data$exc6==1],ex_sum[my_data$exc6==0])$p.value
t.test(ex_sum[my_data$exc11==1],ex_sum[my_data$exc11==0])$p.value
t.test(ex_sum[my_data$exc13==1],ex_sum[my_data$exc13==0])$p.value
t.test(ex_sum[my_data$exc14==1],ex_sum[my_data$exc14==0])$p.value
t.test(ex_sum[my_data$exc15==1],ex_sum[my_data$exc15==0])$p.value
t.test(ex_sum[my_data$exc16==1],ex_sum[my_data$exc16==0])$p.value
t.test(ex_sum[my_data$exc20==1],ex_sum[my_data$exc20==0])$p.value
```

This time they're all significant. So if we want one number that captures people's experience of extortion, this might be our best choice. 

```{r}
exc2 <- data.frame(q=my_data$exc2,w=ex_sum)
exc2 <- na.omit(exc2)
my_lm <- lm(data=exc2, q ~ w)
summary(my_lm)
ggplot(exc2,aes(x=w,y=q)) +
  geom_jitter(color='tomato3',size=5,alpha=0.2) +
  theme_classic() +
  geom_smooth(method='lm',size=2,color='royalblue') +
  annotate('text',label=paste("R-sq:",round(summary(my_lm)$r.squared,2)),  
           size=10,x=-0.5,y=1,hjust=0,vjust=0,color='royalblue') +
  ylab('exc2') +
  xlab('Simple sum') +
  theme(text=element_text(size=20)) 
```

```{r}
exc14 <- data.frame(q=my_data$exc14,w=ex_sum)
exc14 <- na.omit(exc14)
my_lm <- lm(data=exc14, q ~ w)
summary(my_lm)
ggplot(exc14,aes(x=w,y=q)) +
  geom_jitter(color='tomato3',size=5,alpha=0.2) +
  theme_classic() +
  geom_smooth(method='lm',size=2,color='royalblue') +
  annotate('text',label=paste("R-sq:",round(summary(my_lm)$r.squared,2)),  
           size=10,x=-0.5,y=1,hjust=0,vjust=0,color='royalblue') +
  ylab('exc14') +
  xlab('Simple sum') +
  theme(text=element_text(size=20)) 
```