Adding to the list of indices, the following is a proposed Extorsion index. The variables chosen represent questions that relate to the request for a bribe from any civil or governmental institution. (Beginning on page 51 of codebook PDF.)

* `exc2`: Police Officer asked for a bribe (0=No, 1=Yes)
* `exc6`: Government Employee asked for a bribe (0=No, 1=Yes)
* `exc20`: Soldier requested a bribe (0=No, 1=Yes)
* `exc11`: Bribe requested to process document in Municipal Government (0=No, 1=Yes)
* `exc13`: Asked to pay a bribe at work (0=No, 1=Yes)
* `exc14`: Asked to pay a bribe at the courts (0=No, 1=Yes)
* `exc15`: Asked to pay a bribe to use health services (0=No, 1=Yes)
* `exc16`: Asked to pay a bribe at school (0=No, 1=Yes)
* `exc7`: Perception of public corruption (1=common, 4=uncommon)

```{r}
library(ggplot2)
#install.packages("mice") # if you don't have it already
library(mice)
library(plyr) # nice for re-formatting data
#install.packages("GGally") # if you don't have it already
library(GGally) # for plot matrices
set.seed(12345) # makes imputation reproducible
```

```{r}
lapop.2014.GTM <- read.csv("../GTM-2014.csv")
my_data <- lapop.2014.GTM[,c('exc2','exc6','exc20','exc11','exc13','exc14','exc15','exc16','exc7')]
```

Since many of these questions are asking people to report on rare events, we should start by looking to see whether there are any that co-occur.

```{r}
test_data <- lapop.2014.GTM[,c('exc2','exc6','exc20','exc11','exc13','exc14','exc15','exc16')]
names(test_data) <- c('police','gov','soldier','muni','work','courts','health','school')
test_data[test_data>16] <- 0 # assume that non-responses mean no bribe paid
colMeans(test_data)
jitter_data <- data.frame(police=jitter(test_data$police,factor=2),
                          gov=jitter(test_data$gov,factor=2),
                          soldier=jitter(test_data$soldier,factor=2),
                          muni=jitter(test_data$muni,factor=2),
                          work=jitter(test_data$work,factor=2),
                          courts=jitter(test_data$courts,factor=2),
                          health=jitter(test_data$health,factor=2),
                          school=jitter(test_data$school,factor=2))
ggpairs(jitter_data) + theme_classic()
```

It's really disturbing that 14% of people report having paid bribes to the police. The plot is a little confusing, but the main thing to take away from it is that, for each possible pair of bribes, there are at least a few people who paid both. This gives me some hope that this could be the basis for a composite index. 

Because there are so many NA responses, let's try assuming that NA means 'no' for these 8 indices, and only impute exc7.

```{r}
is.na(my_data[my_data>16]) <- TRUE
nrow(na.omit(my_data))
```
Now, here it looks like we're running into trouble. Ignoring the missing values, we have 7 rows left. 

```{r}
sum(is.na(my_data)) / (nrow(my_data)*ncol(my_data))
```
38.2% of values missing

```{r}
my_complete <- na.omit(my_data)
pr_complete <- prcomp(my_complete,center=TRUE,scale=FALSE)
plot(pr_complete)
summary(pr_complete)
```
Without any imputation, 68% of variance is concentrated on the first principal component.

```{r}
my_imp <- mice(my_data,printFlag = F) 
my_imp
```
Now, notice how `exc11`, `exc13`, `exc14`, `exc15`, and `exc16` needed a great amount of imputation This is, because all of these questions were conditional upon the response of some previous questions. e.g. To be asked only if the response to `exc11f`= '1'. Then there is `exc98` that also was missing a lot of values, most likely as a result of people not knowing how to respond to or feeling uncomfortable to answer the question of justification of bribery.

Looks like this one will be difficult!

```{r}
pr <- lapply(1:5,function(x) prcomp(complete(my_imp,x),scale=FALSE,center=TRUE))
plot(pr[[1]])
summary(pr[[1]])
```
With the imputation, we start seeing a slight change, with 61.3% of variance concentrated on the first principal component.

```{r}
pc1 <- pr[[1]]$x[,1]
pc2 <- pr[[1]]$x[,2]
qplot(pc1,pc2)
```

```{r}
all_pc1 <- data.frame(llply(1:5, function(i) pr[[i]]$x[,1]))
names(all_pc1) <- c('imp1','imp2','imp3','imp4','imp5')
ggpairs(all_pc1) + theme_classic()
```
Overall, some good correlations, but nothing perfect.

```{r}
all_pc1$avg <- rowMeans(all_pc1)
quantile(all_pc1$avg)
colMeans(my_data[all_pc1$avg < quantile(all_pc1$avg)[2],],na.rm=TRUE)
colMeans(my_data[all_pc1$avg > quantile(all_pc1$avg)[4],],na.rm=TRUE)
```
I will need some verification here to see if my quantiles are in the proper order; i.e. higher numbers on the higher quantile, and lower numbers in the lower quantile.

```{r}
all_pc1$norm <- -scale(all_pc1$avg) 
predict_data <- data.frame(exc2=c(0,1),exc6=c(0,1),exc20=c(0,1),exc11=c(0,1),
                           exc13=c(0,1),exc14=c(0,1),exc15=c(0,1),exc16=c(0,1),
                           exc7=c(0,1))
minmax <- rowMeans(sapply(1:5, function(i) predict(pr[[i]],predict_data)[,1]))
minmax <- -(minmax - mean(all_pc1$avg)) / sd(all_pc1$avg)  

predict_data2 <- data.frame(diag(9)+1)
names(predict_data2) <- names(my_data)
scores <- rowMeans(sapply(1:5, function(i) predict(pr[[i]],predict_data2)[,1]))
scores <- -(scores-mean(all_pc1$avg)) / sd(all_pc1$avg) 
diff <- data.frame(d=(scores - minmax[1]),n=names(predict_data))
ggplot(diff,aes(x=n,y=d)) +
  geom_bar(stat='identity',fill='skyblue') +
  coord_flip() +
  theme_classic() +
  theme(text=element_text(size=20),
        axis.title.y=element_blank(),
        axis.title.x=element_blank())
```
All variables seem to have a really strong influence in the index.

```{r}
log_data <- my_data[,c('exc2','exc6','exc20','exc11','exc13','exc14','exc15','exc16')]
#log_data <- log_data - 1 #convert to 0-1 scale
p_vals <- sapply(log_data, function(x) coef(summary(glm(x ~ all_pc1$norm,family=binomial(logit))))[2,4])
p_vals
```

None of these correlations are significant.


