---
title: "Geography of fear"
author: "Craig Jolley"
date: "November 5, 2015"
output: html_document
---

Build all of the composite indices


```{r}
source('make_indices.R')
library(ggmap)
library(plyr)
library(XLConnect)
```

OK. So now we've got a fear index for all three countries. Now we need to get the relevant geographic information for each.

```{r}
geo <- c('pais','estratopri','estratosec','prov','municipio')
# HND has Distrito information; GTM doesn't
geo_gtm <- lapop.2014.GTM[,geo]
geo_hnd <- lapop.2014.HND[,geo] 
geo_slv <- lapop.2014.SLV[,geo] 

my_geo <- lapop.2014.all[,geo]
my_geo$fear <- fear_all
```

What I want to do now is aggregate by different geographic units and see how much of the variance can be explained by each level of aggregation. First, the country level (`pais`).

```{r}
pais_avg <- ddply(my_geo,~pais,summarize,x=mean(fear))
my_geo$pais_avg <- pais_avg$x[match(my_geo$pais,pais_avg$pais)]
qplot(my_geo$fear,my_geo$pais_avg)
summary(lm(fear~pais_avg,data=my_geo))
```

So there's a correlation (we would hope!), but the R-squared value is only 0.011; this isn't the right level of aggregation.

Next, let's try `estratopri`, for regional divisions within each country.

```{r}
pri_avg <- ddply(my_geo,~estratopri,summarize,x=mean(fear))
my_geo$pri_avg <- pri_avg$x[match(my_geo$estratopri,pri_avg$estratopri)]
qplot(my_geo$fear,my_geo$pri_avg)
summary(lm(fear~pri_avg,data=my_geo))
```

So now R-squared has climbed to 0.11; we're accounting for 11% of the variance. Still not all that impressive. Let's move on to the province level:

```{r}
prov_avg <- ddply(my_geo,~prov,summarize,x=mean(fear))
my_geo$prov_avg <- prov_avg$x[match(my_geo$prov,prov_avg$prov)]
qplot(my_geo$fear,my_geo$prov_avg)
summary(lm(fear~prov_avg,data=my_geo))
```

With an R-squared of 0.13, provinces are barely better than regions. Now let's get really granular and look at municipalities. Are municipality IDs unique within each country?

```{r}
length(unique(my_geo$municipio)) # 149
length(unique(my_geo[my_geo$pais==2,'municipio'])) # 55
length(unique(my_geo[my_geo$pais==3,'municipio'])) # 52 
length(unique(my_geo[my_geo$pais==4,'municipio'])) # 51 -> 158 total
```

No. This means we'll need to create some kind of unique identifier that combines `pais` and `municipio`.

```{r}
my_geo$muni_uniq <- 10000*my_geo$pais + my_geo$municipio
muni_avg <- ddply(my_geo,~muni_uniq,summarize,x=mean(fear))
my_geo$muni_avg <- muni_avg$x[match(my_geo$muni_uniq,muni_avg$muni_uniq)]
qplot(my_geo$fear,my_geo$muni_avg)
summary(lm(fear~muni_avg,data=my_geo))
```

Even municipalities, the most granular level of geographic resolution we have across all three countries, only explain 22% of the variance in our fear index. Clearly, there's more determining people's level of fear than where they live. The plot is telling, though -- in the least fearful cities, hardly anyone has high a high fear index, while the opposite is true for the most fearful cities.

In Honduras, it's possible to get even more detail and go down to the district level.

```{r}
geo_hnd$distrito <- lapop.2014.HND$hondistrito
geo_hnd$fear <- my_geo[my_geo$pais==4,'fear']
dist_avg <- ddply(geo_hnd,~distrito,summarize,x=mean(fear))
geo_hnd$dist_avg <- dist_avg$x[match(geo_hnd$distrito,dist_avg$distrito)]
qplot(geo_hnd$fear,geo_hnd$dist_avg)
summary(lm(fear~dist_avg,data=geo_hnd))
```

In Honduras, then, the availability of district-level geolocation lets us explain 30% of the variance in fear.

To facilitate mappning, let's roll this up into a data frame that will contain average values of the fear index for each province in the region. Even though this isn't very granular, I have coverage of provinces but not municipalities.

```{r}
# all this comes from place_names.R:
# Country

pais_num <- c(2,3,4)
pais_text <- c('Guatemala','El Salvador','Honduras')
my_geo$pais_text <- pais_text[match(my_geo$pais,pais_num)]

# Province
prov_num <- c(1:18,20:22,301:314,401:413,418:422)                              
prov_text <- c('Guatemala', 'El Progreso', 'Sacatepéquez', 'Chimaltenango', 
               'Escuintla', 'Santa Rosa', 'Sololá', 'Totonicapán', 
               'Quetzaltenango', 'Suchitepéquez', 'Retalhuleu', 'San Marcos', 
               'Huehuetenango', 'Quiché','Baja Verapaz', 'Alta Verapaz', 
               'Petén', 'Izabal', 'Chiquimula', 'Jalapa', 'Jutiapa', 
               'Ahuachapan', 'Santa Ana', 'Sonsonate', 'Chalatenango', 
               'La Libertad','San Salvador', 'Cuscatlan', 'La Paz', 'Cabañas', 
               'San Vicente', 'Usulutan', 'San Miguel', 'Morazan', 'La Union', 
               'Francisco Morazán','Comayagua', 'La Paz', 'Cortés', 
               'Atlántida', 'Colón', 'Yoro', 'Islas de la Bahía', 'Copán', 
               'Intibucá', 'Lempira', 'Ocotepeque', 'Santa Bárbara', 
               'El Paraíso', 'Olancho', 'Gracias a Dios', 'Choluteca', 'Valle')
my_geo$prov_text <- prov_text[match(my_geo$prov,prov_num)]

# Municipality
muni_num_hnd <- c(101,108,109,201,204,208,209,301,401,402,403,405,406,407,408,
                  409,501,503,504,505,603,604,701,702,703,704,705,805,902,905,
                  906,907,1101,1105,1203,1307,1308,1309,1310,1405,1801,1806,
                  1807,1902,1908,2001,2101,2107,2108,2201,2205)
muni_num_gtm <- c(101,106,107,108,110,115,116,117,203,301,314,401,413,501,502,
                  504,509,602,609,701,710,806,901,911,919,920,1001,1006,1013,
                  1104,1204,1212,1213,1301,1304,1311,1319,1401,1404,1413,1504,
                  1601,1606,1609,1614,1705,1708,1803,1805,2004,2101,2102,2201,
                  2202,2211)
muni_num_slv <- c(30101,30105,30107,30110,30201,30205,30207,30210,30302,30306,
                  30315,30407,30408,30503,30511,30512,30515,30521,30601,30602,
                  30604,30607,30608,30610,30613,30614,30617,30618,30619,30702,
                  30703,30708,30816,30821,30822,30903,30906,31002,31003,31010,
                  31111,31117,31123,31202,31209,31217,31219,31309,31312,31403,
                  31408,31416)
muni_text_hnd <- c('Distrito Central','El Porvenir','San Ignacio','Comayagua',
                   'Siguatepeque','El Rosario','San Jerónimo','La Paz',
                   'San Pedro Sula','Choloma','Puerto Cortés','San Manuel',
                   'Santa Cruz de Yojoa','Villanueva','La Lima','Potrerillos',
                   'El Porvenir','Tela','La Ceiba','Arizona','Sonaguera',
                   'Tocoa','Yoro','El Negrito','El Progreso','Morazán',
                   'Olanchito','José Santos Guardiola','El Paraíso',
                   'Nueva Arcadia','San José','Intibucá','Candelaria',
                   'San Juan Guarita','Fraternidad','Arada','Ilama',
                   'Quimistán','San Pedro Zacapa ','Santa Rosa de Copán',
                   'Danlí','Yuscarán','Teupasenti','Catacamas','Guayape',
                   'Brus Laguna','Choluteca','Duyure','Pespire','Nacaome',
                   'Caridad')
muni_text_gtm <- c('Guatemala','Chinautla','San Pedro Ayampuc','Mixco',
                   'San Juan Sacatépequez','Villa Nueva','Villa Canales',
                   'Petapa','San Agustín Acasaguastlán','Antigua Guatemala',
                   'Alotenango','Chimaltenango','San Andrés Itzapa',
                   'Escuintla','Santa Lucía Cotzumalguapa','Siquinalá',
                   'San José','Barberena','Taxisco','Sololá','Panajachel',
                   'Santa María Chiquimula','Quetzaltenango',
                   'Concepción Chiquirichapa','El Palmar','Coatepeque',
                   'Mazatenango','Santo Domingo Suchitepéquez','Chicacao',
                   'San Martín Zapotitlán','Comitancillo','Nuevo Progreso',
                   'El Tumbador','Huehuetenango','Cuilco','La Libertad',
                   'Colotenango','Santa Cruz Del Quiché','Zacualpa','Nebaj',
                   'Cubulco','Cobán','Tucurú','San Pedro Carchá','Chahal',
                   'La Libertad','Dolores','El Estor','Los Amates','Jocotán',
                   'Jalapa','San Pedro Pinula','Jutiapa','El Progreso','Comapa')
muni_text_slv <- c('Ahuachapan','El Refugio','Jujutla','San Pedro Puxtla',
                   'Candelaria De La Frontera','El Porvenir','Metapan',
                   'Santa Ana','Armenia','Izalco','Sonsonate','Chalatenango',
                   'Dulce Nombre De Maria','Colon','Santa Tecla',
                   'Quezaltepeque','San Juan Opico','Tepecoyo','Aguilares',
                   'Apopa','Cuscatancingo','Ilopango','Mejicanos','Panchimalco',
                   'San Martin','San Salvador','Soyapango','Tonacatepeque',
                   'Ciudad Delgado','Cojutepeque','El Carmen','San Cristobal',
                   'San Pedro Nonualco','Zacatecoluca','San Luis La Herradura',
                   'Ilobasco','Sensuntepeque','Guadalupe',
                   'San Cayetano Istepeque','San Vicente','Mercedes Umaña',
                   'San Dionisio','Usulutan','Ciudad Barrios','Moncagua',
                   'San Miguel','Sesori','Guatajiagua','Jocoro',
                   'Concepcion De Oriente','La Union','Santa Rosa De Lima')
my_geo$muni_text <- NA
my_geo[my_geo$pais==2,]$muni_text <- muni_text_gtm[match(my_geo[my_geo$pais==2,]$municipio,muni_num_gtm)]
my_geo[my_geo$pais==3,]$muni_text <- muni_text_slv[match(my_geo[my_geo$pais==3,]$municipio,muni_num_slv)]
my_geo[my_geo$pais==4,]$muni_text <- muni_text_hnd[match(my_geo[my_geo$pais==4,]$municipio,muni_num_hnd)]
```

Export this in a format that I'll (hopefully) be able to import into AGOL. 

```{r}
out_geo <- data.frame(country=my_geo$pais_text,
                      prov=my_geo$prov_text,
                      fear_avg=my_geo$prov_avg,
                      stringsAsFactors = FALSE)
out_geo <- unique(out_geo) # 53 unique provinces left
nrow(out_geo[out_geo$country=='Guatemala',]) # 21 of 22 in Guatemala
nrow(out_geo[out_geo$country=='El Salvador',]) # all 14 in El Salvador
nrow(out_geo[out_geo$country=='Honduras',]) # all 18 in Honduras
write.csv(out_geo,file='fear-prov.csv')
```

Try this with lat/longs also and see how that looks. For now, don't sweat using the Aldeas-level information in Honduras, so that all three countries are comparable.

```{r}
out_geo2 <- data.frame(country=my_geo$pais_text,
                      prov=my_geo$prov_text,
                      muni=my_geo$muni_text,
                      fear_avg=my_geo$muni_avg,
                      stringsAsFactors = FALSE)
out_geo2 <- unique(out_geo2) # 158 unique municipalities left

geocode_txt <- paste(out_geo2$muni,out_geo2$prov,out_geo2$country,sep=", ")
geo_ll <- ldply(geocode_txt,function(x) geocode(x,output='latlon',source='google'))
out_geo2 <- cbind(out_geo2,geo_ll)
write.csv(out_geo2,file='fear-latlong.csv')
```

**TODO:** It would probably also be useful to include a count of survey responses for each municipality and province. This could also be visualized on the map. 