---
title: "Geography of fear"
author: "Craig Jolley"
date: "November 5, 2015"
output: html_document
---

For now, I'm basing this on a version of the fear index where I'm removing `vic45` and `a4`. In addition, the following variables were not collected in El Salvador (although they were in both Honduras and Guatemala):

- `vic40`: Out of fear of crime, has limited the places to go shopping
- `vic41`: Out of fear of crime, has limited the places to go for recreation
- `vic43`: Out of fear of crime, has felt the need to change neighborhoods
- `fear6f`: Insecurity at Schools


```{r}
library(ggplot2)
library(mice)
library(plyr)
library(GGally)

lapop.2014.HND <- read.csv("../2014-HND.csv")
lapop.2014.SLV <- read.csv("../2014-SLV.csv")
lapop.2014.GTM <- read.csv("../2014-GTM.csv")
set.seed(12345) # for reproducibility

incl <- c('vic44','fear10','vic1ext',
         'vic1exta','vic1hogar','aoj11','pese1','pese2','aoj17','diso7',
         'diso8','diso10','diso18','diso14','diso16','diso17','vicbar1', 
         'vicbar1f','vicbar3','vicbar4','vicbar7')

hnd_data <- lapop.2014.HND[,incl]
slv_data <- lapop.2014.SLV[,incl]
gtm_data <- lapop.2014.GTM[,incl]
my_data <- rbind(gtm_data,hnd_data,slv_data)

my_data$vic1exta[my_data$vic1exta == 999999] <- 0
is.na(my_data[my_data>30]) <- TRUE

my_imp <- mice(my_data, printFlag = F)
pr <- lapply(1:5,function(x) prcomp(complete(my_imp,x),scale=FALSE,center=TRUE))

all_pc1 <- data.frame(llply(1:5, function(i) pr[[i]]$x[,1]))
all_pc1$imp1 <- -all_pc1$imp1
all_pc1$imp4 <- -all_pc1$imp4
all_pc1$avg <- rowMeans(all_pc1)
all_pc1$norm <- scale(-all_pc1$avg) # sign flip -- now high values mean more fear
```

Let's see how people's level of fear correlates with their answers to `a4` -- the most serious problem facing the country. This is a little tricky, since the coding is consistent between countries (e.g. `a4=4` means poverty everywhere), but not all of the possible answers were on the questionnaire for any given country. If any country-specific answers turn out to be significant, we'll have to make sure they're significant at a country level as well.

```{r}
a4 <- c(lapop.2014.HND$a4,lapop.2014.SLV$a4,lapop.2014.GTM$a4)
a4[a4==888888] <- NA
a4[a4==988888] <- NA
fear$a4 <- a4

categ_unsorted <- function(f,x,categ) {
  # For a data from f with a continuous variable 'x' and an unsorted 
  # categorical variable 'categ', test whether the value of x is 
  # significantly higher or lower for each value of categ than for the
  # rest of the population. Do this using a two-sample t-test.
  result <- data.frame(var=character(),val=integer(),pval=double(),
                       mean=double(),othermean=double())
  print(unique(f[,categ]))
  for(q in unique(na.omit(f[,categ]))) {
    if(sum(f[,categ]==q,na.rm=TRUE) > 1) {
      yes <- fear[fear[,categ]==q,x]
      no <- fear[fear[,categ]!=q,x]
      tt <- t.test(yes,no,na.rm=TRUE)
      newrow <- data.frame(var=categ,val=q,pval=tt$p.value,
                       mean=tt$estimate[[1]],
                       othermean=tt$estimate[[2]])
      result <- rbind(result,newrow)
    }
  }
  result
}

cu <- categ_unsorted(fear,'w','a4')
cu[cu$pval<0.05,]
```

So it looks like the only strong correlation here is that people who are concerned about human rights violations are significantly *less* fearful than the rest of the population. 

Another unordered categorical variable would be `for1n` (Country with the most influence in the region).

```{r}
for1n <- c(lapop.2014.HND$for1n,lapop.2014.SLV$for1n,lapop.2014.GTM$for1n)
for1n[for1n==888888] <- NA
for1n[for1n==988888] <- NA
fear$for1n <- for1n
categ_unsorted(fear,'w','for1n')
mean(for1n==4,na.rm=TRUE)
mean(for1n==2,na.rm=TRUE)
```

So the 71% of people who feel that the USA (`for1n=4`) is the most influential country in the region are significantly more fearful than the remaining 29%, while the 9% who think Japan has the most influence (`for1n==2`) are less fearful than the remaining 91%. Everything else turns out not to be significant. I don't really know what makes the people who perceive a high Japanese influence different from everyone else, but it's a pretty big difference.

