---
title: "binary"
author: "calbano"
date: "November 29, 2015"
output: html_document
---
```{r}
lapop.2014.HND <- read.csv("../HND-2014.csv",stringsAsFactors=FALSE)
lapop.2014.GTM <- read.csv("../GTM-2014.csv",stringsAsFactors=FALSE)
lapop.2014.SLV <- read.csv("../SLV-2014.csv",stringsAsFactors=FALSE)

library(mice)
library(plyr)
set.seed(12345)
```

```{r}
gtm <- lapop.2014.GTM[,c('vic40', 'vic41', 'vic44', 'vic43',
      'fear10','fear6f','vic1ext','vic1exta',
      'vic1hogar','aoj11','pese1','pese2','aoj17',
      'diso7','diso8','diso10','diso18','diso14',
      'diso16','diso17','vicbar1','vicbar1f',
      'vicbar3','vicbar4','vicbar7')]

hnd <- lapop.2014.HND[,c('vic40', 'vic41', 'vic44', 'vic43', 'vic45',
      'fear10','fear6f','vic1ext','vic1exta',
      'vic1hogar','aoj11','pese1','pese2','aoj17',
      'diso7','diso8','diso10','diso18','diso14',
      'diso16','diso17','vicbar1','vicbar1f',
      'vicbar3','vicbar4','vicbar7')]

slv <- lapop.2014.SLV[,c('vic44',
      'fear10','vic1ext','vic1exta',
      'vic1hogar','aoj11','pese1','pese2','aoj17',
      'diso7','diso8','diso10','diso18','diso14',
      'diso16','diso17','vicbar1','vicbar1f',
      'vicbar3','vicbar4','vicbar7')]

hnd$vic1exta[hnd$vic1exta == 999999] = 0
gtm$vic1exta[gtm$vic1exta == 999999] = 0
slv$vic1exta[slv$vic1exta == 999999] = 0

is.na(hnd[hnd>30]) = TRUE
is.na(gtm[gtm>30]) = TRUE
is.na(slv[slv>30]) = TRUE

complete.hnd = na.omit(hnd)
complete.gtm = na.omit(gtm)
complete.slv = na.omit(slv)

pr_hnd = prcomp(complete.hnd, center = TRUE, scale = FALSE)
pr_gtm = prcomp(complete.gtm, center = TRUE, scale = FALSE)
pr_slv = prcomp(complete.slv, center = TRUE, scale = FALSE)

my_imp_hnd <- mice(hnd, printFlag = F)
my_imp_gtm <- mice(gtm, printFlag = F)
my_imp_slv <- mice(slv, printFlag = F)

pr.hnd <- lapply(1:5,function(x) prcomp(complete(my_imp_hnd,x),scale=FALSE,center=TRUE))
pr.gtm <- lapply(1:5,function(x) prcomp(complete(my_imp_gtm,x),scale=FALSE,center=TRUE))
pr.slv <- lapply(1:5,function(x) prcomp(complete(my_imp_slv,x),scale=FALSE,center=TRUE))

all_pc1_hnd <- data.frame(llply(1:5, function(i) pr.hnd[[i]]$x[,1]))
all_pc1_gtm <- data.frame(llply(1:5, function(i) pr.gtm[[i]]$x[,1]))
all_pc1_slv <- data.frame(llply(1:5, function(i) pr.slv[[i]]$x[,1]))

all_pc1_hnd$avg <- rowMeans(all_pc1_hnd)
all_pc1_gtm$avg <- rowMeans(all_pc1_gtm)
all_pc1_slv$avg <- rowMeans(all_pc1_slv)

all_pc1_hnd$norm <- scale(-all_pc1_hnd$avg) # sign flip -- now high values mean more fear
all_pc1_gtm$norm <- scale(-all_pc1_gtm$avg) # sign flip -- now high values mean more fear
all_pc1_slv$norm <- scale(all_pc1_slv$avg) # no sign flip needed

fear_hnd <- data.frame(w=all_pc1_hnd$norm)
fear_gtm <- data.frame(w=all_pc1_gtm$norm)
fear_slv <- data.frame(w=all_pc1_slv$norm)

fear_hnd = as.numeric(unlist(fear_hnd)) 
fear_gtm = as.numeric(unlist(fear_gtm))
fear_slv = as.numeric(unlist(fear_slv))
```

Before going further, let's do some quick sanity checks to make sure things are pointing in the right direction.

```{r}
mean(fear_hnd[lapop.2014.HND$vic1ext==1]) # 0.52
mean(fear_hnd[lapop.2014.HND$vic1ext==2]) # -0.12 Honduran non-victims less fearful
mean(fear_slv[lapop.2014.SLV$vic1ext==1]) # 0.51
mean(fear_slv[lapop.2014.SLV$vic1ext==2]) # -0.12 Salvadoran non-victims less fearful
mean(fear_gtm[lapop.2014.GTM$vic1ext==1]) # 0.58
mean(fear_gtm[lapop.2014.GTM$vic1ext==2]) # -0.12 Guatemalan non-victims less fearful
```

#### HONDURAS #### 

### Binary
```{r}
pvalue = function(index, data, variable) {
  mydata = cbind(data[,variable], index)
  mydata = mydata[mydata[,1] <3000,]
  regression = lm(mydata[,2]~mydata[,1])
  summary(regression)$coefficients[,c(1,4)]
}

pvalue(fear_hnd, lapop.2014.HND, 'ur') #rural populations are less fearful 
pvalue(fear_hnd, lapop.2014.HND, 'q1')
pvalue(fear_hnd, lapop.2014.HND, 'np1')
pvalue(fear_hnd, lapop.2014.HND, 'np2')
pvalue(fear_hnd, lapop.2014.HND, 'prot3') #people who have not participated in a protest are less fearful 
pvalue(fear_hnd, lapop.2014.HND, 'prot7')
pvalue(fear_hnd, lapop.2014.HND, 'vb2') # voters are less fearful
pvalue(fear_hnd, lapop.2014.HND, 'vb10')
pvalue(fear_hnd, lapop.2014.HND, 'wf1')
pvalue(fear_hnd, lapop.2014.HND, 'cct1b')
pvalue(fear_hnd, lapop.2014.HND, 'q10a') #people who do not receive remittances are less fearful 
pvalue(fear_hnd, lapop.2014.HND, 'q14') #people who dont plan to study abroad are less fearful 
pvalue(fear_hnd, lapop.2014.HND, 'sexi') #people who had a female interviewer are less fearful 
```

### Continuous Variables 
```{r}
pvalue(fear_hnd, lapop.2014.HND, 'ico2') #when police patrols less, participants are more fearful 
pvalue(fear_hnd, lapop.2014.HND, 'pole2n') #more dissatisfied with police means more fearful 
pvalue(fear_hnd, lapop.2014.HND, 'aoj12') #the less confident people are in judicial system to punishthe guilty, the more fearful 
pvalue(fear_hnd, lapop.2014.HND, 'mil1') #less fearful if more confident in training of army/military 
pvalue(fear_hnd, lapop.2014.HND, 'mil3')
pvalue(fear_hnd, lapop.2014.HND, 'sd3new2') #less satisfied with schools, more fearful 
pvalue(fear_hnd, lapop.2014.HND, 'infrax')
pvalue(fear_hnd, lapop.2014.HND, 'ing4') #those who think democracy is better are less fearful 
pvalue(fear_hnd, lapop.2014.HND, 'eff1') #those who think leaders are interested in public opinion are less fearful 
pvalue(fear_hnd, lapop.2014.HND, 'eff2')
pvalue(fear_hnd, lapop.2014.HND, 'mil7')
pvalue(fear_hnd, lapop.2014.HND, 'per4') #anxious personalities are more fearful 
pvalue(fear_hnd, lapop.2014.HND, 'per9')
pvalue(fear_hnd, lapop.2014.HND, 'dem2')
pvalue(fear_hnd, lapop.2014.HND, 'vb1')
pvalue(fear_hnd, lapop.2014.HND, 'pol1') #those less interested in politics are less fearful 
pvalue(fear_hnd, lapop.2014.HND, 'vb20')
pvalue(fear_hnd, lapop.2014.HND, 'for6') #people who think China has less influence are less fearful 
pvalue(fear_hnd, lapop.2014.HND, 'for6b')
pvalue(fear_hnd, lapop.2014.HND, 'mil10a')
pvalue(fear_hnd, lapop.2014.HND, 'mil10e')
pvalue(fear_hnd, lapop.2014.HND, 'q5b')
pvalue(fear_hnd, lapop.2014.HND, 'q2y')
pvalue(fear_hnd, lapop.2014.HND, 'q2')
pvalue(fear_hnd, lapop.2014.HND, 'idio2')
pvalue(fear_hnd, lapop.2014.HND, 'soct2')
pvalue(fear_hnd, lapop.2014.HND, 'ed') #those with more yrs. of education are more fearful 
pvalue(fear_hnd, lapop.2014.HND, 'q12c')
pvalue(fear_hnd, lapop.2014.HND, 'q12bn')
pvalue(fear_hnd, lapop.2014.HND, 'q12') #those with more children are less fearful 
```

Now we should look at which of the variables that correlate with our fear index also correlate with each other.

```{r}
bin_bin <- function(data,var1,var2) {
  v1 <- data[data[,var1] < 1000 & data[,var2] < 1000,var1]
  v2 <- data[data[,var1] < 1000 & data[,var2] < 1000,var2]
  ct <- table(v1,v2)
  ft <- fisher.test(ct)
  c(ft$p.value,ft$estimate)
}
hnd_bin <- c('vb2','q10a','q14','ur','sexi','prot3')
for (x in hnd_bin) {
  for (y in hnd_bin[-1:-which(hnd_bin==x)]) {
    res <- bin_bin(lapop.2014.HND,x,y)
    if (res[1] < 0.01) {
      print(paste(x,y,res[1],res[2]))
    }
  }
}
```

So there are negative relationships between:

* `vb2` and `q14`: Voters less likely to plan to leave 
* `ur` and `sexi`: More female interviewers in urban areas

And positive relationships between:

* `q10a` and `q14`: Remittance receivers more likely to plan to leave
* `q10a` and `ur`: Remittance receivers more urban
* `q14` and `ur`: Urban people more likely to leave

Now, correlate binary and continuous variables:

```{r}
bin_cont <- function(data,bvar,cvar) {
  b <- data[data[,bvar] < 1000 & data[,cvar] < 1000,bvar]
  b <- b - min(b) # convert to 0-1 scale
  c <- data[data[,bvar] < 1000 & data[,cvar] < 1000,cvar]  
  tt <- t.test(c[b==0],c[b==1])
  c(tt$p.value,tt$estimate[2]-tt$estimate[1])
}
hnd_cont <- c('ed','for6','pol1','per4','sd3new2','mil1','aoj12','q12','ico2',
              'pole2n','eff1','ing4')
for (x in hnd_bin) {
  for (y in hnd_cont) {
    res <- bin_cont(lapop.2014.HND,x,y)
    if (res[1] < 0.01) {
      print(paste(x,y,res[1],res[2]))
    }
  }
}
```

Negative correlations:

* `vb2` and `q12`: Voters have more children
* `q10a` and `ed`: Educated people receive more remittances
* `q14` and `ed`: Educated people plan to leave
* `q14` and `sd3new2`: People more dissatisfied with schools plan to leave
* `ur` and `ed`: Urban people more educated
* `ur` and `per4`: Urban people have anxious personalities
* `ur` and `aoj12`: Urban people less confident that guilty are punished
* `sexi` and `for6`: People with male interviewers find China more influential

Positive correlations:

* `vb2` and `pol1`: Voters more interested in politics
* `prot3` and `pol1`: Protesters more interested in politics
* `q14` and `pol1`: People interested in politics plan to leave
* `q14` and `q12`: People with fewer children plan to leave
* `ur` and `for6`: Urban people consider China more influential
* `ur` and `q12`: Urban people have fewer children
* `ur` and `ico2`: More National Police patrols in urban neighborhoods
* `sexi` and `aoj12`: People with male interviewers more convinced that guilty are punished

Finally, let's correlate the continuous variables with each other:

```{r}
cont_cont <- function(data,var1,var2) {
  v1 <- data[data[,var1] < 1000 & data[,var2] < 1000,var1]
  v2 <- data[data[,var1] < 1000 & data[,var2] < 1000,var2]
  reg <- lm(v1~v2)
  summary(reg)$coefficients[2,c(1,4)]
}
for (x in hnd_cont) {
  for (y in hnd_cont[-1:-which(hnd_cont==x)]) {
    res <- cont_cont(lapop.2014.HND,x,y)
    if (res[2] < 0.01) {
      print(paste(x,y,res[2],res[1]))
    }
  }
}
```

Negative correlations:

* `ed` and `for6`: Educated people find China more influential
* `ed` and `pol`: Educated people more interested in politics
* `ed` and `q12`: Educated people have fewer children
* `ed` and `ico2`: Educated people get more National Police patrols
* `for6` and `ing4`: People who see Chinese influence prefer democracy
* `pol1` and `mil1`: People interested in politics have a higher opinion of military
* `pol1` and `eff1`: People interested in politics think leaders are interested
* `sd3new2` and `mil1`: People satisfied with schools have a high opinion of military
* `sd3new2` and `eff1`: People satisfied with schools think leaders are interested
* `sd3new2` and `ing4`: People satisfied with schools like democracy
* `mil1` and `aoj12`: People with a high opinion of the military also trust the judiciary
* `mil1` and `ico2`: People with lots of National Police patrols think highly of the military
* `mil1` and `pole2n`: People who love the police also love the military
* `aoj12` and `eff1`: People who think their leaders care also believe the guilty are punished
* `ico2` and `ing4`: People with lots of National Police patrols prefer democracy
* `pole2n` and `eff1`: People who think their leaders care like the police
* `pole2n` and `ing4`: People who like the police like democracy

Positive correlations:

* `ed` and `aoj12`: More educated people don't think the guilty are punished
* `pol1` and `aoj12`: People interested in politics do think the guilty are punished
* `per4` and `pole2n`: Anxious people dislike the police
* `sd3new2` and `aoj12`: People who are satisfied with schools also think the guilty are punished
* `sd3new2` and `ico2`: People who get lots of National Police patrols are satisfied with schools
* `sd3new2` and `pole2n`: People who like the schools like the cops
* `mil1` and `eff1`: People who think their leaders care also think highly of the military
* `mil1` and `ing4`: People who think highly of the military also like democracy
* `aoj12` and `ico2`: People who get lots of National Police patrols think the guilty are punished
* `aoj12` and `pole2n`: People who think the guilty are punished like the police
* `ico2` and `pole2n`: People who get lots of police patrols love the police
* `eff1` and `ing4`: People who think their leaders don't care disapprove of democracy

This seems to point to a coherent set of beliefs -- high levels of trust in the military, police, schools, and political leaders all seem to go together.

Now this can really get interesting -- we're prepared to look at multiple regressions. The goal here is to *invalidate* correlations; we want to rule out as many as we can. 

Let's start with an easy one: I really don't believe that people are afraid of crime because they think China is influential. Perception of Chinese influence (`for6`) correlates with urbanization (`ur`), education (`ed`), preference for democracy (`ing4`), and having a male interviewer (`sexi`).

```{r}
hnd <- lapop.2014.HND
hnd[hnd > 5000] <- NA
hnd$fear_idx <- fear_hnd
summary(lm(fear_hnd ~ hnd$for6 + hnd$ur))
```

So, when we take urbanization into account, China-watchers are no more afraid of crime than other people at the same level of urbanization.

What if we just toss in everything?

```{r}
summary(lm(fear_hnd ~ hnd$ur + hnd$vb2 + hnd$q10a + hnd$q14 + hnd$sexi + 
             hnd$prot3 + hnd$ed + hnd$for6 + hnd$pol1 + hnd$per4 + 
             hnd$sd3new2 + hnd$mil1 + hnd$aoj12 + hnd$q12 + hnd$ico2 + 
             hnd$pole2n +hnd$eff1 + hnd$ing4))
```

With a strict (p < 0.01) criterion, the only significant correlations left are urbanization, male interviewers, anxious personalities, dissatisfaction with schools, infrequent National Police patrols, and dissatisfaction with police.

In hindsight, it probably wasn't necessary to calculate all of those pairwise correlations, because adding all of them tells us what we really needed to know -- which variables really have an impact on our fear index. Of course there's a risk in adding too many variables -- whenever a variable has `NA` values, those responses get omitted from the multiple regression. With too many variables we'll end up throwing away too much data.

```{r}
nrow(na.omit(hnd[,c(hnd_bin,hnd_cont)])) / nrow(hnd)
```

We've still got about 75% of our data included in that multiple regression.

While we're at it, let's look at unordered categorical variables:

```{r}
categ <- function(f,x,categ) {
  # For a data from f with a continuous variable 'x' and an unordered 
  # categorical variable 'categ', test whether the value of x is 
  # significantly higher or lower for each value of categ than for the
  # rest of the population. Do this using a two-sample t-test.
  result <- data.frame(var=character(),val=integer(),pval=double(),
                       mean=double(),othermean=double())
  for(q in unique(na.omit(f[,categ]))) {
    if(sum(f[,categ]==q,na.rm=TRUE) > 1) {
      yes <- f[f[,categ]==q,x]
      no <- f[f[,categ]!=q,x]
      if (sum(!is.na(no)) < 2) return()
      tt <- t.test(yes,no,na.rm=TRUE)
      if (tt$p.value < 0.01) {
        newrow <- data.frame(var=categ,val=q,pval=tt$p.value,
                         mean=tt$estimate[[1]],
                         othermean=tt$estimate[[2]])
        result <- rbind(result,newrow)
      }
    }
  }
  result
}
unordered_vars <- c('a4','vic2','vic2aa','vb3n','vb4new','vb101','vb11',
                    'for1n','for4','for5','q3c','ocup4a','ocup1a','q11n',
                    'etid','leng1')
categ(hnd,'fear_idx','a4') # people who complain about health services are less afraid
categ(hnd,'fear_idx','vic2') # nothing significant
categ(hnd,'fear_idx','vic2aa') # people who were victimized away from home are less afraid than other victims
categ(hnd,'fear_idx','vb3n') # JOH voters less fearful, Nasralla voters more fearful
categ(hnd,'fear_idx','vb4new') # Nothing significant
categ(hnd,'fear_idx','vb101') # Nada
categ(hnd,'fear_idx','vb11') # Partido Anti Corrupcion (Nasralla's) voters more fearful
categ(hnd,'fear_idx','for1n') # nada
categ(hnd,'fear_idx','for4') # nada
categ(hnd,'fear_idx','for5') # nada
categ(hnd,'fear_idx','q3c') # Catholics less fearful, traditional/Native much less fearful
categ(hnd,'fear_idx','ocup4a') # job seekers more fearful
categ(hnd,'fear_idx','ocup1a') # self-employed less fearful, salaried more fearful
categ(hnd,'fear_idx','q11n') # married people less fearful
categ(hnd,'fear_idx','etid') # nothing
categ(hnd,'fear_idx','leng1') # nothing

hnd$a4_22 <- as.numeric(hnd$a4==22)
hnd$vic2aa_4 <- as.numeric(hnd$vic2aa==4)
hnd$vb3n_401 <- as.numeric(hnd$vb3n==401)
hnd$vb3n_404 <- as.numeric(hnd$vb3n==404)
hnd$vb11_407 <- as.numeric(hnd$vb11==407)
hnd$q3c_1 <- as.numeric(hnd$q3c==1)
hnd$q3c_7 <- as.numeric(hnd$q3c==7)
hnd$ocup4a_3 <- as.numeric(hnd$ocup4a==3)
hnd$ocup1a_4 <- as.numeric(hnd$ocup1a==4)
hnd$ocup1a_2 <- as.numeric(hnd$ocup1a==2)
hnd$ocup1a_1 <- as.numeric(hnd$ocup1a==1)
hnd$q11n_2 <- as.numeric(hnd$q11n==2)

cat_var <- c('a4_22','vic2aa_4','vb3n_401','vb3n_404','vb11_407',
             'q3c_1','q3c_7','ocup4a_3','ocup1a_4','ocup1a_2',
             'ocup1a_1','q11n_2')

for (x in cat_var) {
  s <- lm(fear_hnd ~ hnd$ur + hnd$vb2 + hnd$q10a + hnd$q14 + hnd$sexi + 
             hnd$prot3 + hnd$ed + hnd$for6 + hnd$pol1 + hnd$per4 + 
             hnd$sd3new2 + hnd$mil1 + hnd$aoj12 + hnd$q12 + hnd$ico2 + 
             hnd$pole2n +hnd$eff1 + hnd$ing4 + hnd[,x])
  res <- summary(s)$coefficient[nrow(summary(s)$coefficient),c(4,1)]
  if (res[1] < 0.01) {
    print(paste(c(x,res)))
  }
}
```

With a parsimonious p < 0.01, only `vic2aa_4` is significant. At the p < 0.05 level, `ocup4a_3==3` (job seekers) and `vb11==407` (Anti-corruption party (PAC) adherents) also look interesting.

```{r}
summary(lm(fear_hnd ~ hnd$ur + hnd$vb2 + hnd$q10a + hnd$q14 + hnd$sexi + 
             hnd$prot3 + hnd$ed + hnd$for6 + hnd$pol1 + hnd$per4 + 
             hnd$sd3new2 + hnd$mil1 + hnd$aoj12 + hnd$q12 + hnd$ico2 + 
             hnd$pole2n +hnd$eff1 + hnd$ing4 + hnd$vic2aa_4))
```

Because `vic2aa_4` has `NA` values for all respondents who *weren't* victims in the last 12 months, it ruins most of my other correlations. We do see, however, that out of the population of self-identified victims, those with male interviewers are significantly more fearful.

Looking forward to the social media analyses, it seems that frustration with schools and police, urbanization, anxiety, unemployment, and PAC support could be interesting avenues to explore.

### GUATEMALA ####
```{r}
pvalue(fear_gtm, lapop.2014.GTM, 'ur') #rural populations are less fearful 
pvalue(fear_gtm, lapop.2014.GTM, 'q1')
pvalue(fear_gtm, lapop.2014.GTM, 'np1')
pvalue(fear_gtm, lapop.2014.GTM, 'np2')
pvalue(fear_gtm, lapop.2014.GTM, 'prot3')
pvalue(fear_gtm, lapop.2014.GTM, 'vb2')
pvalue(fear_gtm, lapop.2014.GTM, 'vb10')
pvalue(fear_gtm, lapop.2014.GTM, 'wf1') #people who don't receive govt. assistance are less fearful 
pvalue(fear_gtm, lapop.2014.GTM, 'cct1b') #people who don't receive cash transfers are more fearful 
pvalue(fear_gtm, lapop.2014.GTM, 'q10a')
pvalue(fear_gtm, lapop.2014.GTM, 'q14')
pvalue(fear_gtm, lapop.2014.GTM, 'sexi') #participants with a female interviewer were more fearful 
```

### Continuous Variables 
```{r}
pvalue(fear_gtm, lapop.2014.GTM, 'it1') #people who are less trusting are more fearful 
pvalue(fear_gtm, lapop.2014.GTM, 'aoj12')#those with less confidence in judiciary system to punish the guilty are more fearful
pvalue(fear_gtm, lapop.2014.GTM, 'mil4')
pvalue(fear_gtm, lapop.2014.GTM, 'pn4')#those less satisfied with democracy are more fearful 
pvalue(fear_gtm, lapop.2014.GTM, 'www1')#those who use the internet less are less fearful 
pvalue(fear_gtm, lapop.2014.GTM, 'q10e')
pvalue(fear_gtm, lapop.2014.GTM, 'pol1')#those less interested in politics are less fearful 
pvalue(fear_gtm, lapop.2014.GTM, 'for6')#those who think china has less influence are less fearful 
pvalue(fear_gtm, lapop.2014.GTM, 'for6b')#those who think the U.S. has less influence in their country are less fearful
pvalue(fear_gtm, lapop.2014.GTM, 'for7')
pvalue(fear_gtm, lapop.2014.GTM, 'mil10c')
pvalue(fear_gtm, lapop.2014.GTM, 'q5b')
pvalue(fear_gtm, lapop.2014.GTM, 'ocup4a')#

```

#### EL SALVADOR ####
```{r}
pvalue(fear_slv, lapop.2014.SLV, 'ur') #rural populations are more fearful
pvalue(fear_slv, lapop.2014.SLV, 'q1') #females are more fearful 
pvalue(fear_slv, lapop.2014.SLV, 'np1') #those who have not attended municipal meeting are more fearful 
pvalue(fear_slv, lapop.2014.SLV, 'np2') #those who have not requested help from municipal office are more fearful 
pvalue(fear_slv, lapop.2014.SLV, 'prot3') #those who have not attended a protest are more fearful 
pvalue(fear_slv, lapop.2014.SLV, 'vb2')
pvalue(fear_slv, lapop.2014.SLV, 'vb10') #those who do not identify with a political party are more fearful 
pvalue(fear_slv, lapop.2014.SLV, 'wf1') #those who do not receive govt. assistance are less fearful 
pvalue(fear_slv, lapop.2014.SLV, 'cct1b')
pvalue(fear_slv, lapop.2014.SLV, 'q10a') #those who don't receive remittances are more fearful  
pvalue(fear_slv, lapop.2014.SLV, 'q14') #those who do not plan to study abroad are more fearful 
pvalue(fear_slv, lapop.2014.SLV, 'sexi') #those with female interviewers were less fearful 
```

### Continuous Variables 
```{r}
pvalue(fear_slv, lapop.2014.SLV, 'it1') #those that are less trusting are less fearful 
pvalue(fear_slv, lapop.2014.SLV, 'ico2')
pvalue(fear_slv, lapop.2014.SLV, 'aoj12')#those with less confidence in judicial system are less fearful 
pvalue(fear_slv, lapop.2014.SLV, 'n11')#
pvalue(fear_slv, lapop.2014.SLV, 'n15')#
pvalue(fear_slv, lapop.2014.SLV, 'mil3')
pvalue(fear_slv, lapop.2014.SLV, 'infrax')
pvalue(fear_slv, lapop.2014.SLV, 'ing4')
pvalue(fear_slv, lapop.2014.SLV, 'eff1') #
pvalue(fear_slv, lapop.2014.SLV, 'eff2')#
pvalue(fear_slv, lapop.2014.SLV, 'q10g')
pvalue(fear_slv, lapop.2014.SLV, 'vb3n')
pvalue(fear_slv, lapop.2014.SLV, 'vb4new')
pvalue(fear_slv, lapop.2014.SLV, 'vb20')#
pvalue(fear_slv, lapop.2014.SLV, 'for6')#
pvalue(fear_slv, lapop.2014.SLV, 'for6b')#
pvalue(fear_slv, lapop.2014.SLV, 'mil10a')#
pvalue(fear_slv, lapop.2014.SLV, 'mil10e')
pvalue(fear_slv, lapop.2014.SLV, 'q5b')
pvalue(fear_slv, lapop.2014.SLV, 'q10d')

```


