---
title: "binary"
author: "calbano"
date: "November 29, 2015"
output: html_document
---
```{r}
lapop.2014.HND <- read.csv("../HND-2014.csv")
lapop.2014.GTM <- read.csv("../GTM-2014.csv")
lapop.2014.SLV <- read.csv("../SLV-2014.csv")

library(mice)
library(plyr)
library(ggplot2)
library(GGally)
set.seed(12345)
```

```{r}
gtm <- lapop.2014.GTM[,c('vic40', 'vic41', 'vic44', 'vic43',
      'fear10','fear6f','vic1ext','vic1exta',
      'vic1hogar','aoj11','pese1','pese2','aoj17',
      'diso7','diso8','diso10','diso18','diso14',
      'diso16','diso17','vicbar1','vicbar1f',
      'vicbar3','vicbar4','vicbar7')]

hnd <- lapop.2014.HND[,c('vic40', 'vic41', 'vic44', 'vic43', 'vic45',
      'fear10','fear6f','vic1ext','vic1exta',
      'vic1hogar','aoj11','pese1','pese2','aoj17',
      'diso7','diso8','diso10','diso18','diso14',
      'diso16','diso17','vicbar1','vicbar1f',
      'vicbar3','vicbar4','vicbar7')]

slv <- lapop.2014.SLV[,c('vic44',
      'fear10','vic1ext','vic1exta',
      'vic1hogar','aoj11','pese1','pese2','aoj17',
      'diso7','diso8','diso10','diso18','diso14',
      'diso16','diso17','vicbar1','vicbar1f',
      'vicbar3','vicbar4','vicbar7')]

hnd$vic1exta[hnd$vic1exta == 999999] = 0
gtm$vic1exta[gtm$vic1exta == 999999] = 0
slv$vic1exta[slv$vic1exta == 999999] = 0

is.na(hnd[hnd>30]) = TRUE
is.na(gtm[gtm>30]) = TRUE
is.na(slv[slv>30]) = TRUE

complete.hnd = na.omit(hnd)
complete.gtm = na.omit(gtm)
complete.slv = na.omit(slv)

pr_hnd = prcomp(complete.hnd, center = TRUE, scale = FALSE)
pr_gtm = prcomp(complete.gtm, center = TRUE, scale = FALSE)
pr_slv = prcomp(complete.slv, center = TRUE, scale = FALSE)

my_imp_hnd <- mice(hnd, printFlag = F)
my_imp_gtm <- mice(gtm, printFlag = F)
my_imp_slv <- mice(slv, printFlag = F)

pr.hnd <- lapply(1:5,function(x) prcomp(complete(my_imp_hnd,x),scale=FALSE,center=TRUE))
pr.gtm <- lapply(1:5,function(x) prcomp(complete(my_imp_gtm,x),scale=FALSE,center=TRUE))
pr.slv <- lapply(1:5,function(x) prcomp(complete(my_imp_slv,x),scale=FALSE,center=TRUE))

all_pc1_hnd <- data.frame(llply(1:5, function(i) pr.hnd[[i]]$x[,1]))
all_pc1_gtm <- data.frame(llply(1:5, function(i) pr.gtm[[i]]$x[,1]))
all_pc1_slv <- data.frame(llply(1:5, function(i) pr.slv[[i]]$x[,1]))

all_pc1_hnd$avg <- rowMeans(all_pc1_hnd)
all_pc1_gtm$avg <- rowMeans(all_pc1_gtm)
all_pc1_slv$avg <- rowMeans(all_pc1_slv)

all_pc1_hnd$norm <- scale(-all_pc1_hnd$avg) # sign flip -- now high values mean more fear
all_pc1_gtm$norm <- scale(-all_pc1_gtm$avg) # sign flip -- now high values mean more fear
all_pc1_slv$norm <- scale(all_pc1_slv$avg) # no sign flip needed

fear_hnd <- data.frame(w=all_pc1_hnd$norm)
fear_gtm <- data.frame(w=all_pc1_gtm$norm)
fear_slv <- data.frame(w=all_pc1_slv$norm)

fear_hnd = as.numeric(unlist(fear_hnd)) 
fear_gtm = as.numeric(unlist(fear_gtm))
fear_slv = as.numeric(unlist(fear_slv))
```

Before going further, let's do some quick sanity checks to make sure things are pointing in the right direction.

```{r}
mean(fear_hnd[lapop.2014.HND$vic1ext==1]) # 0.52
mean(fear_hnd[lapop.2014.HND$vic1ext==2]) # -0.12 Honduran non-victims less fearful
mean(fear_slv[lapop.2014.SLV$vic1ext==1]) # 0.51
mean(fear_slv[lapop.2014.SLV$vic1ext==2]) # -0.12 Salvadoran non-victims less fearful
mean(fear_gtm[lapop.2014.GTM$vic1ext==1]) # 0.58
mean(fear_gtm[lapop.2014.GTM$vic1ext==2]) # -0.12 Guatemalan non-victims less fearful
```

#### HONDURAS #### 

### Binary
```{r}
pvalue = function(index, data, variable) {
  mydata = cbind(data[,variable], index)
  mydata = mydata[mydata[,1] <3000,]
  regression = lm(mydata[,2]~mydata[,1])
  summary(regression)$coefficients[,c(1,4)]
}

pvalue(fear_hnd, lapop.2014.HND, 'ur') #rural populations are less fearful 
pvalue(fear_hnd, lapop.2014.HND, 'q1')
pvalue(fear_hnd, lapop.2014.HND, 'np1')
pvalue(fear_hnd, lapop.2014.HND, 'np2')
pvalue(fear_hnd, lapop.2014.HND, 'prot3') #people who have not participated in a protest are less fearful 
pvalue(fear_hnd, lapop.2014.HND, 'prot7')
pvalue(fear_hnd, lapop.2014.HND, 'vb2') # voters are less fearful
pvalue(fear_hnd, lapop.2014.HND, 'vb10')
pvalue(fear_hnd, lapop.2014.HND, 'wf1')
pvalue(fear_hnd, lapop.2014.HND, 'cct1b')
pvalue(fear_hnd, lapop.2014.HND, 'q10a') #people who do not receive remittances are less fearful 
pvalue(fear_hnd, lapop.2014.HND, 'q14') #people who dont plan to study abroad are less fearful 
pvalue(fear_hnd, lapop.2014.HND, 'sexi') #people who had a female interviewer are less fearful 
```

### Continuous Variables 
```{r}
pvalue(fear_hnd, lapop.2014.HND, 'ico2') #when police patrols less, participants are more fearful 
pvalue(fear_hnd, lapop.2014.HND, 'pole2n') #more dissatisfied with police means more fearful 
pvalue(fear_hnd, lapop.2014.HND, 'aoj12') #the less confident people are in judicial system to punishthe guilty, the more fearful 
pvalue(fear_hnd, lapop.2014.HND, 'mil1') #less fearful if more confident in training of army/military 
pvalue(fear_hnd, lapop.2014.HND, 'mil3')
pvalue(fear_hnd, lapop.2014.HND, 'sd3new2') #less satisfied with schools, more fearful 
pvalue(fear_hnd, lapop.2014.HND, 'infrax')
pvalue(fear_hnd, lapop.2014.HND, 'ing4') #those who think democracy is better are less fearful 
pvalue(fear_hnd, lapop.2014.HND, 'eff1') #those who think leaders are interested in public opinion are less fearful 
pvalue(fear_hnd, lapop.2014.HND, 'eff2')
pvalue(fear_hnd, lapop.2014.HND, 'mil7')
pvalue(fear_hnd, lapop.2014.HND, 'per4') #anxious personalities are more fearful 
pvalue(fear_hnd, lapop.2014.HND, 'per9')
pvalue(fear_hnd, lapop.2014.HND, 'dem2')
pvalue(fear_hnd, lapop.2014.HND, 'vb1')
pvalue(fear_hnd, lapop.2014.HND, 'pol1') #those less interested in politics are less fearful 
pvalue(fear_hnd, lapop.2014.HND, 'vb20')
pvalue(fear_hnd, lapop.2014.HND, 'for6') #people who think China has less influence are less fearful 
pvalue(fear_hnd, lapop.2014.HND, 'for6b')
pvalue(fear_hnd, lapop.2014.HND, 'mil10a')
pvalue(fear_hnd, lapop.2014.HND, 'mil10e')
pvalue(fear_hnd, lapop.2014.HND, 'q5b')
pvalue(fear_hnd, lapop.2014.HND, 'q2y')
pvalue(fear_hnd, lapop.2014.HND, 'q2')
pvalue(fear_hnd, lapop.2014.HND, 'idio2')
pvalue(fear_hnd, lapop.2014.HND, 'soct2')
pvalue(fear_hnd, lapop.2014.HND, 'ed') #those with more yrs. of education are more fearful 
pvalue(fear_hnd, lapop.2014.HND, 'q12c')
pvalue(fear_hnd, lapop.2014.HND, 'q12bn')
pvalue(fear_hnd, lapop.2014.HND, 'q12') #those with more children are less fearful 
```

Now we should look at which of the variables that correlate with our fear index also correlate with each other.

```{r}
bin_bin <- function(data,var1,var2) {
  v1 <- data[data[,var1] < 1000 & data[,var2] < 1000,var1]
  v2 <- data[data[,var1] < 1000 & data[,var2] < 1000,var2]
  ct <- table(v1,v2)
  ft <- fisher.test(ct)
  c(ft$p.value,ft$estimate)
}
hnd_bin <- c('vb2','q10a','q14','ur','sexi','prot3')
for (x in hnd_bin) {
  for (y in hnd_bin[-1:-which(hnd_bin==x)]) {
    res <- bin_bin(lapop.2014.HND,x,y)
    if (res[1] < 0.05) {
      print(paste(x,y,res[1],res[2]))
    }
  }
}
```

So there are negative relationships between:

* `vb2` and `q14`: Voters less likely to plan to leave 
* `ur` and `sexi`: More female interviewers in urban areas

And positive relationships between:

* `vb2` and `prot3`: Voters more likely to protest
* `q10a` and `q14`: Remittance receivers more likely to plan to leave
* `q10a` and `ur`: Remittance receivers more urban
* `q10a` and `prot3`: Remittance receivers more likely to protest
* `q14` and `ur`: Urban people more likely to leave

### GUATEMALA ####
```{r}
pvalue(fear_gtm, lapop.2014.GTM, 'ur') #rural populations are less fearful 
pvalue(fear_gtm, lapop.2014.GTM, 'q1')
pvalue(fear_gtm, lapop.2014.GTM, 'np1')
pvalue(fear_gtm, lapop.2014.GTM, 'np2')
pvalue(fear_gtm, lapop.2014.GTM, 'prot3')
pvalue(fear_gtm, lapop.2014.GTM, 'vb2')
pvalue(fear_gtm, lapop.2014.GTM, 'vb10')
pvalue(fear_gtm, lapop.2014.GTM, 'wf1') #people who don't receive govt. assistance are less fearful 
pvalue(fear_gtm, lapop.2014.GTM, 'cct1b') #people who don't receive cash transfers are more fearful 
pvalue(fear_gtm, lapop.2014.GTM, 'q10a')
pvalue(fear_gtm, lapop.2014.GTM, 'q14')
pvalue(fear_gtm, lapop.2014.GTM, 'sexi') #participants with a female interviewer were more fearful 
```

### Continuous Variables 
```{r}
pvalue(fear_gtm, lapop.2014.GTM, 'it1') #people who are less trusting are more fearful 
pvalue(fear_gtm, lapop.2014.GTM, 'aoj12')#those with less confidence in judiciary system to punish the guilty are more fearful
pvalue(fear_gtm, lapop.2014.GTM, 'mil4')
pvalue(fear_gtm, lapop.2014.GTM, 'pn4')#those less satisfied with democracy are more fearful 
pvalue(fear_gtm, lapop.2014.GTM, 'www1')#those who use the internet less are less fearful 
pvalue(fear_gtm, lapop.2014.GTM, 'q10e')
pvalue(fear_gtm, lapop.2014.GTM, 'pol1')#those less interested in politics are less fearful 
pvalue(fear_gtm, lapop.2014.GTM, 'for6')#those who think china has less influence are less fearful 
pvalue(fear_gtm, lapop.2014.GTM, 'for6b')#those who think the U.S. has less influence in their country are less fearful
pvalue(fear_gtm, lapop.2014.GTM, 'for7')
pvalue(fear_gtm, lapop.2014.GTM, 'mil10c')
pvalue(fear_gtm, lapop.2014.GTM, 'q5b')
pvalue(fear_gtm, lapop.2014.GTM, 'ocup4a')#

```

#### EL SALVADOR ####
```{r}
pvalue(fear_slv, lapop.2014.SLV, 'ur') #rural populations are more fearful
pvalue(fear_slv, lapop.2014.SLV, 'q1') #females are more fearful 
pvalue(fear_slv, lapop.2014.SLV, 'np1') #those who have not attended municipal meeting are more fearful 
pvalue(fear_slv, lapop.2014.SLV, 'np2') #those who have not requested help from municipal office are more fearful 
pvalue(fear_slv, lapop.2014.SLV, 'prot3') #those who have not attended a protest are more fearful 
pvalue(fear_slv, lapop.2014.SLV, 'vb2')
pvalue(fear_slv, lapop.2014.SLV, 'vb10') #those who do not identify with a political party are more fearful 
pvalue(fear_slv, lapop.2014.SLV, 'wf1') #those who do not receive govt. assistance are less fearful 
pvalue(fear_slv, lapop.2014.SLV, 'cct1b')
pvalue(fear_slv, lapop.2014.SLV, 'q10a') #those who don't receive remittances are more fearful  
pvalue(fear_slv, lapop.2014.SLV, 'q14') #those who do not plan to study abroad are more fearful 
pvalue(fear_slv, lapop.2014.SLV, 'sexi') #those with female interviewers were less fearful 
```

### Continuous Variables 
```{r}
pvalue(fear_slv, lapop.2014.SLV, 'it1') #those that are less trusting are less fearful 
pvalue(fear_slv, lapop.2014.SLV, 'ico2')
pvalue(fear_slv, lapop.2014.SLV, 'aoj12')#those with less confidence in judicial system are less fearful 
pvalue(fear_slv, lapop.2014.SLV, 'n11')#
pvalue(fear_slv, lapop.2014.SLV, 'n15')#
pvalue(fear_slv, lapop.2014.SLV, 'mil3')
pvalue(fear_slv, lapop.2014.SLV, 'infrax')
pvalue(fear_slv, lapop.2014.SLV, 'ing4')
pvalue(fear_slv, lapop.2014.SLV, 'eff1') #
pvalue(fear_slv, lapop.2014.SLV, 'eff2')#
pvalue(fear_slv, lapop.2014.SLV, 'q10g')
pvalue(fear_slv, lapop.2014.SLV, 'vb3n')
pvalue(fear_slv, lapop.2014.SLV, 'vb4new')
pvalue(fear_slv, lapop.2014.SLV, 'vb20')#
pvalue(fear_slv, lapop.2014.SLV, 'for6')#
pvalue(fear_slv, lapop.2014.SLV, 'for6b')#
pvalue(fear_slv, lapop.2014.SLV, 'mil10a')#
pvalue(fear_slv, lapop.2014.SLV, 'mil10e')
pvalue(fear_slv, lapop.2014.SLV, 'q5b')
pvalue(fear_slv, lapop.2014.SLV, 'q10d')

```


