The following is a proposed index for Political Violence, a phenomenon that has affected all three nations of the Northern Triangle, in varying degrees. While the studies done by LAPOP, do not provide any indicators of political violence for El Salvador and Honduras, there are several elements to take into consideration, so as to better comprehend the fabric of institutionalized violence in Guatemala.
All of the variables chosen reflect the responses of whether or not the people, being interrogated, have witnessed any acts of violence related to political identification or electoral cycles.
* `pv1`: Witnessed an act of violence linked to the 2011 campaign (1=Yes, 2=No)
* `pv2a`: Was attacked because of his/her political position (1=Yes, 2=No)
* `pv2b`: Received threats from a candidate to vote for him (1=Yes, 2=No)
* `pv2c`: Received threats from a party to support it during campaign (1=Yes, 2=No)
* `pv2d`: Violent groups linked to political parties in the neighborhood (1=Yes, 2=No)
* `pv2e`: A candidate in the community was attacked or hurt during the campaign (1=Yes, 2=No)
* `pv2f`: A candidate in the community was assassinated during the campaign (1=Yes, 2=No)
* `pv2g`: There were political lynchings in the community (1=Yes, 2=No)
* `pv2h`: There were political assassinations in the community (1=Yes, 2=No)
* `pv2i`: Party leaders threatened voters in the community (1=Yes, 2=No)
* `pv2j`: Party leaders caused property damages in the community (1=Yes, 2=No)
* `pv2k`: There was political repression in a political event in the community (1=Yes, 2=No)


```{r}
library(ggplot2)
#install.packages("mice") # if you don't have it already
library(mice)
library(plyr) # nice for re-formatting data
#install.packages("GGally") # if you don't have it already
library(GGally) # for plot matrices
set.seed(12345) # makes imputation reproducible
```

```{r}
lapop.2014.GTM <- read.csv("../GTM-2014.csv")
my_data <- lapop.2014.GTM[,c('pv1','pv2a','pv2b','pv2c','pv2d','pv2e','pv2f','pv2g','pv2h','pv2i','pv2j','pv2k')]
```

```{r}
is.na(my_data[my_data>16]) <- TRUE
nrow(na.omit(my_data))
```
With the missing data, we are left with 1,395 rows

```{r}
sum(is.na(my_data)) / (nrow(my_data)*ncol(my_data))
```
About 1.1% of missing data

```{r}
my_complete <- na.omit(my_data)
pr_complete <- prcomp(my_complete,center=TRUE,scale=FALSE)
plot(pr_complete)
summary(pr_complete)
```
Without imputation, 37.8% of variance is concentrated on the first principal component.

```{r}
my_imp <- mice(my_data,printFlag = F) 
my_imp
```
Thus far, not many variables seem to be needing any imputation. `pv1` (witnessed an act of violence linked to the 2011 campaign) is the one variable that is missing the most amount of cells per column. Judging from the options, this may be attributed to lack of awareness or unwillingness to respond.

```{r}
pr <- lapply(1:5,function(x) prcomp(complete(my_imp,x),scale=FALSE,center=TRUE))
plot(pr[[1]])
summary(pr[[1]])
```
With imputation, 36.6% of variance is concentrated the first principal component - so, not much change.

```{r}
pc1 <- pr[[1]]$x[,1]
pc2 <- pr[[1]]$x[,2]
qplot(pc1,pc2)
```

```{r}
all_pc1 <- data.frame(llply(1:5, function(i) pr[[i]]$x[,1]))
names(all_pc1) <- c('imp1','imp2','imp3','imp4','imp5')
ggpairs(all_pc1) + theme_classic()
```
Overall, we have got some strong correlations. 

```{r}
all_pc1$avg <- rowMeans(all_pc1)
quantile(all_pc1$avg)
colMeans(my_data[all_pc1$avg < quantile(all_pc1$avg)[2],],na.rm=TRUE)
colMeans(my_data[all_pc1$avg > quantile(all_pc1$avg)[4],],na.rm=TRUE)
```
So, now, we are entering territory were I may be inclined to pressing the "help" button. I am noticing that the lower 25% quartile has a value of NaN ("not a number") for every variable. However, looking at the higher 75% quartile, I am inclined to say that the majority of people responded "no" on each question. Considering that this is an index of witnessing of political violence, this may not be too alarming. A priori, the cases of political violence may be reserved to particular areas. 

```{r}
all_pc1$norm <- -scale(all_pc1$avg) 
predict_data <- data.frame(pv1=c(1,2),pv2a=c(1,2),pv2b=c(1,2),pv2c=c(1,2),
                           pv2d=c(1,2),pv2e=c(1,2),pv2f=c(1,2),
                           pv2g=c(1,2),pv2h=c(1,2),pv2i=c(1,2),pv2j=c(1,2),
                           pv2k=c(1,2))
minmax <- rowMeans(sapply(1:5, function(i) predict(pr[[i]],predict_data)[,1]))
minmax <- -(minmax - mean(all_pc1$avg)) / sd(all_pc1$avg) 

predict_data2 <- data.frame(diag(12)+1)
names(predict_data2) <- names(my_data)
scores <- rowMeans(sapply(1:5, function(i) predict(pr[[i]],predict_data2)[,1]))
scores <- -(scores-mean(all_pc1$avg)) / sd(all_pc1$avg) 
diff <- data.frame(d=(scores - minmax[1]),n=names(predict_data))
ggplot(diff,aes(x=n,y=d)) +
  geom_bar(stat='identity',fill='skyblue') +
  coord_flip() +
  theme_classic() +
  theme(text=element_text(size=20),
        axis.title.y=element_blank(),
        axis.title.x=element_blank())
```
I am noticing that `pv1`is by far the indicator with the strongest influence in the index. Such is the opposite for `pv2f`(Candidate in the community assassinated during the campaign) and `pv2h` (Political assassinations in the community)

```{r}
log_data <- my_data[,c('pv1','pv2a','pv2b','pv2c','pv2d','pv2e','pv2f','pv2g','pv2h','pv2i','pv2j','pv2k')]
log_data <- log_data - 1 #convert to 0-1 scale
p_vals <- sapply(log_data, function(x) coef(summary(glm(x ~ all_pc1$norm,family=binomial(logit))))[2,4])
p_vals
```
Coincidentally, `pv2f`and `pv2h`have also got the lowest p-values of the group.

```{r}
pv2f <- data.frame(q=my_data$pv2f,w=all_pc1$norm)
pv2f <- na.omit(pv2f)
my_lm <- lm(data=pv2f, q ~ w)
summary(my_lm)
ggplot(pv2f,aes(x=w,y=q)) +
  geom_jitter(color='tomato3',size=5,alpha=0.2) +
  theme_classic() +
  geom_smooth(method='lm',size=2,color='royalblue') +
  annotate('text',label=paste("R-sq:",round(summary(my_lm)$r.squared,2)),  
           size=10,x=-0.5,y=2.5,hjust=0,vjust=0,color='royalblue') +
  ylab('pv2f') +
  xlab('Composite index') +
  theme(text=element_text(size=20))
```

```{r}
pv2h <- data.frame(q=my_data$pv2h,w=all_pc1$norm)
pv2h <- na.omit(pv2h)
my_lm <- lm(data=pv2h, q ~ w)
summary(my_lm)
ggplot(pv2h,aes(x=w,y=q)) +
  geom_jitter(color='tomato3',size=5,alpha=0.2) +
  theme_classic() +
  geom_smooth(method='lm',size=2,color='royalblue') +
  annotate('text',label=paste("R-sq:",round(summary(my_lm)$r.squared,2)),  
           size=10,x=-0.5,y=2.5,hjust=0,vjust=0,color='royalblue') +
  ylab('pv2h') +
  xlab('Composite index') +
  theme(text=element_text(size=20))
``` 
Very low r-squared values for both variables.
