---
title: "Trust_Index.rmd"
author: "Jorge L. Perez"
date: "31 October 2015"
output: html_document
---

The following is a proposal for a Trust in Government Index, for which the following variables were considered to be the most relevant in determining the citizens' trust in all civic, public, political and governmental structures:

* `b1`: Belief in judicial system delivering just sentences (1=None at all, 7=High)
* `b2`: Respect for political institutions (1=None at all, 7=High)
* `b3`: Human rights are protected by the political system (1=None at all, 7=High)
* `b4`: Sense of pride in national political system (1=None at all, 7=High)
* `b6`: Support for national political system (1=None at all, 7=High)
* `b10a`: Trust in judicial system (1=None at all, 7=High)
* `b11`: Trust in Supreme Electoral Tribunal (1=None at all, 7=High)
* `b12`: Trust in Armed Forces (1=None at all, 7=High)
* `b13`: Trust in Congress / Parliament (1=None at all, 7=High)
* `b18`: Trust in law enforcement / police force (1=None at all, 7=High)
* `b21`: Trust in political parties (1=None at all, 7=High)
* `b21a`: Trust in President (1=None at all, 7=High)
* `b32`: Trust in municipality (1=None at all, 7=High) 
* `b37`: Trust in media (1=None at all, 7=High)
* `b47a`: Trust in national electoral system (1=None at all, 7=High)
* `b14`: Trust in Central Government (1=None at all, 7=High)
* `b15`: Trust in Ministry of Public Affairs (1=None at all, 7=High)
*`b19`: Trust in Superior Tribunal of Accounts (1=None at all, 7=High)
`b46`: Trust in National Council Against Corruption (1=None at all, 7=High) 
`honb51`: Trust in Human Rights Commission (1=None at all, 7=High)
`venb11`: Belief that votes are kept secret (1=None at all, 7=High)
`epp1`: Political parties represent voters (1=None at all, 7=High)
`epp3`: Political parties listen to individuals (1=None at all, 7=High)
`n9`: Current administration's efforts in fighting corruption (1=None at all, 7=High)
`n11`: Current administration's efforts in improving security (1=None at all, 7=High) 
`n15`: Current administration is handling the economy well (1=None at all, 7=High)
`b3milx`: The Armed Forces respect human rights (1=None at all, 7=High) 

Variables measuring the levels of trust in religious institutions, such as the Catholic and the Protestant Churches, were left out of this proposed index, since they do not partake in the structures of government and political and civil affairs - regardless, of their influence in the citizens' daily lives. 

The `set.seed(12345)` command was not included in this index or the wealth index, but it's a good idea. This specifies the starting point for the random number generator used in multiple imputation and ensures that the results of your analysis will be identical every time.

```{r}
library(ggplot2)
#install.packages("mice") # if you don't have it already
library(mice)
library(plyr) # nice for re-formatting data
#install.packages("GGally") # if you don't have it already
library(GGally) # for plot matrices
set.seed(12345) # makes imputation reproducible
```


```{r}
lapop.2014.HND <- read.csv('../2014-HND.csv')
my_data <- lapop.2014.HND[,c('b1','b2','b3','b4','b6','b10a','b11','b12','b13','b18','b21','b21a','b32', 'b37','b47a','b14','b15','b19','b46','honb51','venb11','venhonvb10','epp1','epp3','n9','n11','n15', 'b3milx')]
```


```{r}
is.na(my_data[my_data>16]) <- TRUE
nrow(na.omit(my_data))
```

```{r}
sum(is.na(my_data)) / (nrow(my_data)*ncol(my_data))
```

```{r}
my_complete <- na.omit(my_data)
pr_complete <- prcomp(my_complete,center=TRUE,scale=FALSE)
plot(pr_complete)
summary(pr_complete)
```

If we don't do any imputation, 39.6% of variance is contained in the first principal component; the others all appear to be much lower.

```{r}
my_imp <- mice(my_data,printFlag = F) 
my_imp
```

None of the variables really stand out as needing a lot of imputation -- `b19` (Trust in the Superior Court of Accounts) is a little higher than others, possibly because some people don't know enough about it to know whether to trust it or not.

```{r}
pr <- lapply(1:5,function(x) prcomp(complete(my_imp,x),scale=FALSE,center=TRUE))
plot(pr[[1]])
summary(pr[[1]])

```

After imputation, things don't change much -- 38.1% of variance is in the first PC.

```{r}
pc1 <- pr[[1]]$x[,1]
pc2 <- pr[[1]]$x[,2]
qplot(pc1,pc2)

```

First indication that things seem to be going well... (Extra points if you can say what about this plot makes me feel that way.)

```{r}
all_pc1 <- data.frame(llply(1:5, function(i) pr[[i]]$x[,1]))
names(all_pc1) <- c('imp1','imp2','imp3','imp4','imp5')
ggpairs(all_pc1) + theme_classic()

```

Beautiful distributions, high correlations (probably because not much imputation was required).

```{r}
all_pc1$avg <- rowMeans(all_pc1)
quantile(all_pc1$avg)
colMeans(my_data[all_pc1$avg < quantile(all_pc1$avg)[2],],na.rm=TRUE)
colMeans(my_data[all_pc1$avg > quantile(all_pc1$avg)[4],],na.rm=TRUE)
```

I see changes, some of them quite large, but without an explanation (at the top of the document, ideally) of which variables we're looking at, it's hard  to know what these changes mean and whether they make intuitive sense.

Without interpreting changes, it's hard to know whether a high value of this index corresponds to high or low trust in government.

```{r}
all_pc1$norm <- scale(all_pc1$avg)
```

```{r}
predict_data <- data.frame(b1=c(1,7),b2=c(1,7),b3=c(1,7),b4=c(1,7),
                           b6=c(1,7),b10a=c(1,7),b11=c(1,7),b12=c(1,7),
                           b13=c(1,7),b18=c(1,7),b21=c(1,7),b21a=c(1,7),
                           b32=c(1,7),b37=c(1,7),b47a=c(1,7),b14=c(1,7),
                           b15=c(1,7), b19=c(1,7), b46=c(1,7), honb51=c(1,7), venb11=c(1,7), venhonvb10=c(1,7), epp1=c(1,7), epp3=c(1,7), n9=c(1,7), n11=c(1,7), n15=c(1,7), b3milx=c(1,7))
minmax <- rowMeans(sapply(1:5, function(i) predict(pr[[i]],predict_data)[,1]))
minmax <- (minmax-mean(all_pc1$avg)) / sd(all_pc1$avg)

predict_data2 <- data.frame(diag(28)+1)
names(predict_data2) <- names(my_data)
scores <- rowMeans(sapply(1:5, function(i) predict(pr[[i]],predict_data2)[,1]))
scores <- (scores-mean(all_pc1$avg)) / sd(all_pc1$avg)
diff <- data.frame(d=(scores - minmax[1]),n=names(predict_data))
ggplot(diff,aes(x=n,y=d)) +
  geom_bar(stat='identity',fill='skyblue') +
  coord_flip() +
  theme_classic() +
  theme(text=element_text(size=20),
        axis.title.y=element_blank(),
        axis.title.x=element_blank())
```

There are a few potential issues here:
- The first value in `minmax` (ostensibly the minimum) is higher than the second (maximum). This could be an indication that `predict_data` wasn't set up correctly, or that high values of your index correspond to low trust, rather than high. The first row in `predict_data` should be the answers that a person with the minimum possible level of trust in government would have given; the second should be the answers showing a maximum level of trust.
- If you look at `quantile(all_pc1$norm)`, you'll see that the minimum value is slightly higher than the lower value in `minmax`, and the maximum slightly lower than the higher value in `minmax`. This is good.
- When you make the bar plot, you'll notice that all of the values are negative. This is because `predict_data2` is full of 1's and 2's -- that means you're starting with a baseline of no trust (1) and moving up to a tiny bit of trust (2). The resulting scores (stored in the variable `scores`) are all close to the maximum value of your index (which is stored in `minmax[1]`). Since you're subtracting the maximum from these perturbed scores (in line 110), everything comes out negative, which is sort of funny. It also tells us that higher values of the index mean less trust, which isn't very intuitive.

I'd recommend replacing the code you have above with the following:

```{r}
all_pc1$norm <- -scale(all_pc1$avg) # reversed; now low scores mean low trust
predict_data <- data.frame(b1=c(1,7),b2=c(1,7),b3=c(1,7),b4=c(1,7),
                           b6=c(1,7),b10a=c(1,7),b11=c(1,7),b12=c(1,7),
                           b13=c(1,7),b18=c(1,7),b21=c(1,7),b21a=c(1,7),
                           b32=c(1,7),b37=c(1,7),b47a=c(1,7),b14=c(1,7),
                           b15=c(1,7), b19=c(1,7), b46=c(1,7), honb51=c(1,7), 
                           venb11=c(1,7), venhonvb10=c(1,7), epp1=c(1,7), 
                           epp3=c(1,7), n9=c(1,7), n11=c(1,7), n15=c(1,7), 
                           b3milx=c(1,7))
minmax <- rowMeans(sapply(1:5, function(i) predict(pr[[i]],predict_data)[,1]))
minmax <- -(minmax - mean(all_pc1$avg)) / sd(all_pc1$avg) # make sure to reverse here also

predict_data2 <- data.frame(diag(28)+1)
names(predict_data2) <- names(my_data)
scores <- rowMeans(sapply(1:5, function(i) predict(pr[[i]],predict_data2)[,1]))
scores <- -(scores-mean(all_pc1$avg)) / sd(all_pc1$avg) # make sure to reverse here as well
diff <- data.frame(d=(scores - minmax[1]),n=names(predict_data))
ggplot(diff,aes(x=n,y=d)) +
  geom_bar(stat='identity',fill='skyblue') +
  coord_flip() +
  theme_classic() +
  theme(text=element_text(size=20),
        axis.title.y=element_blank(),
        axis.title.x=element_blank())
```

Now everything seems to be pointing the right direction.

```{r}
lmp <- function (modelobject) {
    if (class(modelobject) != "lm") stop("Not an object of class 'lm' ")
    f <- summary(modelobject)$fstatistic
    p <- pf(f[1],f[2],f[3],lower.tail=F)
    attributes(p) <- NULL
    return(p)
}
sapply(c('b1','b2','b3', 'b4','b6','b10a','b11','b12','b13',
                             'b18','b21','b21a','b32', 'b37','b47a','b14','b15','b19','b46', 'honb51', 'venb11', 'venhonvb10', 'epp1', 'epp3', 'n9', 'n11', 'n15', 'b3milx'),
       function(x) lmp(lm(all_pc1$norm ~ my_data[,x])))

```

Beautiful!

```{r}

b37 <- data.frame(q=my_data$b37,w=all_pc1$norm)
b37 <- na.omit(b37)
my_lm <- lm(data=b37, q ~ w)
summary(my_lm)
ggplot(b37,aes(x=w,y=q)) +
  geom_jitter(color='tomato3',size=5,alpha=0.2) +
  theme_classic() +
  geom_smooth(method='lm',size=2,color='royalblue') +
  annotate('text',label=paste("R-sq:",round(summary(my_lm)$r.squared,2)),  
           size=10,x=-0.5,y=2.5,hjust=0,vjust=0,color='royalblue') +
  ylab('b37') +
  xlab('Composite index') +
  theme(text=element_text(size=20))

```

So... which variables did you omit here and why?
`names(my_data)[!(names(my_data) %in% names(my_data2))]` tells me you left out b1, b2, b6, and b37 -- these had less influence than the others, but not by very much. I'd suggest leaving them in

```{r}
my_data2 <- lapop.2014.HND[,c('b3', 'b4','b10a','b11','b12','b13', 'b18','b21','b21a','b32', 'b47a','b14','b15','b19','b46', 'honb51', 'venb11', 'venhonvb10', 'epp1', 'epp3', 'n9', 'n11', 'n15', 'b3milx')]
```

```{r}
is.na(my_data2[my_data2>16]) <- TRUE
my_complete2 <- na.omit(my_data2)
pr_complete2 <- prcomp(my_complete2,center=TRUE,scale=FALSE)
plot(pr_complete2)
summary(pr_complete2)
```

```{r}
my_imp2 <- mice(my_data2,printFlag = F) 
pr2 <- lapply(1:5,function(x) prcomp(complete(my_imp2,x),scale=FALSE,center=TRUE))
plot(pr2[[1]]) 
summary(pr2[[1]])

```

`critics_tol` is probably not a very good name for what you're doing here.

```{r}
critics_tol <- data.frame(llply(1:5, function(i) pr2[[i]]$x[,1]))
names(critics_tol) <- c('imp1','imp2','imp3','imp4','imp5')
ggpairs(critics_tol) + theme_classic()
```

```{r}
critics_tol$avg <- rowMeans(critics_tol)
critics_tol$norm <- scale(critics_tol$avg)
colMeans(my_data2[critics_tol$norm < quantile(critics_tol$norm)[2],],na.rm=TRUE)
colMeans(my_data2[critics_tol$norm > quantile(critics_tol$norm)[4],],na.rm=TRUE)
```

```{r}
sapply(c('b3','b4','b10a','b11','b12','b13','b18','b21','b21a','b32','b47a','b14','b15','b19','b46', 'honb51', 'venb11', 'venhonvb10', 'epp1', 'epp3', 'n9', 'n11', 'n15', 'b3milx'),
 function(x) lmp(lm(critics_tol$norm ~ my_data[,x])))

```

```{r}
venhonvb10 <- data.frame(q=my_data$venhonvb10,w=critics_tol$norm)
venhonvb10 <- na.omit(venhonvb10)
my_lm <- lm(data=venhonvb10, q ~ w)
summary(my_lm)
ggplot(venhonvb10,aes(x=w,y=q)) +
  geom_jitter(color='tomato3',size=5,alpha=0.2) +
  theme_classic() +
  geom_smooth(method='lm',size=2,color='royalblue') +
  annotate('text',label=paste("R-sq:",round(summary(my_lm)$r.squared,2)),  
           size=10,x=-0.5,y=2.5,hjust=0,vjust=0,color='royalblue') +
  ylab('venhonvb10') +
  xlab('Composite index') +
  theme(text=element_text(size=20))


```























































































































































































