---
title: "Verify correlations"
author: "Craig"
date: "12 December 2015"
output: html_document
---

Rather than re-calculating all of the indices here, we've got the code to do this stashed in a separate R file.

``` {r}
source('make_indices.R')
library(plyr)
library(MASS)
```

To systematically explore all possible correlations with our fear index, we first need to see which haven't been accounted for in one of our other composite indices.

```{r}
n_fear_more <- c(fear_common,'vic40','vic41','vic43','vic45','fear6f')
n_ca_common <- c('cp5','cp7','cp8','cp13','cp20')
n_ca_more <- c(n_ca_common,'honcp22','honcp21a','cp21')
n_pv_gtm <- c('pv1','pv2a','pv2b','pv2c','pv2d','pv2e','pv2f',
            'pv2g','pv2h','pv2i','pv2j','pv2k')
n_ex_common <- c('exc2','exc6','exc20','exc11','exc13','exc14',
               'exc15','exc16','exc7')
n_tr_more <- c(tr_common,'pr4','m1','b11','esb48','epp1','epp3','pr4',
             'epn3a','epn3b','epn3c','b11','b37','b14','b15','b19',
             'b46','honb51','venb11','venhonb51',
             'venhonvb10','epp1','epp3','aoj18')
n_w_more <- c(w_common,'inf3a')
n_aut_hnd <- c('dem2','dem11','aut1','jc13','jc10','jc15a',
             'jc16a','honjc17')
n_aut_common <- c('dem2','dem11','jc13','jc10','jc15a')
n_geo_more <- c('pais','estratopri','estratosec','upm','prov','municipio',
              'cluster','tamano','hondistrito')
n_dont_use <- c('idnum','fecha','wt','uniq_id','nationality','vb3n','vb11',
              'leng1')

used <- unique(c(n_fear_more,n_ca_more,n_pv_gtm,n_ex_common,n_tr_more,
                 n_w_more,crit_common,n_aut_hnd,n_geo_more,n_dont_use))

unused_common <- names(lapop.2014.all)[!(names(lapop.2014.all) %in% used)]
unused_gtm <- names(lapop.2014.GTM)[!(names(lapop.2014.GTM) %in% used) & 
                                    !(names(lapop.2014.GTM) %in% unused_common)]
unused_slv <- names(lapop.2014.SLV)[!(names(lapop.2014.SLV) %in% used) & 
                                      !(names(lapop.2014.SLV) %in% unused_common)]
unused_hnd <- names(lapop.2014.HND)[!(names(lapop.2014.HND) %in% used) & 
                                      !(names(lapop.2014.HND) %in% unused_common)]
# common questions with country-specific answers
addback <- c('vb3n','vb11','leng1')
unused_gtm <- c(unused_gtm,addback)
unused_slv <- c(unused_slv,addback)
unused_hnd <- c(unused_hnd,addback)
```

We now have a list of unused variables that are common to all three countries, as well as a list of the unique unused variables for each country. Next, we need to categorize these variables as binary, ordered, or unordered categorical.

```{r}
# Assume a function is binary if only two responses are present
is_binary <- function(data,var) {
  u <- unique(data[,var])
  length(u[u<888888]) == 2
}

bin_common <- unused_common[sapply(unused_common,function(x) 
  is_binary(lapop.2014.all,x))]
bin_gtm <- unused_gtm[sapply(unused_gtm,function(x) 
  is_binary(lapop.2014.GTM,x))]
bin_slv <- unused_slv[sapply(unused_slv,function(x) 
  is_binary(lapop.2014.SLV,x))]
bin_hnd <- unused_hnd[sapply(unused_hnd,function(x) 
  is_binary(lapop.2014.HND,x))]

# Note that, while the indicators vb3n, vb11, and leng1 are used in all three
# countries, the answers are country-specific and shouldn't be used as 
# regional indicators.
unord_common <- c('idiomaq','a4','vic2','vic2aa','aoj22','env1','vb1',
                  'vb4new','vb101','vb20','for1n','for4',
                  'for5','q3c','ocup4a','ocup1a','q11n','etid')
ord_common <- unused_common[!(unused_common %in% bin_common) &
                            !(unused_common %in% unord_common)]
unord_gtm <- c('aoj21','chipart107n','parclien','guaetid2n','leng4','vb3n','vb11','leng1')
ord_gtm <- unused_gtm[!(unused_gtm %in% bin_gtm) &
                        !(unused_gtm %in% unord_gtm)]
unord_slv <- c('esexc16a','elsvb48','pr1','vb3n','vb11','leng1')
ord_slv <- unused_slv[!(unused_slv %in% bin_slv) &
                        !(unused_slv %in% unord_slv)]
unord_hnd <- c('dst1','vb3n','vb11','leng1')
ord_hnd <- unused_hnd[!(unused_hnd %in% bin_hnd) &
                              !(unused_hnd %in% unord_hnd)]
```

The appopriate method for finding correlations will be different for each of these three types of variables. I've defined three functions that take the same inputs and produce similar outputs, so that all can be combined and compared at the end.

Note that all of the functions below set the p-value cutoff, by default, to `1/ncol(data)` This is appropriate in cases (like this one) where we're looking at a large number of variables -- if we have more variables then we're more likely to observe spurious correlations so we need to be stricter. 

```{r}
bin_cor <- function(data,var1,var2,cutoff=0) {
  # Determine the correlation between a variable var1 and a binary variable 
  # var2. This uses a Fisher test conditioned on the values of var2, and will
  # be applicable if var1 is continuous (i.e. one of our composite indices).
  res = data.frame(var=character(),est=numeric(),pval=numeric(),
                   stringsAsFactors=FALSE)
  if (var1 != var2) {
    if (cutoff == 0) {
      cutoff <- 1 / ncol(data)
    }
    tmp <- data.frame(v1=data[,var1],v2=data[,var2])
    is.na(tmp[tmp>800000]) <- TRUE
    tmp$v2 <- tmp$v2 - min(tmp$v2,na.rm=TRUE) # convert to 0-1 scale
    if (sum(tmp$v2==0,na.rm=TRUE) > 1 & sum(tmp$v2==1,na.rm=TRUE) > 1) {
      tt <- t.test(tmp$v1[tmp$v2==0],tmp$v1[tmp$v2==1])
      if (tt$p.value < cutoff) {
        res=data.frame(var=var2,est=tt$estimate[2]-tt$estimate[1],pval=tt$p.value,
                       stringsAsFactors=FALSE)
      }
    }
  }
  res
}

ord_cor <- function(data,var1,var2,cutoff=0) {
  # Determine the correlation between two ordered variables, var1 and var2. 
  # This uses linear regression, and will be applicable when we're dealing
  # with ordered categorical variables or composite indices.
  # As a matter of convention, use the composite index as var1 so that the
  # output makes sense.
  res = data.frame(var=character(),est=numeric(),pval=numeric(),
                   stringsAsFactors=FALSE)
  if (var1 != var2) {
    if (cutoff == 0) {
      cutoff <- 1 / ncol(data)
    }
    tmp <- data.frame(v1=data[,var1],v2=data[,var2])
    is.na(tmp[tmp>800000]) <- TRUE
    reg <- lm(v1 ~ v2,data=tmp)
    p <- summary(reg)$coefficients[2,4]
    if (p < cutoff) {
      res=data.frame(var=var2,est=summary(reg)$coefficients[2,1],pval=p,
                     stringsAsFactors=FALSE)
    }
  }
  res
}

unord_cor <- function(data,var1,var2,cutoff=0) {
  # Determine the correlation between an ordered variable var1 and an unordered
  # categorical variable var2. Do this by creating a binary variable for each
  # possible value of var2 and calling bin_cor().
  vals <- na.omit(unique(data[data[,var2] < 800000,var2]))
  if (cutoff == 0) {
    cutoff <- 1 / ncol(data)
  }
  tmp <- data.frame(v1=data[,var1])
  is.na(tmp[tmp>800000]) <- TRUE
  for (x in vals) {
    tmp[,paste(var2,x,sep='_')] <- as.numeric(data[,var2] == x)
  }
  ldply(names(tmp),function(x) bin_cor(tmp,'v1',x,cutoff))
}
```

**Region-wide**

Now we're going to look at the region-wide correlations. First, define a data frame that will contain all of the variables that were not used to generate composite indices, along with the composite indices.

```{r}
a <- lapop.2014.all[,unused_common]
is.na(a[a>800000]) <- TRUE
# add in all of the composite indices
a$fear_idx <- fear_all
a$ca_idx <- ca_all
a$tr_idx <- tr_all
a$w_idx <- w_all
a$crit_idx <- crit_all
a$aut_idx <- aut_all
idxs <- c('fear_idx','ca_idx','tr_idx','w_idx','crit_idx','aut_idx')
```

Next, we'll use the functions defined above to get all of our correlations. We'll keep those with p-values below the cutoff of `r 1/ncol(a)`

```{r}
cor_all_bin <- ldply(bin_common, function(x) bin_cor(a,'fear_idx',x))
cor_all_ord <- ldply(ord_common, function(x) ord_cor(a,'fear_idx',x))
cor_all_idx <- ldply(idxs, function(x) ord_cor(a,'fear_idx',x))
cor_all_unord <- ldply(unord_common, function(x) unord_cor(a,'fear_idx',x))
cor_all <- rbind(cor_all_bin,cor_all_ord,cor_all_unord,cor_all_idx)
cor_all <- cor_all[order(cor_all$pval),]
```

While we have a list of which values of unordered categorical variables are significantly correlated with the fear index, the variables aren't formatted in a way that will be useful for multiple regression. We'll do this by creating dummy variables, with the same names as the `var` column in `cor_all`. (For example, `ocup1a_4` for `ocup1a == 4`.)

```{r}
unord_vars <- ldply(cor_all_unord$var, function(x)
  data.frame(var=strsplit(x,'_')[[1]][1],
             val=as.numeric(strsplit(x,'_')[[1]][2]),
             str=x,
             stringsAsFactors=FALSE))
for (i in 1:nrow(unord_vars)) {
  a[,unord_vars$str[i]] <- as.numeric(a[,unord_vars$var[i]] == unord_vars$val[i])
}
a2 <- a[,cor_all$var]
a2$fear_idx <- a$fear_idx
```

One thing that can cause trouble in multiple regression is missing values -- we can only include rows in our dataframe that have a non-NA value for all of the variables compared. This means that as we add in more and more variables, we throw out more and more data, and our final results can be biased. We'll get around this by using multiple imputation. We'll need to create a predictor matrix (named `pm`) to account for the fact that we don't want to be using our fear index to impute values of the other variables -- this would be sort of circular logic!

```{r}
pm <- 1 - diag(ncol(a2))
pm[,which(names(a2)=='fear_idx')] <- 0
b <- mice(a2,printFlag=F,predictorMatrix=pm)
```

We can get to a decent set of variables using stepwise regression. What we'll do is run backward stepwise regression on each imputed dataset and keep only the ones that appear for all five sets. One can get pretty much the exact same result using forward regression (I checked).

```{r}
b1 <- complete(b,1)
b2 <- complete(b,2)
b3 <- complete(b,3)
b4 <- complete(b,4)
b5 <- complete(b,5)
lm1 <- lm(fear_idx ~ .,data=na.omit(b1))
lm2 <- lm(fear_idx ~ .,data=na.omit(b2))
lm3 <- lm(fear_idx ~ .,data=na.omit(b3))
lm4 <- lm(fear_idx ~ .,data=na.omit(b4))
lm5 <- lm(fear_idx ~ .,data=na.omit(b5))
library(MASS)
step1 <- stepAIC(lm1,trace=F) # final AIC = -967.9723
step2 <- stepAIC(lm2,trace=F) # final AIC = -946.0618
step3 <- stepAIC(lm3,trace=F) # final AIC = -951.4168
step4 <- stepAIC(lm4,trace=F) # final AIC = -951.9888
step5 <- stepAIC(lm5,trace=F) # final AIC = -941.9365
counts <- as.data.frame(table(c(names(coef(step1)),names(coef(step2)),
                                names(coef(step3)),names(coef(step4)),
                                names(coef(step5)))),stringsAsFactors=F)
counts[counts$Freq==5,'Var1']
nrow(counts[counts$Freq==5,]) / nrow(counts) # 55% appear in all 5

b.reg2 <- with(data=b,exp=lm(fear_idx ~ a4_5 + aoj22_1 + aut_idx + ca_idx + 
                               clien1n + d6 + ed2 + for1n_7 + for4_4 + 
                               for5_6 + for6 + it1 + mil10c + ocup4a_3 + 
                               pol1 + pole2n + q3c_5 + sd3new2 + tr_idx + 
                               ur + vb20_4 + w14a + www1))
s <- summary(pool(b.reg2))
plotme <- data.frame(s[-1,c(1,5)])
plotme$log_p <- -log10(plotme[,2])
m <- max(plotme$log_p[plotme$log_p < Inf])
plotme$log_p[plotme$log_p > m] <- 1.1*m
scale <- c(1,1,1,1,1,10,8,1,1,1,4,4,4,1,4,4,1,4,1,1,1,1,5)
plotme$est <- plotme$est * scale
n = 0.02 # nudge factor
nx = c(0,0,0,0,0,0,10*n,-5*n,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0)
ny = c(0,n,0,0,0,-n,2*n,n,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0)
ggplot(data=plotme,aes(x=log_p,y=est,color=est,label=rownames(plotme))) +
  geom_point(size=20,alpha=0.2) +
  geom_text(color='black',aes(x=log_p+nx, y=est+ny)) +
  geom_segment(x=-log10(0.01),xend=-log10(0.01),
               y=1.2*min(plotme$est),yend=0.95*max(plotme$est),color='red') +
  annotate("text",x=-log10(0.01)+0.2,y=max(plotme$est),label='p=0.01',
           color='red') +
  geom_segment(x=-log10(0.05),xend=-log10(0.05),
               y=1.2*min(plotme$est),yend=0.95*max(plotme$est),color='blue') +
  annotate("text",x=-log10(0.05)-0.2,y=max(plotme$est),label='p=0.05',
           color='blue') +
  scale_color_gradientn(colours=rainbow(4)) +
  theme_classic() +
  theme(legend.position='none') +
  xlab('Significance (-log(p))') +
  ylab('Influence on fear index')
```

Variable  |Description                                    |Effect
----------|-----------------------------------------------|-------
`it1`     |Interpersonal trust                            |Less trust = more fear
`ur`      |Urbanization                                   |Urban = more fear  
`tr_idx`  |Trust in government                            |Less trust = more fear
`ca_idx`  |Community activity                             |More involvement = more fear
`aut_idx` |Authoritarianism                               |More authoritarian = more fear
`q3c = 5` |Evangelical/Protestant                         |More fearful
`pole2n`  |Satisfaction with police performance           |Less satisfied = more fearful
`for6`    |Influence of China in country                  |More influence = more fear
`pol1`    |Interest in politics                           |More interest = more fear
`a4 = 5`  |Believes crime is the most serious problem     |More fearful
`mil10c`  |Trustworthiness of Iranian government          |More trustworthy = more fear
`aoj22=1` |Reduce crime via preventative measures         |Less fearful
`clien1n` |Knows someone offered benefit for vote         |Yes = more fear
`ocup4a=3`|Actively looking for a job                     |More fearful
`www1`    |Internet usage                                 |Frequent = more fear
`w14a`    |Abortion justified when mother's health at risk|Yes = more fear
`sd3new2` |Satisfaction with public schools               |Dissatisfaction = more fear
`ed2`     |Education level of mother                      |More educated = more fear
`vb20 = 4`|Plans to leave ballot blank in next election   |More fearful
`d6`      |Approval of same-sex marriage                  |Approving = more fear 
`for5 = 6`|Best model for future development is Russia    |More fearful

```{r}
geo <- c('pais','estratopri','estratosec','prov','municipio')
my_geo <- lapop.2014.all[,geo]
my_geo$muni_uniq <- 10000*my_geo$pais + my_geo$municipio
my_geo$fear <- fear_all
muni_avg <- ddply(my_geo,~muni_uniq,summarize,x=mean(fear))
my_geo$muni_avg <- muni_avg$x[match(my_geo$muni_uniq,muni_avg$muni_uniq)]

b.reggeo <- with(data=b,exp=lm(fear_idx ~ a4_5 + aoj22_1 + aut_idx + ca_idx +
                                 clien1n + d6 + ed2 + for1n_7 + for4_4 + 
                               for5_6 + for6 + it1 + mil10c + ocup4a_3 + 
                               pol1 + pole2n + q3c_5 + sd3new2 + tr_idx + 
                               ur + vb20_4 + w14a + www1 + my_geo$muni_avg))
s_geo <- summary(pool(b.reggeo))
plotme$geo_est <- s_geo[c(-1,-25),1]
plotme$geo_logp <- -log10(s_geo[c(-1,-25),5])
m <- max(plotme$geo_logp[plotme$geo_logp < Inf])
plotme$geo_logp[plotme$geo_logp > m] <- 1.1*m
plotme$geo_est <- plotme$geo_est * scale
nx = c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,5*n,0,0,0,0,0,0)
ny = c(0,0,0,0,0,2*n,2.5*n,0,0,0,-n,0,0,0,0,0,0,0,0,0,-1.5*n,-2*n,0)
ggplot(data=plotme,aes(x=geo_logp,y=geo_est,color=geo_est,
                       label=rownames(plotme))) +
  geom_point(size=20,alpha=0.2) +
  geom_segment(aes(x=plotme$log_p,y=plotme$est,
                   xend=plotme$geo_logp,yend=plotme$geo_est,
                   color=geo_est)) +
  geom_text(color='black',aes(x=geo_logp+nx, y=geo_est+ny)) +
  geom_segment(x=-log10(0.01),xend=-log10(0.01),
               y=1.2*min(plotme$est),yend=0.95*max(plotme$est),color='red') +
  annotate("text",x=-log10(0.01)+0.2,y=max(plotme$est),label='p=0.01',
           color='red') +
  geom_segment(x=-log10(0.05),xend=-log10(0.05),
               y=1.2*min(plotme$est),yend=0.95*max(plotme$est),color='blue') +
  annotate("text",x=-log10(0.05)-0.2,y=max(plotme$est),label='p=0.05',
           color='blue') +
  scale_color_gradientn(colours=rainbow(4)) +
  theme_classic() +
  theme(legend.position='none') +
  xlab('Significance (-log(p))') +
  ylab('Influence on fear index')
```

When we include the municipality averages in our multiple correlation, we're asking about the factors that make a person more fearful than others in their community. The straight lines attached to each circle illustrate how much they moved from the previous plot -- for example, `tr_idx` became dramatically less significant while `mil10c` hardly moved at all. This suggests that people's trust in their own government (`tr_idx`) correlates strongly with their municipality, while views on the Iranian government (`mil10c`) serve more to differentiate people within the same community. In general, most variables have a smaller estimated influence, but most significant variables remain significant -- some actually even increase. This means that we're measuring more than just attributes of places; these variables are telling us about people within heterogeneous places.

# Guatemala #

Next, let's see if any of the Guatemala-specific variables are interesting. 

```{r}
g <- lapop.2014.GTM[,c(unused_common,unused_gtm)]
is.na(g[g>800000]) <- TRUE
g$fear_idx <- fear_gtm
g$ca_idx <- ca_all[lapop.2014.all$pais==2] # there is no ca_gtm
g$tr_idx <- tr_gtm
g$w_idx <- w_gtm
g$crit_idx <- crit_all[lapop.2014.all$pais==2]
g$aut_idx <- aut_all[lapop.2014.all$pais==2]
g$pv_idx <- pv_gtm
idxs <- c('fear_idx','ca_idx','tr_idx','w_idx','crit_idx','aut_idx','pv_idx')

cor_gtm_bin <- ldply(c(bin_common,bin_gtm), function(x) bin_cor(g,'fear_idx',x))
cor_gtm_ord <- ldply(c(ord_common,ord_gtm), function(x) ord_cor(g,'fear_idx',x))
cor_gtm_idx <- ldply(idxs, function(x) ord_cor(g,'fear_idx',x))
cor_gtm_unord <- ldply(c(unord_common,unord_gtm), function(x) 
  unord_cor(g,'fear_idx',x))
cor_gtm <- rbind(cor_gtm_bin,cor_gtm_ord,cor_gtm_unord,cor_gtm_idx)
cor_gtm <- cor_gtm[order(cor_gtm$pval),]
```

We definitely see a few variables that weren't on the menu before; some of these we might want to add into the indices we have.

```{r}
unord_vars_g <- ldply(cor_gtm_unord$var, function(x)
  data.frame(var=strsplit(x,'_')[[1]][1],
             val=as.numeric(strsplit(x,'_')[[1]][2]),
             str=x,
             stringsAsFactors=FALSE))
for (i in 1:nrow(unord_vars_g)) {
  g[,unord_vars_g$str[i]] <- as.numeric(g[,unord_vars_g$var[i]] == unord_vars_g$val[i])
}
g2 <- g[,cor_gtm$var]
g2$fear_idx <- g$fear_idx

pm_g <- 1 - diag(ncol(g2))
pm_g[,which(names(g2)=='fear_idx')] <- 0
g_imp <- mice(g2,printFlag=F,predictorMatrix=pm_g)
```

Let's cut right to the chase and use stepwise regression on our multiply-imputed data:

```{r}
lm1g <- lm(fear_idx ~ .,data=na.omit(complete(g_imp,1)))
lm2g <- lm(fear_idx ~ .,data=na.omit(complete(g_imp,2)))
lm3g <- lm(fear_idx ~ .,data=na.omit(complete(g_imp,3)))
lm4g <- lm(fear_idx ~ .,data=na.omit(complete(g_imp,4)))
lm5g <- lm(fear_idx ~ .,data=na.omit(complete(g_imp,5)))
step1g <- stepAIC(lm1g,direction="both",trace=F) # final AIC = -1016.8753
step2g <- stepAIC(lm2g,direction="both",trace=F) # final AIC = -997.2276
step3g <- stepAIC(lm3g,direction="both",trace=F) # final AIC = -1014.1022
step4g <- stepAIC(lm4g,direction="both",trace=F) # final AIC = -1005.7721
step5g <- stepAIC(lm5g,direction="both",trace=F) # final AIC = -993.3492
counts_g <- as.data.frame(table(c(names(coef(step1g)),names(coef(step2g)),
                                names(coef(step3g)),names(coef(step4g)),
                                names(coef(step5g)))),stringsAsFactors=F)
counts_g[counts_g$Freq==5,'Var1']
nrow(counts[counts_g$Freq==5,]) / nrow(counts_g) # 55% appear in all 5

g.reg <- with(data=g_imp,exp=lm(fear_idx ~ a4_27 + a4_5 + aoj21_1 + aoj21_2 + 
                                  aoj21_8 + aut_idx + d6 + fear6e + for4_1 + 
                                  guaetid2n_1 + guaetid2n_16 + it1 + leng4_1 + 
                                  per9 + pn4 + pole2n + pv_idx + pv3 + q3c_2 + 
                                  tr_idx + ur + vb20_3 + vb20_4 + w_idx))
summary(g.reg)
```

TODO: Add scatter plot visualization for GTM.

So what are the strongest correlates in Guatemala?

- `fear6e`: Insecurity on public transportation (add to index?)
- `aoj21 == 2`: Gangs are biggest threat to security
- `ur`: Urbanizaton
- `aoj21 == 8`: Biggest threat = none of the above (much less fearful)
- `it1`: Interpersonal trust
- `aut_idx`: Authoritarianism index
- `pv_idx`: Political violence index
- `leng4 == 1`: Parents' language (Spanish vs. indigenous)
- `aoj21 == 8`: Biggest threat = your neighbors (less fearful)
- `tr_idx`: Trust in government index
- `vb20 == 4`: Plans to vote but leave ballot blank
- `a4 == 5`: Crime the biggest problem
- `guaetid2n == 16`: Ethnic group = Q'eqchi'
- `a4 == 27`: Most serious problem = lack of security
- `pv3`: Believes violence was used in 2015 elections
- `d6`: Approval of same-sex marriage (higher = more fear)
- `pole2n`: Satisfaction with police performance
- `guaetid2n == 1`: Ethnic group = Achí
- `pn4`: Satisfaction with democracy

We see some new themes, as well as some overlap with the region-wide trends.

# El Salvador #

```{r}
s <- lapop.2014.SLV[,c(unused_common,unused_slv)]
is.na(s[s>800000]) <- TRUE
s$fear_idx <- fear_slv
s$ca_idx <- ca_all[lapop.2014.all$pais==3] # there is no ca_gtm
s$tr_idx <- tr_slv
s$w_idx <- w_slv
s$crit_idx <- crit_all[lapop.2014.all$pais==3]
s$aut_idx <- aut_all[lapop.2014.all$pais==3]
idxs <- c('fear_idx','ca_idx','tr_idx','w_idx','crit_idx','aut_idx')

cor_slv_bin <- ldply(c(bin_common,bin_slv), function(x) bin_cor(s,'fear_idx',x))
cor_slv_ord <- ldply(c(ord_common,ord_slv), function(x) ord_cor(s,'fear_idx',x))
cor_slv_idx <- ldply(idxs, function(x) ord_cor(s,'fear_idx',x))
cor_slv_unord <- ldply(c(unord_common,unord_slv), function(x) 
  unord_cor(s,'fear_idx',x))
cor_slv <- rbind(cor_slv_bin,cor_slv_ord,cor_slv_unord,cor_slv_idx)
cor_slv <- cor_slv[order(cor_slv$pval),]
head(cor_slv,15)
```

TODO: elsdiso18 and elsdiso19 should probably be made part of the index for SLV, given their topics and incredibly low p-values. Rather than adding on to the end of fear.Rmd, rename it as fear_HND.Rmd and make a new copy for GTM.

We see a lot of variables that should look familiar, plus some new ones:
- elsvb55c: Uses newspaper as info source for voting decision (yes=more fearful)
- elsvb55d: Uses internet...
- elsvb55e: Uses facebook...
- elsvb55i: Uses candidate forums
- vicbar7f: # of occurrences of murders (TODO: should also be part of index)
- info1: knows about Law of Access to Public Information (yes=more fear)

A lot of these seem like measures for political engagement, with the more-engaged and better-informed people being more concerned about crime.

```{r}
unord_vars_s <- ldply(cor_slv_unord$var, function(x)
  data.frame(var=strsplit(x,'_')[[1]][1],
             val=as.numeric(strsplit(x,'_')[[1]][2]),
             str=x,
             stringsAsFactors=FALSE))
for (i in 1:nrow(unord_vars_s)) {
  s[,unord_vars_s$str[i]] <- as.numeric(s[,unord_vars_s$var[i]] == unord_vars_s$val[i])
}
s2 <- s[,cor_slv$var]
s2$fear_idx <- s$fear_idx

pm_s <- 1 - diag(ncol(s2))
pm_s[,which(names(s2)=='fear_idx')] <- 0
s_imp <- mice(s2,printFlag=F,predictorMatrix=pm_s)
```

Let's cut right to the chase and use stepwise regression on our multiply-imputed data:

```{r}
lm1s <- lm(fear_idx ~ .,data=na.omit(complete(s_imp,1)))
lm2s <- lm(fear_idx ~ .,data=na.omit(complete(s_imp,2)))
lm3s <- lm(fear_idx ~ .,data=na.omit(complete(s_imp,3)))
lm4s <- lm(fear_idx ~ .,data=na.omit(complete(s_imp,4)))
lm5s <- lm(fear_idx ~ .,data=na.omit(complete(s_imp,5)))
step1s <- stepAIC(lm1s,direction="both",trace=F) # final AIC = 
step2s <- stepAIC(lm2s,direction="both",trace=F) # final AIC = 
step3s <- stepAIC(lm3s,direction="both",trace=F) # final AIC = 
step4s <- stepAIC(lm4s,direction="both",trace=F) # final AIC = 
step5s <- stepAIC(lm5s,direction="both",trace=F) # final AIC = 
counts_s <- as.data.frame(table(c(names(coef(step1s)),names(coef(step2s)),
                                names(coef(step3s)),names(coef(step4s)),
                                names(coef(step5s)))),stringsAsFactors=F)
counts_s[counts_s$Freq==5,'Var1']
nrow(counts_s[counts_s$Freq==5,]) / nrow(counts_s) # 55% appear in all 5

s.reg <- with(data=s_imp,exp=lm(fear_idx ~ ca_idx + cpss1 + elsdiso18 + elsdiso19 + elsvb55a + elsvb55e + gi0 + infra2 + it1 + pol1 + q11n_6 + sexi + tr_idx))
summary(s.reg)
```

TODO: Add scatter plot visualization here also.

In El Salvador, our most significant correlations (other than the elsdisos, which I need TODO something with) are:

- `ca_idx`



**Honduras**

```{r}
h <- lapop.2014.HND[,c(unused_common,unused_hnd)]
is.na(h[h>800000]) <- TRUE
h$fear_idx <- fear_hnd
h$ca_idx <- ca_hnd 
h$tr_idx <- tr_hnd
h$w_idx <- w_hnd
h$crit_idx <- crit_hnd
h$aut_idx <- aut_hnd
idxs <- c('fear_idx','ca_idx','tr_idx','w_idx','crit_idx','aut_idx')

cor_hnd_bin <- ldply(c(bin_common,bin_hnd), function(x) bin_cor(h,'fear_idx',x))
cor_hnd_ord <- ldply(c(ord_common,ord_hnd), function(x) ord_cor(h,'fear_idx',x))
cor_hnd_idx <- ldply(idxs, function(x) ord_cor(h,'fear_idx',x))
cor_hnd_unord <- ldply(c(unord_common,unord_hnd), function(x) 
  unord_cor(h,'fear_idx',x))
cor_hnd <- rbind(cor_hnd_bin,cor_hnd_ord,cor_hnd_unord,cor_hnd_idx)
cor_hnd <- cor_hnd[order(cor_hnd$pval),]
```


```{r}
unord_vars_h <- ldply(cor_hnd_unord$var, function(x)
  data.frame(var=strsplit(x,'_')[[1]][1],
             val=as.numeric(strsplit(x,'_')[[1]][2]),
             str=x,
             stringsAsFactors=FALSE))
for (i in 1:nrow(unord_vars_h)) {
  h[,unord_vars_h$str[i]] <- as.numeric(h[,unord_vars_h$var[i]] == unord_vars_h$val[i])
}
h2 <- h[,cor_hnd$var]
h2$fear_idx <- h$fear_idx

pm_h <- 1 - diag(ncol(h2))
pm_h[,which(names(h2)=='fear_idx')] <- 0
h_imp <- mice(h2,printFlag=F,predictorMatrix=pm_h)
```

Let's cut right to the chase and use stepwise regression on our multiply-imputed data:

```{r}
lm1h <- lm(fear_idx ~ .,data=na.omit(complete(h_imp,1)))
lm2h <- lm(fear_idx ~ .,data=na.omit(complete(h_imp,2)))
lm3h <- lm(fear_idx ~ .,data=na.omit(complete(h_imp,3)))
lm4h <- lm(fear_idx ~ .,data=na.omit(complete(h_imp,4)))
lm5h <- lm(fear_idx ~ .,data=na.omit(complete(h_imp,5)))
step1h <- stepAIC(lm1h,direction="both",trace=F) # final AIC = 
step2h <- stepAIC(lm2h,direction="both",trace=F) # final AIC = 
step3h <- stepAIC(lm3h,direction="both",trace=F) # final AIC = 
step4h <- stepAIC(lm4h,direction="both",trace=F) # final AIC = 
step5h <- stepAIC(lm5h,direction="both",trace=F) # final AIC = 
counts_h <- as.data.frame(table(c(names(coef(step1h)),names(coef(step2h)),
                                names(coef(step3h)),names(coef(step4h)),
                                names(coef(step5h)))),stringsAsFactors=F)
counts_h[counts_h$Freq==5,'Var1']
nrow(counts_h[counts_h$Freq==5,]) / nrow(counts_h) # 67% appear in all 5

h.reg <- with(data=h_imp,exp=lm(fear_idx ~ a4_22 + aut_idx + b20 + clien1na + coer1 + gix4 + ico2 + it1 + mil10c + ocup4a_3 + per4 + pole2n + pr3e + prot3 + sd3new2 + tr_idx + ur))
h_s <- summary(h.reg)
```

Now let's visualize the same way we did with the regional correlations.

```{r}
plotme_h <- data.frame(h_s[-1,c(1,5)])
plotme_h$log_p <- -log10(plotme_h[,2])
m <- max(plotme_h$log_p[plotme_h$log_p < Inf])
plotme_h$log_p[plotme_h$log_p > m] <- 1.1*m
# TODO: these numbers aren't right anymore; check for the real ones
scale <- c(1,1,1,1,1,10,8,1,1,1,4,4,4,1,4,4,1,4,1,1,1,1,5)
plotme_h$est <- plotme_h$est * scale
# TODO: Add title to plot (all of them)
ggplot(data=plotme_h,aes(x=log_p,y=est,color=est,label=rownames(plotme_h))) +
  geom_point(size=20,alpha=0.2) +
  geom_text(color='black') +
  scale_color_gradientn(colours=rainbow(4)) +
  theme_classic() +
  theme(legend.position='none') +
  xlab('Significance (-log(p))') +
  ylab('Influence on fear index')
```
