---
title: "Wealth time series"
author: "Craig Jolley"
date: "January 27, 2016"
output: html_document
---

I'm curious which of the variables in our wealth index have historical information available, and how these vary over time.

```{r}
setwd("C:/Users/Craig/Desktop/SMA/VSFS/LAPOP-SMA/index_ts")
source('../utils/make_trends.R')
source('../indices/make_indices.R')
my_data <- lapop.trends
# In 2004, some people answered "2" or "3" to r1; I'm not sure what this means. Maybe 
# they have multiple TVs. Change these to 1, since that was the usage in previous years.
my_data[!is.na(my_data$r1) & my_data$r1 > 1,'r1'] <- 1
w_common_ts <- intersect(names(my_data),w_common)
w_ts_idx <- make_idx(my_data,w_common_ts,sgn=1)
```

We can visualize changes in wealth (as measured in terms of assets) for each country:

```{r}
df3 <- data.frame(year=lapop.trends$year,pais=lapop.trends$pais,w_idx=w_ts_idx)
plotme <- ddply(df3,.(year,pais),summarize,mean=mean(w_idx),
      lo=t.test(w_idx)$conf.int[1],hi=t.test(w_idx)$conf.int[2])
pais_text <- c('','Guatemala','El Salvador','Honduras')
plotme$pais <- pais_text[plotme$pais]
ggplot(plotme,aes(x=year,y=mean,group=pais,color=pais)) +
  geom_point(size=5) +
  geom_line(size=3) +
  geom_errorbar(aes(ymin=lo,ymax=hi,width=0.5)) +
  ylab('Average wealth index') +
  scale_x_continuous(limits=c(2004,2014)) +
  theme_classic()
```

On average things are getting better, most dramatically in Honduras. 

I'm not sure that `estratopri` labels will be consistent for files loaded the way I've done it here; if I want to drill down to more geographic resolution I should use LAPOP's "Grand Merge" dataset and subset out the Northern Triangle countries.