---
title: "LAPOP Summary"
author: "Craig Jolley"
date: "January 20, 2016"
output: html_document
---

Calculating some high-level summary points for the LAPOP datasets.

```{r,message=FALSE}
library(plyr)
library(ggplot2)
library(lubridate)
library(VennDiagram)
library(RColorBrewer)
source('../utils/make_trends.R')

# Load data
setwd("C:/Users/Craig/Desktop/SMA/VSFS/LAPOP-SMA/viz")
lapop.2014.HND <- read.csv("../../HND-2014.csv",stringsAsFactors=FALSE)
lapop.2014.GTM <- read.csv("../../GTM-2014.csv",stringsAsFactors=FALSE)
lapop.2014.SLV <- read.csv("../../SLV-2014.csv",stringsAsFactors=FALSE)
# Get useful non-numeric variables
date_time <- rbind(lapop.2014.GTM[,c('pais','fecha','ti')],
                   lapop.2014.SLV[,c('pais','fecha','ti')],
                   lapop.2014.HND[,c('pais','fecha','ti')])


fixdata <- function(x) {
  y <- x
  # vic1exta asks about the frequency of crime victimization and is NR if 
  # respondant was never victimized; change this to zero
  y$vic1exta[y$vic1exta == 999999] <- 0
  # vicbar1f is similar; asks about the number of occurrences of burglaries
  y$vicbar1f[y$vicbar1f == 999999] <-  0
  # vicbar7f asks about the number of occurences of murders (SLV only)
  if ('vicbar7f' %in% names(y)) {
    y$vicbar7f[y$vicbar7f == 999999] <- 0
  }
  # set uniq_id to 0 (so it doesn't become NA)
  y$uniq_id <- 0
  # somehow ages got imported as strings
  y$q2 <- as.numeric(y$q2)
  # set non-responses to NA
  y[y > 800000] <- NA
  y
}

lapop.2014.GTM <- fixdata(lapop.2014.GTM)
lapop.2014.SLV <- fixdata(lapop.2014.SLV)
lapop.2014.HND <- fixdata(lapop.2014.HND)

# create a combined dataset
common <- Reduce(intersect,list(names(lapop.2014.GTM),names(lapop.2014.SLV),
                                names(lapop.2014.HND)))
lapop.2014.all <- rbind(lapop.2014.GTM[,common],lapop.2014.SLV[,common],
                        lapop.2014.HND[,common])

```

# Basic features #

```{r,message=FALSE,warning=FALSE}
basic <- data.frame(country=c('Guatemala','El Salvador','Honduras'),
                    nq=c(ncol(lapop.2014.GTM),ncol(lapop.2014.SLV),
                         ncol(lapop.2014.HND)),
                    nr=c(nrow(lapop.2014.GTM),nrow(lapop.2014.SLV),
                         nrow(lapop.2014.HND)))
# num of questions
ggplot(basic,aes(x=country,y=nq,fill=country)) +
  geom_bar(stat="identity") +
  ggtitle("Number of questions") +
  geom_text(aes(y=nq,label=nq,vjust=1.5),size=10) +
  scale_fill_manual(values=brewer.pal(3,"Dark2")) +
  theme_classic() +
  theme(legend.position='none',
        axis.ticks=element_blank(),
        axis.text.y=element_blank(),
        axis.line=element_blank(),
        axis.title=element_blank(),
        text=element_text(size=20))

# num of responses
ggplot(basic,aes(x=country,y=nr,fill=country)) +
  geom_bar(stat="identity") +
  ggtitle("Number of responses") +
  geom_text(aes(y=nr,label=nr,vjust=1.5),size=10) +
    scale_fill_manual(values=brewer.pal(3,"Dark2")) +
  theme_classic() +
  theme(legend.position='none',
        axis.ticks=element_blank(),
        axis.text.y=element_blank(),
        axis.line=element_blank(),
        axis.title=element_blank(),
        text=element_text(size=20))
# Venn diagram of overlapping questions
plot.new()
draw.triple.venn(ncol(lapop.2014.GTM),ncol(lapop.2014.SLV),ncol(lapop.2014.GTM),
                 sum(names(lapop.2014.GTM) %in% names(lapop.2014.SLV)),
                 sum(names(lapop.2014.SLV) %in% names(lapop.2014.HND)),
                 sum(names(lapop.2014.GTM) %in% names(lapop.2014.HND)),
                 ncol(lapop.2014.all),
                 category=c('Guatemala','El Salvador','Honduras'),
                 fill=brewer.pal(3,"Dark2")[c(2,1,3)],
                 cex=rep(3,7),cat.cex=rep(3,3),
                 alpha=rep(0.1,3),col=rep(NA,3),
                 cat.just=list(c(0,0.5),c(1,0.5),c(0.5,0)),
                 euler.d=TRUE,scaled=TRUE)

# Questions with time series

# Create test data.
ts = data.frame(count=c(ncol(lapop.trends), ncol(lapop.2014.all)-ncol(lapop.trends)), 
                category=c('trend','notrend'))

# Add addition columns, needed for drawing with geom_rect.
ts$fraction = ts$count / sum(ts$count)
ts = ts[order(ts$fraction), ]
ts$ymax = cumsum(ts$fraction)
ts$ymin = c(0, head(ts$ymax, n=-1))
ggplot(ts, aes(fill=category, ymax=ymax, ymin=ymin, xmax=4, xmin=2)) +
     geom_rect() +
     coord_polar(theta="y") +
     xlim(c(0, 4)) +
     annotate('text',x=4,y=0.25,
              label='2004-2014 time series',size=10,hjust=0.85) +
     annotate('text',x=4,y=0.82,
              label='No time series',size=10,hjust=0.3) +
     theme_classic() +
     theme(legend.position='none',
           axis.ticks=element_blank(),
           axis.text=element_blank(),
           axis.line=element_blank(),
           axis.title=element_blank())
```

# Missing values #

```{r}
# Average missing values
mean(is.na(lapop.2014.GTM)) # 8.9% in GTM
mean(is.na(lapop.2014.SLV)) # 9.0% in SLV
mean(is.na(lapop.2014.HND)) # 9.9% in HND
mean(is.na(lapop.2014.all)) # 8.2% in overlap
# Missing values per row
missing_row <- rbind(data.frame(x=rowMeans(is.na(lapop.2014.GTM)),country='GTM'),
                     data.frame(x=rowMeans(is.na(lapop.2014.SLV)),country='SLV'),
                     data.frame(x=rowMeans(is.na(lapop.2014.HND)),country='HND'),
                     data.frame(x=rowMeans(is.na(lapop.2014.all)),country='all'))
ggplot(missing_row,aes(x=x,fill=country)) +
  geom_density(alpha=0.2) +
  theme_classic() +
  xlab('Missing values per row') +
  scale_fill_manual(values=brewer.pal(4,"Dark2")[c(2,1,3,4)]) +
  scale_x_continuous(limits=c(0,0.3)) +
  theme(axis.ticks=element_blank(),
        axis.text.y=element_blank(),
        axis.line=element_blank(),
        axis.title.y=element_blank(),
        text=element_text(size=20))

head(sort(colMeans(is.na(lapop.2014.GTM)),decreasing=TRUE))
head(sort(colMeans(is.na(lapop.2014.SLV)),decreasing=TRUE))
head(sort(colMeans(is.na(lapop.2014.HND)),decreasing=TRUE))
head(sort(colMeans(is.na(lapop.2014.all)),decreasing=TRUE))
quantile(colMeans(is.na(lapop.2014.GTM)))
quantile(colMeans(is.na(lapop.2014.SLV)))
quantile(colMeans(is.na(lapop.2014.HND)))
quantile(colMeans(is.na(lapop.2014.all)))
```

The most common missing responses are for:

- `vb101`: Reason for nullifying ballot (98.3%)
- `exc14`: Paid a bribe at the courts (95.6%)
- `vic2`: Type of victimization (82.3%)
- `vic2aa`: Location of victimization (82.2%)
- `exc11`: Bribe required to process document in municipal government (77.5%)
- `vb4new`: Reason for not voting (74.7%)

In general, questions with many missing answers were requesting information about fairly unusual experiences. In all cases, 50% of questions have missing-answer rates of less than 1-2%.

# Survey dates #

For some reason, a single response in Guatemala was from 1/1/2012, while the rest are from the spring of 2014

```{r}
Sys.setlocale("LC_TIME", "usa")
date_time <- date_time[-1407,] # remove outlier
date_time$date <- dmy(date_time$fecha)
countries <- c(NA,'Guatemala','El Salvador','Honduras')
date_time$country <- countries[date_time$pais]
ggplot(date_time,aes(x=date,fill=country)) +
  geom_density(alpha=0.2,adjust=0.25) +
  scale_fill_manual(values=brewer.pal(3,"Dark2")) +
  theme_classic() +
  xlab('Survey date (2014)') +
  theme(axis.ticks=element_blank(),
        axis.text.y=element_blank(),
        axis.line=element_blank(),
        axis.title.y=element_blank(),
        text=element_text(size=20))


ggplot(date_time,aes(x=ti,fill=country)) +
  geom_density(alpha=0.2,adjust=0.25) +
  scale_x_continuous(limits=c(15,90)) +
  scale_fill_manual(values=brewer.pal(3,"Dark2")) +
  theme_classic() +
  xlab('Survey time (minutes)') +
  theme(axis.ticks=element_blank(),
        axis.text.y=element_blank(),
        axis.line=element_blank(),
        axis.title.y=element_blank(),
        text=element_text(size=20))
```

Most people took 30-60 minutes to complete the survey.

# Basic Demographics #

```{r}
mean(lapop.2014.all$sex==2) # 51.8% female
ages <- lapop.2014.all[,c('pais','q2')]
ages$country <- countries[ages$pais]
ggplot(ages,aes(x=q2,fill=country)) +
  geom_density(alpha=0.2) +
  theme_classic() +
  xlab('Age (years)') +
  scale_fill_manual(values=brewer.pal(3,"Dark2")) +
  theme(axis.ticks=element_blank(),
        axis.text.y=element_blank(),
        axis.line=element_blank(),
        axis.title.y=element_blank(),
        text=element_text(size=20))

